### 2019-11-25 Mathematical Statistics/ms_ch_02.rmd
ms, ch3
ms, ch3

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Exploratory Data Analysis}
   \rfoot{\color{headergrey}Chapter 2}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(copula, quietly = TRUE, warn.conflicts = FALSE)
library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(tseries, quietly = TRUE, warn.conflicts = FALSE)
library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
library(EnvStats, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 2

### 2.1

Compute the mean $\bar{x}$ and median, $m$ ,of the six numbers: 3, 5, 8, 15, 20, 21, 24, then apply the natural log to the data.

```{r, c2_1, echo = T}

x <- c(3, 5, 8, 15, 20, 21, 24)
xt <- log(x)

```

Does $\bar{x}$ = $\tilde{x}$ ?

```{r echo = T}

log(mean(x)) == mean(xt)

```

Does $m$ = $\tilde{m}$ ?

```{r echo = T}

log(median(x)) == median(xt)

```

### 2.2

Compute the median $\bar{x}$ and median of the eight numbers: 1, 2, 4, 5, 6, 8, 11, 15.

Let $f(x) = \sqrt{x}$

Apply the transformation, then compute the mean, $\tilde{x}$ and median, m, of the transformed data.

```{r echo = T}

x <- c(1, 2, 4, 5, 6, 8, 11, 15)
xt <- sqrt(x)

```

+ Is $f(\bar{x}) = \tilde{x}$?

```{r echo = T}

sqrt(mean(x)) == mean(xt)

```

+ Is $f(m) = \tilde{m}$?

```{r echo = T}

sqrt(median(x)) == median(xt)

```

### 2.4

Import the flights data.

```{r echo = T}

flights <- data.table(read.csv(paste0(data.dir, "FlightDelays.csv"),
                 header = T))

```

a.) Create a table and bar chart of the departure times _(DepartTimes)_

```{r echo = T}

table(flights$DepartTime)

```

```{r echo = T, fig.height=3.5}

ggplot(flights, aes(DepartTime)) +
   geom_bar(aes(fill = ..count..)) +
   labs(title = "Depart Times")

```

\newpage

b.) Create a contingency table of the variables _Day_ and _Delay30_.

```{r echo = T}

delay <- table(flights$Day, flights$Delayed30)

pretty_kable(delay, "Flight Delays")

```

Show the proportions of delayed flights, by day:

```{r echo = T}

pretty_kable(round(prop.table(delay), 4) * 100, "Flight Delays Proportions")

```

\newpage

c.) Create side-by-side boxplots of the lengths of flight times, grouped by wether or not the flight was delayed at least 30 minutes:

```{r echo = T, fig.height=3.5}

ggplot(flights) +
   geom_boxplot(aes(Delayed30, FlightLength, fill = Delayed30)) +
   labs(title = "Flight Length by Delayed30")

```

d.) Do you think there is a relationship between the length of the flight and whether or not the departure time is delayed by at least 30 minutes?

The average flight time is the same, however, the flights that are delayed 30 minutes or more seem to be shorter overall.

### 2.5

Import the General Social Survey data.

```{r echo = T}

gss <- data.table(read.csv(paste0(data.dir, "GSS2002.csv"),
                 header = T))

```

a.) Create a table and a bar chart of the response to the question about the death penalty.

```{r echo = T}

table(gss$DeathPenalty)

```

```{r echo = T, fig.height=3.5}

ggplot(gss, aes(DeathPenalty)) +
   geom_bar(aes(fill = ..count..)) +
   labs(title = "Death Penalty Survey Results")

```

b.) Use the _table_ command and the summary command in R on the gun ownership variable. What additional information does the summary command give that the table does not?

```{r echo = T}

table(gss$OwnGun)

```

```{r echo = T}

summary(gss$OwnGun)

```

The summary tells us how many people didn't respond at all to the question.

c.) Create a contingency table displaying the relationship between opinions about the death penalty to that about gun ownership.

```{r echo = T}

with(gss, {
   table(OwnGun, DeathPenalty)
})

```

d.) What proportion of gun owners favor the death penalty? Does it appear to be different from the proportion among those who do not own guns?

```{r echo = T}

round(prop.table(with(gss, {
   table(OwnGun, DeathPenalty)
})) * 100, 2)

```

It does seem that gun owners are overwhelmingly in favor of the death penalty compared to those who do not own a gun.

### 2.6

Import the data from the recidivism case study in section 1.4.

Import the General Social Survey data.

```{r echo = T}

recid <- data.table(read.csv(paste0(data.dir, "Recidivism.csv"),
                 header = T))

```

a.) Create a table and bar chart  of the __Recid__ variable.

```{r echo = T}
table(recid$Recid)
```

```{r echo = T}
ggplot(recid, aes(Recid)) +
   geom_bar(aes(fill = ..count..)) +
   labs(title = "Recidivism")
```

b.) Create a contingency table summarizing the relationship between recidivism (__Recid__) by age (__Age25__).

```{r echo = T}
with(recid, {
   table(Recid, Age25)
})
```

Of those over 25 years of age, what proportion were sent back to prision?

```{r echo = T}
round(prop.table(with(recid, {
   table(Recid, Age25)
})) * 100, 2)
```

c.) Create side-by-side boxplots of the number of days to recidivism grouped by type of violation, and give three comparative statements about the distributions.

```{r echo = T, fig.height=3.5}
ggplot(recid[!is.na(Days)], aes(Type, Days)) +
   geom_boxplot(aes(fill = Type)) +
   labs(title = "Days Until Recidivism by Crime Type")

```

+ 1.) Technical violations seem to happen quicker after initial release than new crimes.

+ 2.) There are more outliers present in the technical violations, with a cluster of them occuring after 900 days.

+ 3.) The variance is days until recidivism is larger for new crimes. The chances of a technical violation occuring after 450 days drop dramatically.

```{r echo = T}
recid[!is.na(Days), .(Variance = comma(var(Days))), by = Type]
```

d.) Use the quantile command to obtain the quartiles of the number of days to recidivism. Since there are missing values (__NA__) for those released offeneders who had not recidivated, you will need to add the argument __na.rm = T__ to the __quartile__ command to exclude those observations.

```{r echo = T}
quantile(recid$Days, na.rm = T)
```

e.) Create ecdf's of days to recidivism for those under 25 years of age and those 25 years of age or older.

```{r echo = T}

with(recid, {
   recid$U25 <- Age25 == "Under 25"

   plot.ecdf(recid[U25 == T]$Days)
   plot.ecdf(recid[U25 == F]$Days, col="cornflowerblue", add = T)
})

```

Approximately what proportion in each age group were sent back to prision 400 days after release?

Of those who were sent back to prision,

```{r echo = T}

recid[!is.na(Days), 
      .(Days, O400 = Days >= 400), by = Age25][,
                                               .(Proportion = sum(O400) / .N), by = Age25]

```

went back after being release 400 days or more.

### 2.7

Import the data from the black spruce case study in 1.10.

```{r echo=T}
spruce <- data.table(read.csv(paste0(data.dir, "Spruce.csv"),
                 header = T))
```

a.) Compute the numeric summaries for the hight changes of the seedlings.

```{r echo=T}
summary(spruce$Ht.change)
```

b.) Create a histogram and normal quantile plot for the height changes of the seedlings.

Is the distribution approximately normal?

```{r echo=T, fig.height=3}
ggplot(spruce) +
   geom_histogram(aes(x = Ht.change, y = ..density.., fill = ..count..), bins = 20) +
   geom_density(aes(x = Ht.change, y = ..density.., col = "darkred"), lwd = 1) +
   guides(col = "none") +
   labs(title = "Height-Change Densities")
```
```{r echo=T, fig.height=3}
ggplot(spruce) +
   geom_qq(aes(sample = Ht.change)) +
   geom_qq_line(aes(sample = Ht.change)) +
   labs(title = "Height Change QQ-Plot")
```

For the height change variable, we see an uneven density plot, qq-plot with outliers in the tails. The data appers to be "approximately" normal.

c.) Create a boxplot to compare the distribution of the change in diameters of the seedlings (__Di.change__), grouped by wheather or not they were in fertilized plots.

```{r echo=T, fig.height = 3}
ggplot(spruce) +
   geom_boxplot(aes( y = Di.change, fill = Fertilizer ))
```

d.) Use the __tapply__ command to find the numeric summaries of the diameter changes for the two levels of fertilization.

```{r echo=T}
tapply(spruce$Di.change, spruce$Fertilizer, summary)
```

e.) Create a scatter plot of the hight change against the diameter changes, and describe the relationship.

```{r echo=T, fig.height=3.5}
ggplot(spruce, aes(Di.change, Ht.change)) +
   geom_point() +
   geom_smooth(method = "lm") +
   labs(title = "Height Change ~ Diameter Change")
```

\newpage

### 2.8

Import the mobile ads data from section 1.12.

```{r echo=T}
mobile <- data.table(read.csv(paste0(data.dir, "MobileAds.csv"),
                 header = T))
```

a.) Create histograms of the variables __m.cpc_pre__ and __m.cpc_post__, and describe their distributions.

```{r echo=T, fig.height=3.5}

fit <- fitdistr(mobile$m.cpc_pre, "lognormal")

ggplot(mobile, aes(m.cpc_pre)) +
   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
   geom_density(aes(y = ..density.., col = "darkred"), lwd = 1) +
   stat_function(fun = dlnorm, size = 1, color = 'gray',
                    args = list(mean = fit$estimate[1], sd = fit$estimate[2])) +
   guides(col = "none") +
   labs(title = "CPC Pre vs lognormal")

```

```{r echo=T, fig.height=3.5}

pct_zero <- sum(mobile$m.cpc_post == 0) / nrow(mobile) # 7% of values zero

# we will replace the zeros with half the min value here to fit a lognormal.

t_cpc_post <- mobile$d.cpc_post
rep_val <- min( t_cpc_post[t_cpc_post > 0] ) / 2
t_cpc_post[t_cpc_post == 0] <- rep_val

fit <- fitdistr(t_cpc_post, "lognormal")

ggplot(mobile, aes(t_cpc_post)) +
   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
   geom_density(aes(y = ..density.., col = "darkred"), lwd = 1) +
   stat_function(fun = dlnorm, size = 1.2, color = "darkgrey",
                    args = list(mean = fit$estimate[1], sd = fit$estimate[2])) +
   guides(col = "none") +
   labs(title = "CPC Post vs lognormal") +
   theme(legend.position = "none")
```

The distribution of cpc pre and post seem to follow an approximately lognormal distribution.

b.) Compute the difference between these two variables, create a histogram, and describe the distribution.

```{r echo = T, fig.height=3.5}

diff <- data.table(Diff = with(mobile, { m.cpc_pre - m.cpc_post}))[, Index := .I ]

suppressWarnings({
   fit <- fitdistr(diff$Diff, "t")
})

ggplot(diff, aes(Diff)) +
   geom_histogram(aes(y = ..density.., fill = ..count..))+
   geom_density(aes(y = ..density..), col = "cornflowerblue", lwd = 1.2) +
   stat_function(fun = dt, size = 1.2, color = "darkred",
                    args = list(df = fit$estimate[3])) +
   stat_function(fun = dstd, size = 1.2, color = "darkgreen",
                    args = list(mean = fit$estimate[1], sd = fit$estimate[2])) +   
   guides(col = "none") +
   labs(title = "CPC Pre-Post vs t-dist") +
   theme(legend.position = "none")

```

We fit the delta data to a t-distribution, estimating the parameters with __fitdist__. The distribution is way too tail heavy for a normal, and the standard t-distribution has tails that are too light, while the normal t-distribution has tails that are too heavy.

We also note that the true tails in the distribution are polynomial, so getting a close fit would be a bit challenging.

d.) Create a normal quantile plot of the difference. Does it appear to be normally distributed?

```{r echo = T, fig.height=3.5}

ggplot(diff, aes(sample = Diff)) +
   geom_qq() +
   geom_qq_line() +
   labs(title = "Normal QQ-Plot of Pre-Post")

```

This data definitely does not fit a normal distribution. As we saw with the density plots, there are numerious problems with fitting the tails of the data. A simple normal estimate is a terrible fit, while even parametric approaches have problems due to the polynomial tails.

### 2.9

Let $x_1 < x_2 < \ldots < x_n$ and $y_1 < y_2 < \ldots < y_n$ be two sets of data with means $\bar{x}, \bar{y}$ and means $m_x, m_y$, respectively.

Let $w_i = x_i + y_i$ for $i = 1, 2, \ldots, n$.

a.) Prove or give a counterexample: $\bar{x} + \bar{y}$ is the mean of $w_1, w_2, ..., w_n$.

\setlength{\leftskip}{2cm}

$\bar{w} = \bar{x} + \bar{y} = \frac{1}{n}\sum_{i=1}^{n}x_n + \frac{1}{n}\sum_{i=1}^{n}y_n$

$\ldots = \frac{1}{n}\sum_{i=1}^{n}[x_n + y_n]$

\setlength{\leftskip}{0pt}

b.) Prove or give a counterexample: $m_x + m_y$ is the median of $w_1, w_2, ..., w_n$.

\setlength{\leftskip}{2cm}

$m_x = \frac{x_{[(\#x+1)\div2]} + x_{[(\#x+1)\div2]}}{2}$

$m_y = \frac{y_{[(\#y+1)\div2]} + y_{[(\#y+1)\div2]}}{2}$

$m_x + m_y = \frac{x_{[(\#x+1)\div2]} + x_{[(\#x+1)\div2]}}{2} + \frac{y_{[(\#y+1)\div2]} + y_{[(\#y+1)\div2]}}{2}$

$\ldots = \frac{1}{2}(x_{[(\#x+1)\div2]} + x_{[(\#x+1)\div2]} + y_{[(\#y+1)\div2]} + y_{[(\#y+1)\div2]})$

\setlength{\leftskip}{0pt}

\newpage

### 2.10

Find the median __m__ and first and third quartiles for the random variable $X$ having:

a.) The exponential distribution with $pdf f(x) = \lambda e^{-\lambda x}$

\setlength{\leftskip}{2cm}

__m__ = $\frac{ln(2)}{\lambda}$ =

$q_p = -\frac{ln(1 - p)}{\lambda}$, where q in .25, .5, .75.

\setlength{\leftskip}{0cm}

b.) The Pareto distribution with parameter $\alpha > 0$ with $pdf f(x) = \frac{\alpha}{X^{\alpha+1}}$

### 2.11

Let the random variable X have a Cauchy distribution with pdf $f(x) = \frac{1}{\pi(1 + x(x - \theta)^2))}$ for $-\infty < x < \infty$.

a.) Show that the mean of X does not exist.

b.) More generally, will $\mathbb{E}[X^k]$ exist? (k = 1, 2, 3, ...).

c.) Show that $\theta$ is the median of the distribution.

### 2.12

Find:

a.) The $30^{th}$ and $60^{th}$ percentiles for $N(10, 17^2)$.

```{r, echo = T}

# 30th percentile
qnorm(.3, mean = 10, sd = 17)

# 60th percentile
qnorm(.6, mean = 10, sd = 17)

```

b.) The 0.10 and 0.90 quantile for $N(25, 32^2)$.

```{r, echo = T}

# .1 quantile
qnorm(.1, mean = 25, sd = 32)

# .9 quantile
qnorm(.9, mean = 25, sd = 32)

```

c.) The point that marks off the upper 25% in $N(25, 32^2)$.

```{r, echo = T}

qnorm(.75, mean = 25, sd = 32, lower.tail = T)

```

### 2.13

The cdf of the exponential distribution is $F(t) = 1 - e^{-\lambda t}$

a.) Find an expression for the 0.05 quantile $q_{0.05}$.

\setlength{\leftskip}{2cm}

$q_t = \frac{ln(1 - t)}{-\lambda}$

$q_{0.05} = \frac{0.513}{\lambda}$

\setlength{\leftskip}{0cm}

b.) Let $\lambda = 4$, and use your answer from (a) to find $q_0.05$, and then check your answer in R using __qexp__.

```{r, echo = T}
qexp(0.05, 4)

-1/4*log(1 - 0.05)

```

### 2.14

Let $X$ be a random variable with cdf $F(x) = \frac{x^2}{a^2}$ for $0 \le x \le a$.

Find an expression for the $\frac{a}{2}$ and $(1 - \frac{a}{2})$ quantiles, where 0 < a < 1.

$\sqrt{\frac{a}{2}}x$, $1 - \sqrt{\frac{a}{2}}x$

### 2.15

Let X be a random variable with cdf $F(x) = 1 - \frac{9}{x^2}$ for $x \ge 3$.

Find an expression for the $q^{th}$ quantile of $X$.

$\frac{3}{\sqrt{1-x}}$

\newpage

### 2.16

Let $X \sim {\sf Binom}(20, 0.3)$ and let __F__ denotes its cdf.

+ Does there exist a __q__ such that $F(q) = 0.05$?

No, the discrete nature of this distribution doesn't have a value that falls in this interval.

### 2.17

In this exercise, we investigate normal quantile plots using R.

a.) Draw a random sample of size $n = 15$ from $N(0, 1)$, and plot both a normal quantile plot and a histogram.

```{r, echo = T, fig.height=3.5, fig.width=8}

plot_norm_sample <- function( n ) {
   dat <- data.table( X = rnorm(n) )[, N := .I]

   p1 <- ggplot(dat, aes(sample = X)) +
      geom_qq() +
      geom_qq_line() +
      labs(title = paste0( "Normal Distribution, N=", n ))

   p2 <- ggplot(dat, aes(X)) +
      geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
      geom_density(col = "darkorange", fill = "darkorange", alpha = .3)

   grid.arrange(p1, p2, nrow = 2)
}

```

```{r, echo = T, fig.height=3.5, fig.width=8}

plot_norm_sample(15)

```

+ Do the points on the quantile plot appear to fall on a straight line? Is the histogram symmetric, unimodal and bell shaped?

No, the lines do not fit the QQ plot and the histogram is skewed all over.

+ Do this several times.

b.) Repeat part (a) for samples of size n = 30, n = 60, and n = 100.

```{r, echo = T, fig.height=3.5, fig.width=8}

plot_norm_sample(30)

```

```{r, echo = T, fig.height=3.5, fig.width=8}

plot_norm_sample(60)

```

```{r, echo = T, fig.height=3.5, fig.width=8}

plot_norm_sample(100)

```

c.) What lesson do you draw about using graphs to assess whether or not a data set follows a normal distribution?

The random samples from a normal distirbution do not follow the expectation until the sample size is relatively larger (n>30).

### 2.18

Plot by hand the empirical cumulative distribution function for the set of values:

4, 7, 8, 9, 9, 13, 18, 18, 18, 21

### 2.19

The ecdf for a data set with $n = 20$ values is given in figure 2.18.

a.) How many values are less than or equal to 7?

3

b.) How many times does the value 8 occur?

4

c.) In a histogram of these values, how many values fall in the bin (20, 25]?

6

### 2.20

The data set __ChiMarathonMen__ has a sample of times ffor men between 20 and 39 years of age who completed the Chicago Marathon in 2015.

Graph the ecdf's of the times for men in the 25-29-age division and men in the 35-39-age division.

```{r, echo = T, fig.height=4, fig.width=8}
marathon <- data.table(read.csv(paste0(data.dir, "ChiMarathonMen.csv"),
                 header = T))

p1 <- ggplot(marathon[Division == "25-29"], aes(FinishMin)) +
   stat_ecdf(geom = "point") +
   geom_vline(xintercept = 160, color = "cornflowerblue") +
   labs(title = "Men 25-29 Finish Times", y = "")

p2 <- ggplot(marathon[Division == "35-39"], aes(FinishMin)) +
   stat_ecdf(geom = "point") +
   geom_vline(xintercept = 160, color = "cornflowerblue") +
   labs(title = "Men 35-39 Finish Times", y = "")

grid.arrange(p1, p2, nrow = 2)

```

+ Approximately what proportion of men in these two divisions finished in 160 min or less?

25-29 is around 70%, and 35-39 is around 40%.