### 2019-12-03 Mathematical Statistics/04_sampling_distributions.rmd
mathematical statistics, ch4 work
mathematical statistics, ch4 work

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Sampling Distributions}
   \rfoot{\color{headergrey}Chapter 4}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(copula, quietly = TRUE, warn.conflicts = FALSE)
library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(tseries, quietly = TRUE, warn.conflicts = FALSE)
library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
library(gtools, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

setwd("D:/Projects/Statistical-Computing/RDS")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### 4.1

Consider the population {1, 2, 5, 6, 10, 12}.

Find (and plot) the sampling distribution of medians for samples of size 3 without replacement.

```{r, echo = T, fig.height=4.5, fig.width=8}

p <- c(1, 2, 5, 6, 10, 12)
c <- combinations(v = p, n = 6, r = 3)
t <- apply(c, 1, median)

ggplot(data.table(value = t), aes(value, fill = ..count..)) +
   geom_histogram(bins = 30) +
   labs(title = "Median Sampling Distribution of p")

```

Compare the median of the population to the mean of the medians.

Median of p = __`r median(p)`__. Mean of Medians of p = __`r mean(t)`__

\newpage

### 4.2

Consider the population {3, 6, 7, 9, 11, 14}.

For samples of size 3 without replacement, find (and plot) the sampling distribution for the minimum.

```{r, echo = T, fig.height=4.5, fig.width=8}

p <- c(3, 6, 7, 9, 11, 14)
c <- combinations(v = p, n = 6, r = 3)
t <- apply(c, 1, min)

ggplot(data.table(value = t), aes(value, fill = ..count..)) +
   geom_histogram(bins = 30) +
   labs(title = "Minimum Sampling Distribution of p")

```

What is the mean of the sampling distribution? __`r mean(t)`__

The statistic is an estimate of some parameter - what is the value of that parameter?

This is an estimation of the minimum, which is: __`r min(p)`__

\newpage

### 4.3

Let _A_ denote the population {1, 3, 4, 5} and _B_ the population {5, 7, 9}.

```{r, echo = T}

A <- c(1, 3, 4, 5)
B <- c(5, 7, 9)

```

Let _X_ be a random value from _A_, and _Y_ and random value from _B_.

a.) Find the sampling distribution of _X + Y_.

```{r, echo = T}

result = numeric(12)
index <- 1
for(j in 1:length(A))
{
   for(k in 1:length(B))
   {
      result[index] <- A[j] + B[k]
      index <- index + 1
   }
}

sort(result)

```

b.) In this example, does the sampling distribution depend on whether you sample with or without replacement?

_No._

Why or why not?

_Because 5 in is both sets._

c.) Compute the mean of the values for each of _A_ and _B_ and the values in the sampling distribution of _X + Y_.

Mean of _A_: __`r mean(A)`__. Mean of _B_: __`r mean(B)`__.

Mean of _A + B_: __`r mean(result)`__

How are the means related?

mean(A) + mean(B) = mean(A + B).

d.) Suppose you draw a random value from _A_ and a random value from _B_.

```{r, echo = T}

prob <- sum(result >= 13) / length(result)

```

What is the probability that the sum is 13 or larger? __`r prob*100`%__

\newpage

### 4.4

Consider the population {3, 5, 6, 6, 8, 11, 13, 15, 19, 20}.

a.) Compute the mean and standard deviation and create a dot plot of its distribution.

```{r, echo = T, fig.height=3.5, fig.width=8}

p <- c(3, 5, 6, 6, 8, 11, 13, 15, 19, 20)

mu <- mean(p)
sigma <- sd(p)

ggplot(data.table(value = p)) +
   geom_dotplot(aes(value, fill = ..count..), binwidth = 1) +
   labs(title = "Population Dot Plot")

```

$\mu = `r mu`$, $\sigma = `r sigma`$

b.) Simulate the sampling distribution of $\bar{X}$ by taking random samples of size 4 and plot your results.

```{r, echo = T, fig.height=3.5, fig.width=8}

N <- 10e2
results <- numeric(N)

for( i in 1:N)
{
   index <- sample(length(p), size = 4, replace = F)
   results[i] <- mean( p[index] )
}

ggplot(data.table(value = results)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Sample Means")

xbar <- mean(results)
se <- sd(results) / sqrt(N)

```

Compute the mean and standard error, and compare to the population mean and standard deviation.

mean: `r xbar`, standard error: `r se`

c.) Use the simulation to find $P(\bar{X} < 11)$.

```{r, echo = T}

prob <- mean(results < 11)

```

$P(\bar{X} < 11) = `r prob*100`$%

\newpage

### 4.5

Consider two populations A = {3, 5, 7, 9, 10, 16}, B = {8, 10, 11, 15, 18, 25, 28}.

```{r, echo = T}
A <- c(3, 5, 7, 9, 10, 16)
B <- c(8, 10, 11, 15, 18, 25, 28)
```

a.) Using R, draw random samples (without replacement) of size 3 from each population, and simulate the sampling distribution of the sum of their maximums.

```{r, echo = T, fig.height=3.5, fig.width=8}
N <- 10e2
results <- numeric(N)

for(i in 1:N)
{
   samp.a <- sample(A, 3, replace = F)
   samp.b <- sample(B, 3, replace = F)

   results[i] <- max(samp.a) + max(samp.b) 
}

ggplot(data.table(value = results)[, index := .I]) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Sampling Distribution: max(A) + max(B)")

```

b.) Use your simulation to estimate the probability that the sum of the maximums is less than 20.

```{r, echo = T}
prob <- mean(results < 20)
```

Probability: `r prob*100`%

c.) Draw random samples of size 3 from each population, and find the maximum of the union of these two sets.

Simulate the sampling distribution of the maximums of this union.

```{r, echo = T, fig.height=3.5, fig.width=8}
results <- numeric(N)

for(i in 1:N)
{
   samp.a <- sample(A, 3, replace = F)
   samp.b <- sample(B, 3, replace = F)

   results[i] <- max(union(samp.a, samp.b))
}

ggplot(data.table(value = results)[, index := .I]) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Sampling Distribution: max(union(A,B))")
```

d.) Use simulation to find the probability that the maximum of the union is less than 20.

```{r, echo = T}
prob <- mean(results < 20)
```

Probability: `r prob*100`%

\newpage

### 4.6

The data set _Recidivism_ contains the poopulation of all Iowa offenders convicted of either a felony or misdemeanor who were released in 2010 (case study in Section 1.4).

```{r, echo = T}
Recidivism <- data.table(read.csv(paste0(data.dir, "Recidivism.csv"),
                               header = T))
```

Of these, 31.6% recidivated and were sent back to prision.

Simulate the sampling distribution of $\hat{p}$, the sample proportion of offeneders who recidivated, for random samples of size 25.

```{r, echo = T}

mean(Recidivism$Recid == "Yes")

N <- 10e2
results <- numeric(N)

for(i in 1:N)
{
   samp <- sample(Recidivism$Recid, 25)
   results[i] <- mean(samp == "Yes")
}

```

a.) Create a histogram and describe the simulated sampling distribution of $\hat{p}$.

```{r, echo = T, fig.height=3.5, fig.width=8}
ggplot(data.table(value = results)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Recidivism Sampling Distribution")
```

Estimate the mean and standard error.

```{r, echo = T}
mu <- mean(results)
se <- sd(results) / sqrt(25)
```

$\mu = `r mu`$, $\sigma = `r se`$

b.) Compare your estimate of the standard error with the theoretical standard error (_Corollary 4.3.2_).

```{r, echo = T}
tse <- mu * ( 1 - mu ) / sqrt(25)
```

Theoretical: `r tse`

c.) Repeat the above using samples of size 250, and compare with the $n = 25$ case.

```{r, echo = T, fig.height=3.5, fig.width=8}
N <- 10e2
results <- numeric(N)

for(i in 1:N)
{
   samp <- sample(Recidivism$Recid, 250)
   results[i] <- mean(samp == "Yes")
}

ggplot(data.table(value = results)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Recidivism Sampling Distribution")

mu <- mean(results)
se <- sd(results) / sqrt(250)

```

$\mu = `r mu`$, $\sigma = `r se`$

### 4.7

The data set _FlightDelays_ contains the population of all flight departures by United Airlines and American Airlines out of LGA during May and June 2009 (case study in Section 1.1).

```{r, echo = T}
Flights <- data.table(read.csv(paste0(data.dir, "FlightDelays.csv"),
                               header = T))
```

a.) Create a histogram of _Delay_ and describe the distribution.

```{r, echo = T, fig.height=3.5, fig.width=8}

ggplot(Flights, aes(Delay)) +
   geom_histogram(aes(fill = ..count..), bins = 30) +
   scale_y_continuous(labels = comma) +
   labs(title = "Flight Delays")

```

Compute the mean and standard deviation.

```{r, echo = T}
mu <- mean(Flights$Delay)
sigma <- sd(Flights$Delay)
```

$\mu = `r mu`$, $\sigma = `r sigma`$

b.) Simulate the sampling distribution of $\bar{x}$, the sample mean of the length of the flight delays (_Delay_), for sample size 25.

```{r, echo = T}

N <- 10e2
results <- numeric(N)

for(i in 1:N)
{
   samp <- sample(Flights$Delay, 25, replace = F)
   results[i] <- mean(samp)
}

```

Create a histogram and describe the simulated sampling distribution of $\bar{x}$.

```{r, echo = T, fig.height=3.5, fig.width=8}
ggplot(data.table(value = results)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Flight Delay Sampling Distribution")
```

Estimate the mean and standard error.

```{r, echo = T}
mu <- mean(results)
se <- sd(results) / sqrt(25)
```

$\mu = `r mu`$, $\Sigma = `r se`$

c.) Compare your estimate of the standard error with the theoretical standard error (_Corollary A.4.1_).

```{r, echo = T}
tse <- var(results) / 25
```

Theoretical: `r tse`

d.) Repeat with sample size 250.

```{r, echo = T, fig.height=3.5, fig.width=8}

N <- 10e2
results <- numeric(N)

for(i in 1:N)
{
   samp <- sample(Flights$Delay, 250, replace = F)
   results[i] <- mean(samp)
}

ggplot(data.table(value = results)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Flight Delay Sampling Distribution")

mu <- mean(results)
se <- sd(results) / sqrt(250)
tse <- var(results) / 250
```

$\mu = `r mu`$, $\Sigma = `r se`$

Theoretical: `r tse`

### 4.8

Let $X_1, X_2, \ldots, X_{25}$ be a random sample from some distribution and $W = T(X_1, X_2, \ldots, X_n)$ be a statistic.

Suppose the _sampling distribution_ of W has a pdf given by $f(x) = \frac{2}{x^2}$, for 1 < x < 2.

Find $P(w < 1.5)$

__Solution__:

```{r, echo = T, fig.height=3.5, fig.width=8}

f <- function(x) 2 / x^2

x <- seq( from = 1.0001, to = 1.999, by = 0.0001)

y <- f(x)

ggplot(data.table(x, y)) +
   geom_point(aes(x, y), color = "cornflowerblue", size = .6) +
   labs(title = "pdf f(x) = 2/x^2")

a <- cumsum(y) / sum(y)
p <- round( a[x == 1.5], 4 ) * 100

d <- data.table(x, y = a)
ggplot(d) +
   geom_point(aes(x, y), color = "cornflowerblue", size = .6) +
   geom_area(aes(x, y), data = d[x < 1.5], fill = "cornflowerblue", alpha = .3) +
   labs(title = paste("cdf f(x) = 2/x^2, A =", p ))

```

Numerical solution: `r p`%

Analytical Solution: ${\int_{1}^{1.5}}\frac{2}{x^2} = \frac{2}{3}$

### 4.9

Let $X_1, X_2, \ldots, X_{n}$ be a random sample from some distribution and $Y = T(X_1, X_2, \ldots, X_n)$ be a statistic.

Suppose the _sampling distribution_ of Y has pdf $f(y) = (3/8)y^2 \ for \ 0 \le y \le 2$.

Find $P(0 \le Y \le \frac{1}{5})$

__Solution__:

```{r, echo = T, fig.height=3.5, fig.width=8}

f <- function(x) (3/8)*x**2

x <- seq( from = 0, to = 2, by = 0.001)

y <- f(x)

ggplot(data.table(x, y)) +
   geom_point(aes(x, y), color = "cornflowerblue", size = .6) +
   labs(title = paste("pdf: ", paste0(deparse(f), collapse = " ")))

a <- cumsum(y) / sum(y)
p <- round( a[x == 1/5], 4 ) * 100

d <- data.table(x, y = a)

ggplot(d) +
   geom_point(aes(x, y), color = "cornflowerblue", size = .6) +
   geom_area(aes(x, y), data = d[x < 1/5], fill = "cornflowerblue", alpha = .3) +
   labs(title = paste("cdf f(x) = 2/x^2, A =", p ))

```

Numerical Solution: `r p`%

Analytical Solution: ${\int_{0}^{\frac{1}{5}}}\frac{x^3}{8} = \frac{.008}{8} = .001 = .1$ %

### 4.10

Suppose the heights of boys in a certain large city follow a distribution with mean 48 in. and variance $9^2$.

Use the CLT approximation to estimate the probability that in a random sample of 30 boys, the mean height is more than 51 in.

```{r, echo = T}

z <- (51 - 48) / (9^2 / sqrt(30))
p <- pnorm(z, lower.tail = F)

```

Probability: __`r round(p, 4)*100`%__

### 4.11

Let $X_1, X_2, \ldots, X_{36} \sim Bern(.55)$ be independent, and let $\hat{p}$ denote the sample proportion.

Use the CLT approximation with continuity correction to find the probability that $\hat{p} \le 0.5$.

```{r, echo = T}

z <- ( .5 - .55 ) / sqrt(.55 * (1 - .55) / 36)

p <- pnorm(z, lower.tail = T)

```

Probability: `r round(p, 4)*100`%

### 4.12

A random sample of size $n = 20$ is drawn from a distribution with mean 6 and variance 10.

Use the CLT approximation to estimate $P(\bar{X} \le 4.6)$.

```{r, echo = T}

z <- ( 4.6 - 6 ) / ( 10 * sqrt(20) )

p <- pnorm(z, lower.tail = T)

```

Probability: `r round(p, 4)*100`%

### 4.13

A random sample of size $n = 244$ is drawn from a distribution with pdf $f(x) = (3/16)(x - 4)^2, 2 \le x \le 6$.

Use the CLT approximation to estimate $P(X \ge 4.2)$.

```{r, echo = T, fig.height=2.6, fig.width=8}
pdf <- function(x) (3/16)*(x - 4)^2
x <- seq(from = 2, to = 6, by = 0.001)
y <- pdf(x)

ggplot(data.table(x,y)) +
   geom_point(aes(x, y), col = "cornflowerblue", lwd = .8) +
   labs(title = paste("PDF: ", paste0(deparse(f), collapse = " ")))

cdf <- function(x) (3/8)*(x - 4)

y <- cumsum(y) / sum(y)

ev <- x[min(which(y > .5))]

ggplot(data.table(x,y)) +
   geom_point(aes(x, y), col = "cornflowerblue", lwd = .8) +
   geom_vline(xintercept = ev) +
   labs(title = paste("CDF: ", paste0(deparse(f), collapse = " ")))

z <- ( 4.2 - ev ) / sqrt(244)
pnorm(z, lower.tail = F)

```

### 4.14

According to the 2000 census, 28.6% of the US adult population recieved a high school diploma.

In a random sample of 800 US adults, what is the probability that between 220 and 230 (inclusive) people have a high school deploma?

Use the CLT approximation with continuity correction, and compare with the exact probability.

__Solution__:

The sampling distribution of $\hat{p}$ is approximately normal with:

```{r, echo = T}

n <- 800
mu <- .286

ev <- 800 * mu
sigma <- sqrt(n*mu*(1-mu))

```

$\mathbb{E}[X] = `r ev`$ and $\sigma = \sqrt{800(.286)(1 - .286)} = `r round(sigma, 4)`$

```{r, echo = T}

l <- pnorm((ev - 219.5) / sigma)
h <- pnorm((ev - 230.5) / sigma)

p <- l - h

```

Probability: `r round(p, 4)`

### 4.15

If $X_1, \ldots, X_n$ are i.i.d. from Unif[0, 1], how large should n be so that $P(\bar{X} - \frac{1}{2} < 0.05) \ge 0.90$,

that is, is there at least a 90% chance that the sample mean is within 0.05 of $\frac{1}{2}$? Use the CLT approximation.

```{r, echo = T}

```

### 4.16

Maria claims that she has drawn a random sample of size 30 from the exponential distribution with $\lambda = 1/10$.

The mean of her sample is 12.

a.) What is the expected value of a sample mean?

$X \sim Exp(\frac{1}{10}), \mathbb{E}(x) = 10$

b.) Run a simulation by drawing 1000 random samples, each of size 30, from Exp(1/10), and compute the mean for each sample.

```{r, echo = T, fig.height=3.5, fig.width=12}

N <- 1000
result <- numeric(N)

for( i in 1:N)
{
   samp <- rexp( n = 30, rate = 1/10)
   result[i] <- mean(samp)
}

ggplot(data.table(result), aes(result)) +
   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
   geom_vline(xintercept = 12, col = "darkorange") +
   stat_density( kernel = "gaussian", fill = "darkorange", alpha = .3) +
   labs(title = "Exp(1/10) Sampling Distribution")

p <- mean(result > 12)

```

What proportion of the sample means is as large or larger than 12? __`r p*100`%__

c.) Is a mean of 12 unusual for a sample of size 30 from Exp(1/10)?

__Yes__, only ~13% of the sample means have a value of 12 or higher.

### 4.17

Let $X \sim N(15, 3^2)$ and $Y \sim N(4, 2^2)$ be independent random variables.

a.) What is the exact sampling distribution of $W = X - 2Y$?

$W \sim N(7, 5^2)$

b.) Use R to simulate the sampling distribution of $W$ and plot your results.

```{r, echo = T, fig.height=3.5, fig.width=8}

X <- rnorm(10e3, 15, 3)
Y <- rnorm(10e3, 4, 2)

W <- X - 2*Y

ggplot(data.table(x = W)) +
   geom_histogram(aes(x, fill = ..count..), bins = 30) +
   geom_vline(xintercept = 7, col = "darkorange", lwd = 1.5) +
   scale_y_continuous(labels = comma) +
   labs(title = "Sampling Distribution")

```

Check that the simulated mean and standard error are close to the theoretical mean and standard error.

```{r, echo = T}
mu <- mean(W)
sigma <- sd(W)
```

$\mu = `r mu`$, $\sigma = `r sigma`$

c.) Use the simulated sampling to estimate $P(W \le 10)$, and then check your estimate with an exact calculation.

```{r, echo = T}
phat <- mean(W <= 10)

p <- pnorm(10, mean = 7, sd = 5)

```

$\hat{p}= `r phat*100`$%

$P(W \le 10) = `r round(p, 4)*100`$%

### 4.18

Let $X \sim Pois(4)$, $Y \sim Pois(12)$, $U \sim Pois(3)$ be independent random variables.

a.) What is the exact sampling distribution of $W = X + Y + U$?

$W \sim Pois(19)$

b.) Use R to simulate the sampling distribution of $W$ and plot your results.

```{r, echo = T, fig.height=3.5, fig.width=8}
W <- rpois(10e3, lambda = 19)

ggplot(data.table(x = W)) +
   geom_histogram(aes(x, fill = ..count..), bins = 30) +
   geom_vline(xintercept = 19, col = "darkorange", lwd = 1.5) +
   scale_y_continuous(labels = comma) +
   labs(title = "Sampling Distribution")

```

Check that the simulated mean and standard error are close to the theoretical mean and standard error.

```{r, echo = T}
mu <- mean(W)
sigma <- sd(W)
```

$\mu = `r mu`$, $\sigma = `r sigma`$

c.) Use the simulated sampling distribution to estimate $P(W \le 14)$ and then check your estimate with an exact calculation.

```{r, echo = T}
phat <- mean(W <= 14)

p <- ppois(14, lambda = 19)

```

$\hat{p} = `r phat*100`$%

$P(W \le 14) = `r round(p, 4)*100`$%

### 4.19

Let $X_1, X_2, \ldots, X_{10} \sim^{i.i.d} N(20, 8^2)$ and $Y_1, Y_2, \ldots, Y_{15} \sim^{i.i.d} N(16, 7^2)$.

Let $W = \bar{X} + \bar{Y}$

a.) Give the exact sampling distribution of W.

$W \sim N(36, 3.1^2)$

b.) Simulate the sampling distribution in R and plot your results.

```{r, echo = T, fig.height=3.5, fig.width=8}

N <- 10e3
result <- numeric(N)

for( i in 1:N)
{
   X <- rnorm(10, 20, 8)
   Y <- rnorm(15, 16, 7)

   result[i] <- mean(X) + mean(Y)
}

ggplot(data.table(x = result)) +
   geom_histogram(aes(x, fill = ..count..), bins = 30) +
   geom_vline(xintercept = 36, col = "darkorange", lwd = 1.5) +
   scale_y_continuous(labels = comma) +
   labs(title = "Sampling Distribution")

```

Check that the simulated mean and standard error are close to the exact mean and standard error.

```{r, echo = T}
mu <- mean(result)
sigma <- sd(result)
```

$\mu = `r mu`$, $\sigma = `r sigma`$

c.) Use your simulation to find $P(W < 40)$. Calculate an exact answer and compare.

```{r, echo = T}
phat <- mean(result <= 40)

p <- pnorm(40, 36, 3)
```

$\hat{p} = `r phat*100`$%

$P(W < 40) = `r round(p, 4)*100`$%