### 2019-12-08 Mathematical Statistics/05_conf_int_bootstrap.rmd
statistical methods, ch2
statistical methods, ch2

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Confidence Intervals - The Bootstrap}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 5}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(gtools, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

setwd("D:/Projects/Statistical-Computing/RDS")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### 5.1

Consider the samples 1-6. Use a six-sided die to obtain three different bootstrap samples and their corresponding means.

```{r,echo=T}
pop <- seq(from = 1, to = 6, by = 1)

n <- 6

s1 <- mean( sample(pop, n, replace = T) )
s2 <- mean( sample(pop, n, replace = T) )
s3 <- mean( sample(pop, n, replace = T) )

```

$\bar{x}^*_1 = `r s1`$, $\bar{x}^*_2 = `r s2`$, $\bar{x}^*_3 = `r s3`$

#### 5.2

Consider the samples 1, 3, 4, and 6 from some distribution.

```{r, echo = T}

pop <- c(1, 3, 4, 6)

samples <- permutations(n = 4, r = 4, pop, repeats.allowed = T)

```

a.) For one random bootstrap sample, find the probability that the mean is one.

```{r, echo = T}

means <- apply(samples, 1, mean)

p <- mean( means == 1 )

```

Probability: __`r round(p, 4) *100`%__

b.) For one random bootstrap sample, find the probability that the maximum is 6.

```{r, echo = T}
maxes <- apply(samples, 1, max)

p <- mean( maxes == 6 )
```

Probability: __`r round(p, 4) *100`%__

c.) For one random bootstrap sample, find the probability that exactly two elements in the sample are less than 2.

```{r, echo = T}

lt2 <- apply(t(apply(samples, 1, function(x) { x < 2})), 1, sum)

p <- mean( lt2 == 2 )

```

Probability: __`r round(p, 4) *100`%__

### 5.3

Consider the sample 1-3.

a.) List all the (ordered) bootstrap samples from this sample. How many are there?

```{r, echo = T}
samples <- permutations(n = 3, r = 3, 1:3, repeats.allowed = T)

n <- nrow(samples)
```

Samples: = $3^3$ = __`r n`__

b.) How many unordered bootstrap samples are there? For example, {1, 2, 2} and {2, 1, 2} are considered to be the same.

```{r, echo = T}
samples <- combinations(n = 3, r = 3, 1:3, repeats.allowed = T)

n <- nrow(samples)

choose(3^3, 3)

```

Samples: = $3^3$ = __`r n`__

c.) How many ordered bootstrap samples have one occurrence of 1 and two occurences of 3?

```{r, echo = T}

samples <- permutations(n = 3, r = 3, 1:3, repeats.allowed = T)

n.ones <- apply(t(apply(samples, 1, FUN = function(x) { x == 1 })), 1, function(x) sum(x) )
n.threes <- apply(t(apply(samples, 1, FUN = function(x) { x == 3 })), 1, function(x) sum(x) )

sum((n.ones == 1 & n.threes == 2) == T)
```

Is this the same number of bootstrap samples that have each of 1, 2 and 3 occuring exactly once?

```{r, echo = T}

n.ones <- apply(t(apply(samples, 1, FUN = function(x) { x == 1 })), 1, function(x) sum(x) )
n.twos <- apply(t(apply(samples, 1, FUN = function(x) { x == 2 })), 1, function(x) sum(x) )
n.threes <- apply(t(apply(samples, 1, FUN = function(x) { x == 3 })), 1, function(x) sum(x) )

sum((n.ones == 1 & n.twos == 1 & n.threes == 1) == T)
```

No, 3 != 6.

d.) Is the probability of obtaining a bootstrap sample with one 1 and two 3's the same as the probabiliity of obtaining a bootstrap sample with each of 1, 2 and 3 occuring exactly once?

```{r, echo = T}

( sum((n.ones == 1 & n.threes == 2)) / n ) == ( sum((n.ones == 1 & n.twos == 1 &  n.threes == 1)) / n )

```

No, 3% and 6% chances respectfully.