### 2019-12-10 Statistical Methods/02_numerical-summaries.rmd
statistical methods, ch2
statistical methods, ch2

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Numerical Summaries}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 2}
   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

setwd("D:/Projects/Statistical-Computing/RDS")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 2

### 2.1

Suppose,

$X_1 = 1, X_2 = 3, X_3 = 0,$
$X_4 = -2, X_5 = 4, X_6 = -1,$
$X_7 = 5, X_8 = 2, X_9 = 10.$

```{r, echo = T}
x <- c(1, 3, 0, -2, 4, -1, 5, 2, 10)
```

Find:

a.) $\sum{X_i}$

```{r, echo = T}
sum(x)
```

b.) $\sum^5_{i=3}X_i$

```{r, echo = T}
sum( x[3:5] )
```

c.) $\sum^4_{i=1}X^3_i$

```{r, echo = T}
sum(x[1:4]^3)
```

d.) $(\sum{X_i})^2$

```{r, echo = T}
(sum(x))^2
```

e.) $\sum{3}$

```{r, echo = T}
3 * length(x)
```

f.) $\sum{(X_i - 7)}$

```{r, echo = T}
sum(x - 7)
```

g.) $3\sum^5_{i=1}X_i - \sum^9_{i=6}X_i$

```{r, echo = T}
3 * sum( x[1:5] ) - sum( x[6:9] )
```

h.) $\sum{10X}$

```{r, echo = T}
sum( 10 * x)
```

i.) $\sum^6_{i=2}iX_i$

```{r, echo = T}
i <- 2:6
sum( i * x[i]  )
```

j.) $\sum6$

```{r, echo = T}
6 * length(x)
```

### 2.2

Express the following in summation notation.

a.) $X_1 + \frac{X_2}{2} + \frac{X_3}{3} + \frac{X_4}{4}$

$\ldots = \frac{X_1}{1} + \frac{X_2}{2} + \frac{X_3}{3} + \frac{X_4}{4}$

$\ldots = \sum{^4}_{i=1}{\frac{X_i}{i}}$

b.) $U_1 + U^2_2 + U^3_3 + U^4_4$

$\ldots = U^1_1 + U^2_2 + U^3_3 + U^4_4$

$\ldots = \sum^4_{i=1}{U^i_i}$

c.) $(Y_1 + Y_2 + Y_3)^4$

$(\sum^3_{i=1}{Y_i})^4$

### 2.3

Show by numerical example that $\sum{X^2_i}$ is not necessarily equal to $(\sum{X_i}^2)$.

```{r, echo = T}
x <- 1:10

sum(x^2)

(sum(x))^2
```

### 2.4

Find the mean and median of the following sets of numbers.

a.) -1, 0, 3, 0, 2, -5.

```{r, echo = T}
x <- c(-1, 0, 3, 0, 2, -5)

mean(x)
median(x)
```

b.) 2, 2, 3, 10, 100, 1,000

```{r, echo = T}
x <- c(2, 2, 3, 10, 100, 1000)

mean(x)
median(x)
```

### 2.5

The final exam scores for 15 students are: 73, 74, 92, 98, 100, 72, 74, 85, 76, 94, 89, 73, 76, 99.

Compute the mean, 20% trimmed mean, and median using R.

```{r, echo = T}
scores <- c(73, 74, 92, 98, 100, 72, 74, 85, 76, 94, 89, 73, 76, 99)

mean(scores)
mean(scores, trim = .2)
median(scores)
```

### 2.6

The average of 23 numbers is 14.7. What is the sum of these numbers?

```{r, echo = T}
23 * 14.7
```

### 2.7

Consider the 10 values: 3, 6, 8, 12, 23, 26, 37, 42, 49, 63.

The mean is $\bar{X} = 26.9$

```{r, echo = T}
x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 63)
mean(x)
```

a.) What is the value of the mean if the largest value, 63, is increased to 100?

```{r, echo=T}
x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 100)
mean(x)
```

b.) What is the mean if 633 is increased to 1,000?

```{r, echo=T}
x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 1000)
mean(x)
```

c.) What do these results illustrate about the mean?

_The mean is very sensitive to outliers._

### 2.8

Repeat the previous exercise, only compute the median instead.

```{r, echo=T}
x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 100)
median(x)
```

b.) What is the mean if 633 is increased to 1,000?

```{r, echo=T}
x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 1000)
median(x)
```

### 2.9

In general, how many values must be altered to make the sample mean arbitrarily large?

_One._

### 2.10

What is the minimum number of values that must be altered to make the 20% trimmed mean and sample median arbitrarily large?

mean = $g = (.2N), g + 1$

median = ~$.5N$

### 2.11

For the values 0, 23, -1, 12, -10, -7, 1, -19, -6, 12, 1, -3 compute the lower and upper quartiles using the ideal fourths.

```{r, echo = T}
# Used in 11 & 12
idealf <- function( x ) {
   n <- length(x)
   sorted <- sort(x)
   i <- (n / 4) + 5/12
   j <- floor(i)
   h <- i - j
   q1 <- (1 - h)*sorted[j] + h*sorted[j + 1]

   k <- n - j + 1
   q2 <- (1 - h)*sorted[k] + h*sorted[k - 1]

   c(x[1], q1, q2, sorted[n])
}

```

```{r, echo = T}
x <- c(0, 23, -1, 12, -10, -7, 1, -19, -6, 12, 1, -3)

idealf(x)
```

### 2.12

For the values: -1, -10, 2, 2, -7, -2, 3, 3, -6, 12, -1, -12, -6, 8, 6 compute the lower and upper quartiles (the ideal fourths).

```{r, echo = T}
x <- c(-1, -10, 2, 2, -7, -2, 3, 3, -6, 12, -1, -12, -6, 8, 6)

idealf(x)
```

### 2.13

Approximately how many values must be altered to make _q2_, the estimate of the upper quartile based on the ideal fourths, arbitrarily large?

_About $\frac{1}{4}$_

### 2.14

Argue that the smallest observed value, $X_1$, satisfies the definition of a measure of location.

_The smallest (or largest) value in a set defines the boundaries of the set. This is by definition a measure of location._

### 2.15

The height of 10 plants is measured in inches and found to be 12, 6, 15, 3, 12, 6, 21, 15, 18 and 12.

```{r, echo = T}

x <- c(12, 6, 15, 3, 12, 6, 21, 15, 18, 12)

xbar <- mean(x)

stopifnot( sum( x - xbar ) == 0 )

```

Verify that $\sum(X_i - \bar{X}) = 0$

### 2.16

For the data in the previous exercise, compute the range, variance and standard deviation.

```{r, echo = T}

n <- length(x)

range <- max(x) - min(x)
variance <- (1/(n-1))*sum( ( x - xbar )^2)
stdDev <- sqrt(variance)

```

$\bar{X} = `r xbar`$, Range = `r range`, $Var = `r variance`$, $\sigma = `r stdDev`$

\newpage

### 2.17

Use the rules of summation notation to show that it is always the case that $\sum(X_i - \bar{X}) = 0$.

$\bar{X} = \frac{1}{n}\sum{X_i}$

$\sum{X_i} - \frac{1}{n}\sum{X_i}$

$\ldots = \frac{1}{n}\sum{X_i} - \sum{X_i}$

$\ldots = \frac{1}{n}(\sum{X_i - X_i})$

$\ldots = \frac{1}{n}(0)$

$\ldots = 0$

### 2.18

Seven different thermometers were used to measure the temperature of a substance. The reading in degrees Celsius are -4.10, -4.13, -5.09, -4.08, -4.10, -4.09 and -4.12.

```{r, echo = T}
x <- c(-4.10, -4.13, -5.09, -4.08, -4.10, -4.09, -4.12)
n <- length(x)
xbar <- 1/n*sum(x)
variance <- (1/(n-1))*sum( (x - xbar)^2 )
stdDev <- sqrt(variance)
```

Find the variance and standard deviation.

$Var = `r variance`$, $\sigma = `r stdDev`$

### 2.19

A weightlifter's maximum bench press (in pounds) in each of 6 successive weaks was 280, 295, 275, 305, 300, 290.

Find the standard deviation.

```{r, echo = T}
x <- c(280, 295, 275, 305, 300, 290)
n <- length(x)
xbar <- 1/n*(sum(x))
variance <- (1/(n-1))*sum( (x - xbar)^2)
sqrt(variance)
```

### 2.20

For the values,

20, 121, 132, 123, 145, 151, 119, 133, 134, 130, 200

use the classic outlier detection rule to determine whether any outliers exist.

```{r, echo = T}
x <- c(20, 121, 132, 123, 145, 151, 119, 133, 134, 130, 200)

x[ abs(x - mean(x)) / sd(x) > 2 ]

```

### 2.21

Apply the boxplot rule and the MAD-median rule using the values in the preceding exercise. Note that the results differ, compared to using the classic rule.

```{r, echo = T}

x[ abs( x - median(x) ) / mad(x) > 2.27 ]

```

Explain why this happened.

_The classic method masks the value 200 as an outlier._

### 2.22

Consider the values,

0, 121, 132, 123, 145, 151, 119, 133, 134, 130, 250.

```{r, echo = T}
x <- c(0, 121, 132, 123, 145, 151, 119, 133, 134, 130, 250)

x[ ( abs(x - mean(x)) /sd(x) ) > 2 ]

```

Are the values 0 and 250 declared outliers using the classic outlier detection rule?

_Yes._

### 2.23

Verify that for the data in the previous exercise, the boxplot rule declares the values 0 and 250 outliers.

```{r, echo = T}

bounds <- quantile(x)[c(2,4)]

lower <- bounds[1] - 1.5*(bounds[2] - bounds[1])
upper <- bounds[2] + 1.5*(bounds[2] - bounds[1])

x[ x < lower | x > upper]

```

### 2.24

Consider the values

20, 121, 132, 123, 145, 151, 119, 133, 134, 240, 250

Verify that no outliers are found using the classic outlier detection rule.

```{r, echo = T}
x <- c(20, 121, 132, 123, 145, 151, 119, 133, 134, 240, 250)

stopifnot( length( x[ ( abs( x - mean(x) ) / sd(x) ) > 2] ) == 0 )

```

### 2.25

Verify that for the data in the previous exercise, the boxplot rule declares the values 20, 240 and 250 outliers.

```{r, echo = T}
x <- c(20, 121, 132, 123, 145, 151, 119, 133, 134, 240, 250)

bounds <- quantile(x)[c(2,4)]

lower <- bounds[1] - 1.5*(bounds[2] - bounds[1])
upper <- bounds[2] + 1.5*(bounds[2] - bounds[1])

x[ x < lower | x > upper]

```

### 2.26

What do the last three exercises suggest about the boxplot rule versus the classic rule for detecting outliers?

_The classic boxplot rule masks outliers._

### 2.27

What is the typical pulse rate (beats per minute) among adults? Imagine that you sample 21 adults, measure their pulse rate, and get:

80, 85, 81, 75, 77, 79, 74, 86, 79, 55, 82, 89, 73, 79, 83, 82, 88, 79, 77, 81, 82

```{r, echo = T}
x <- c(80, 85, 81, 75, 77, 79, 74, 86, 79, 55, 82, 89, 73, 79, 83, 82, 88, 79, 77, 81, 82)
```

compute the 20% trimmed mean.

```{r, echo = T}

n <- length(x)
o <- sort(x)
g <- floor(.2*n)
t <- o[ (g+1):(n-g) ]
tbar <- 1/(n - 2*g)*sum(t)

stopifnot(tbar == mean(x, trim = .2))
```

### 2.28

For the observations,

21, 36, 42, 24, 25, 36, 35, 49, 32

```{r, echo = T}
x <- c(21, 36, 42, 24, 25, 36, 35, 49, 32)
n <- length(x)

xbar <- 1/n * sum(x)

o <- sort(x)
g <- floor(.2*n)
t <- o[ (g+1):(n-g) ]
tbar <- 1/(n - 2*g)*sum(t)

M <- o[ (n + 1)/2 ]

```

Verify that the sample mean, 20% trimmed mean, and median are $\bar{X} = 33.33, \bar{X}_t = 32.9, and \ M = 35$

$\bar{X} = `r xbar`$, $\bar{X}_t = `r tbar`$, $M = `r M`$