### 2019-12-12 Statistical Methods/04_probability.rmd
statistical methods, ch 4
statistical methods, ch 4

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Probability and Related Concepts}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 4}
   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

setwd("D:/Projects/Statistical-Computing/RDS")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 4

### 4.1

If the possible values for x are 0, 1, 2, 3, 4, 5, and the corresponding values for $P(x)$ are 0.2, 0.2, 0.15, 0.3, 0.35, 0.2, 0.1, respectively, does $P(x)$ qualify as a probability function?

```{r, echo = T}
x <- c(0, 1, 2, 3, 4, 5)
p <- c(0.2, 0.2, 0.15, 0.3, 0.35, 0.2, 0.1)

sum(p)
```

_No. P > 1._

### 4.2

If the possible values for x are 2, 3, 4, and the corresponding values for $P(x)$ are 0.2, -0.1, 0.9, respectively, does $P(x)$ qualify as a probability function?

_No. $0 \le x \le 1$_

### 4.3

If the possible values for x are 1, 2, 3, 4, and the corresponding values for $P(x)$ are 0.1, 0.15, 0.5, 0.25, respectively, does $P(x)$ qualify as a probability function?

```{r, echo = T}
x <- c(1, 2, 3, 4)
p <- c(0.1, 0.15, 0.5, 0.25)

sum(p)

```

Yes, this is a valid probability function.

### 4.4

If the possible values for x are 2, 3, 4, 5, and the corresponding values for $P(x)$ are 0.2, 0.3, 0.4, 0.1, respectively, what is the probability of observing a value less than or equal to 3.4?

```{r, echo = T}
x <- c(2, 3, 4, 5)
p <- c(0.2, 0.3, 0.4, 0.1)

stopifnot( sum(p) == 1 )

prob <- sum( p[ x <= 3.4 ] )

```

Probability: __`r prob*100`%__

### 4.5

For the previous distribution, what is the probability of observing a 1?

_Zero._

### 4.6

For the previous distribution, what is the probability of observing a value greater than 3?

```{r, echo = T}
prob <- sum( p[ x > 3 ])
```

Probability: __`r prob*100`%__

### 4.7

For the previous distribution, what is the probability of observing a value greater than or equal to 3?

```{r, echo = T}
prob <- sum( p[ x >= 3 ])
```

Probability: __`r prob*100`%__

### 4.8

If the probability of observing a value less than or equal to 6 is 0.3, what is the probability of observing a value greater than 6?

```{r, echo = T}
prob <- 1 - .3
```

Probability: __`r prob*100`%__

### 4.9

For the probability function:

$x \ : \ 0, 1$

$P(x) \ : 0.7, 0.3$

Verify that the mean and variance are 0.3 and 0.21, respectively.

```{r, echo = T}
x <- c(0, 1)
p <- c(0.7, 0.3)

mu <- sum( x * p )
variance <- sum( (x - mu)^2 * p)

```

$\mu = `r mu`$, $\sigma^2 = `r variance`$

What is the probability of getting a value less than the mean?

_50%_

### 4.10

Imagine that an auto manufacturer wants to evaluate how potential customers will rate handling for a new car being considered for production. Also, suppose that if all potential customers were to rate handling on a four-point scale, 1 being poor and 4 being excellent, the corresponding probabilities associated with these ratings would be:

$P(1) = 0.2, P(2) = 0.4, P(3) = 0.3, P(4) = 0.1$

Determine the population mean, variance and standard deviation.

```{r, echo = T}
x <- 1:4
p <- c(0.2, 0.4, 0.3, 0.1)

stopifnot(sum(p) == 1)

mu <- sum(x * p)
variance <- sum( (x - mu)^2 * p)
stdDev <- sqrt(variance)

```

$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$

### 4.11

If the possible values for x are 1, 2, 3, 4, 5, with probabilities 0.2, 0.1, 0.1, 0.5, 0.1, respectfully, what are the population mean, variance, and standard deviation?

```{r, echo = T}
x <- 1:5
p <- c(0.2, 0.1, 0.1, 0.5, 0.1)

mu <- sum( x * p )
variance <- sum( (x - mu)^2 * p )
sigma <- sqrt(variance)
```

$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$

### 4.12

In the previous exercise, determine the probability of getting a value within one standard deviation of the mean.

That is, $\mu - \sigma \le x \le \mu + \sigma$

```{r, echo = T}

vals <- mu + c(-1, 1)*sigma
round(vals, 4)

sum( p[ x >= vals[1] & x <= vals[2] ] )

```

### 4.13

If the possible values for x are 1, 2, 3, with probabilities 0.2, 0.6, and 0.2, respectively, what is the mean and standard deviation?

```{r, echo = T}
x <- 1:3
p <- c(0.2, 0.6, 0.2)

mu <- sum( x * p )
variance <- sum( (x - mu)^2 * p)
stdDev <- sqrt(variance)

```

$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$

### 4.14

In the previous excersize, suppose the possible values for x are now 0, 2, 4 with the same probabilities as before.

Will the standard deviation increase, decrease or stay the same?

_Increase._

```{r, echo = T}
x <- c(0, 2, 4)
p <- c(0.2, 0.6, 0.2)

mu <- sum( x * p )
variance <- sum( (x - mu)^2 * p)
stdDev <- sqrt(variance)

```

$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$

### 4.15

For the probability function:

$x : 1, 2, 3, 4, 5$
$P(x) : 0.15, 0.2, 0.3, 0.2, 0.15$

Determine the mean, the variance, and the probability that a value is less than the mean.

```{r, echo = T}
x <- 1:5
p <- c(0.15, 0.2, 0.3, 0.2, 0.15)

mu <- sum( x * p)
variance <- sum( ( x - mu)^2 * p )
stdDev <- sqrt(variance)

sum( p[x < mu] )
```

$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$

### 4.16

For the probability function:

$x : 1, 2, 3, 4, 5$
$P(x) : 0.1, 0.25, 0.3, 0.25, 0.1$

Would you expect the variance to be larger or smaller than the previous pdf?

_Larger._

### 4.17

For the probability function:

$x : 1, 2, 3, 4, 5$
$P(x) : 0.2, 0.2, 0.2, 0.2, 0.2$

Would you expect the variance to be larger or smaller than the previous pdf?

_Smaller._

### 4.18

For the following probabilities:

\begin{center}
 \begin{tabular}{||c c c c||}
 \hline
 \multicolumn{4}{|c|}{Income} \\ 
 \hline
 Age & High & Medium & Low \\ [0.5ex] 
 \hline\hline
 < 30 & 0.030 & 0.180 & 0.090 \\ 
 \hline
 30-50 & 0.052 & 0.312 & 0.156 \\
 \hline
  Over 50 & 0.018 & 0.108 & 0.054 \\
 \hline
 \end{tabular}
\end{center}

a.) The probability that someone is under 30.

_.03 + 0.18 + 0.09 = .30_

b.) The probability that someone has a high income given that they are under 30.

_.03 / .3 = .01_

c.) The probability of someone having a low income given that they are under 30.

_0.09 / .3 = 0.3_

d.) The probability of a medium income given that they are over 50.

_0.018 + 0.108 + 0.054 = .18_

_.108 / .18 = .6_

### 4.19

For the previous data, are income and age independent?

_Yes._

### 4.20

\begin{center}
 \begin{tabular}{||c c c||}
 \hline
 \multicolumn{3}{|c|}{Attitude} \\ 
 \hline
 Member & 1 & 0 \\ [0.5ex] 
 \hline\hline
 Yes & 757 & 496 \\ 
 \hline
 No & 1,071 & 1,074 \\
 \hline
 \end{tabular}
\end{center}

```{r, echo = T}
d <- matrix(c(757, 496, 1071, 1074), nrow = 2, byrow = T)

prop.table(data.table(d))

```

a.) Probability of boy choosing "yes".

.4

b.) P(yes|1)

.22

c.) P(1|yes)

.41

d.) is yes independent of attitude?

No, the probabilities are disproportionate

### 4.21

Let Y be the cost of a home and let X be a measure of the crime rate. If the variance of the cost of a home changes with X, does this mean that the cost of a home and the crime rate are dependent?

_Yes, this can only happen when the conditional probabilites change when told X._

### 4.22

If the probability of Y < 6 is .4 given that X = 2, and if the probability of Y < 6 is .3 given that X = 4, does this mean that X and Y are dependent?

_Yes._

### 4.23

If the range of possible Y values varies with X, does this mean that X and Y are dependent?

_Absolutely._

### 4.24

For a binomial with n = 10 and p = .4, determine:

a.) $P(0)$

```{r, echo = T}
dbinom(0, size = 10, prob = .4)
```

b.) $P(X \le 3)$

```{r, echo = T}
pbinom(3, size = 10, prob = .4)
```

c.) $P(X < 3)$

```{r, echo = T}
pbinom(2, size = 10, prob = .4)
```

d.) $P(X > 4)$

```{r, echo = T}
1 - pbinom(4, size = 10, prob = .4)
```

e.) $P(2 \le X \le 5)$

```{r, echo = T}
pbinom(5, size = 10, prob = .4) - pbinom(1, size = 10, prob = .4)
```

### 4.25

For a binomial with n = 15 and p = 0.3, determine.

a.) $P(0)$

```{r, echo = T}
dbinom(x = 0, prob = .3, size = 15)
```

b.) $P( X \le 3 )$

```{r, echo = T}
pbinom( q = 3, prob = .3, size = 15)
```

c.) $P(X < 3)$

```{r, echo = T}
pbinom(2, size = 15, prob = .3)
```

d.) $P(X > 4)$

```{r, echo = T}
pbinom(4, size = 15, prob = .3, lower.tail = F)
```

e.) $P(2 \le X \le 5)$

```{r, echo = T}
pbinom(5, size = 15, prob = .3) - pbinom(1, size = 15, prob = .3)
```

### 4.26

For a binomial with n = 15, p = 0.6 determine the probability of exactly 10 successes.

```{r, echo = T}
dbinom(10, size = 15, prob = .6)
```

### 4.27

For a binomial with n = 7 and p = 0.35, what is the probability of exactly 2 successes?

```{r, echo = T}
dbinom(2, size = 7, p = .35)
```

### 4.28

For a binomial with n = 18 and p = 0.6, determine the mean, variance of X, the total number of successes.

```{r, echo = T}
n <- 18
p <- 0.6
q <- 1 - p

mu <- n * p
variance <- mu * q

```

$\mu = `r mu`$, $\sigma^2 = `r variance`$

### 4.29

For a binomial with n = 22 and p = .2, determine the mean and variance of X, the total number of successes.

```{r, echo = T}
n <- 22
p <- .2
q <- 1 - p

mu <- n * p
variance <- mu * q
```

$\mu = `r mu`$, $\sigma^2 = `r variance`$

### 4.30

For a binomial with n = 20 and p = .7, determine the mean and variance of $\hat{p}$, the proportion of observed success.

```{r, echo = T}
n <- 20
p <- .7
q <- 1 - p

mu <- n * p
variance <- mu * q
```

### 4.31

For a binomial with n = 30 and p = 0.3, determine the mean and variance of $\hat{p}$.

```{r, echo = T}
n <- 30
p <- .3
q <- 1 - p

phat <- p / n
variance <- p*q / n
```

$\hat{p} = `r phat`$, $\sigma^2 = `r variance`$

### 4.32

For a binomial with n = 10 and p = 0.8, determine:

```{r, echo = T}
n <- 10
p <- 0.8
q <- 1 - p

variance <- p*q / n
```

a.) the probability that $\hat{p}$ is less than or equal to 0.7.

```{r, echo = T}

```

b.) the probability that $\hat{p}$ is greater than or equal to 0.8.

```{r, echo = T}

```

c.) the probability that $\hat{p}$ is exactly equal to 0.8.

```{r, echo = T}

```

### 4.33

A coin is rigged so that when it is flipped, the probability of a head is 0.7. If the coin is flipped three times, which is the more likely outcome, exactly three heads or two heads and a tail?

```{r, echo = T}

dbinom(3, 3, .7) # 3 heads

dbinom(2, 3, .7) # 2 heads 1 tail

```

_Two heads, 1 tail._

### 4.34

Imagine that the probability of heads when flipping a coin is given by the binomial probability function with p = 0.5.

If you flip the coin nine times and get nine heads, what is the probability of a head on the 10th flip?

```{r, echo = T}
# independent events.
dbinom(1, 1, .5)
```

### 4.35

The Department of Agriculture of the United States reports that 75% of all people who invest in the futures market lose money. Based on the binomial probability function, with n = 5, determine:

a.) the probability that all 5 lose money.

$P(x) = 0$

```{r, echo = T}
dbinom(5, size = 5, prob = .75)
```

b.) the probability that all five make money.

```{r, echo = T}
dbinom(5, size = 5, prob = .25)

```

c.) the probability that at least two lose money.

```{r, echo = T}
pbinom(q = 3, size = 5, prob = .25)
```

### 4.36

If for a binomial distribution p = 0.4 and n = 25, determine:

```{r, echo = T}
n <- 25
p <- .4
q <- 1 - p

```

a.) $P(X < 11)$

```{r, echo = T}
pbinom(10, size = n, prob = p)
```

b.) $P(X \le 11)$

```{r, echo = T}
pbinom(11, size = n, prob = p)
```

c.) $P(X > 9)$

```{r, echo = T}
pbinom(9, size = n, prob = p, lower.tail = F)
```

d.) $P(X \ge 9)$

```{r, echo = T}
pbinom(8, size = n, prob = p, lower.tail = F)
```

### 4.37

In the previous problem, determine the mean of X, the variance of X, the mean of $\hat{p}$, and the variance of $\hat{p}$.

```{r, echo = T}

mu <- n * p
variance <- mu * q

phat <- p
v <- p*q/n
```

$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\hat{p} = `r phat`$, $\sigma^2 = `r v`$

### 4.38

Given that Z has a standard normal distribution, determine:

a.) $P(Z \ge 1.5)$

```{r, echo = T}
pnorm(1.5, lower.tail = F)
```

b.) $P(Z \le -2.5)$

```{r, echo = T}
pnorm(-2.5)
```

c.) $P(Z < -2.5)$

```{r, echo = T}
pnorm(-2.5)
```

d.) $P(-1 \le Z \le 1)$

```{r, echo = T}
# P(Z > -1) - P(Z > 1)
pnorm(-1, lower.tail = F) - pnorm(1, lower.tail = F)

# 1 - 2*tail_area
1 - 2 * pnorm(1, lower.tail = F)
```

### 4.39

If Z has a standard normal distribution, determine:

a.) $P(Z \le 0.5)$

```{r, echo = T}
pnorm(0.5)
```

b.) $P(Z > -1.25)$

```{r, echo = T}
pnorm(-1.25, lower.tail = F)
```

c.) $P(-1.2 < Z < 1.2)$

```{r, echo = T}
1 - 2 *pnorm(1.2, lower.tail = F)
```

d.) $P(-1.8 \le Z \le 1.8)$

```{r, echo = T}
pnorm(1.8, lower.tail = T) - pnorm(-1.8)
```

### 4.40

```{r, echo = T}

```

### 4.41

```{r, echo = T}

```

### 4.42

```{r, echo = T}

```

### 4.43

```{r, echo = T}

```

### 4.44

```{r, echo = T}

```

### 4.45

```{r, echo = T}

```

### 4.46

```{r, echo = T}

```

### 4.47

```{r, echo = T}

```

### 4.48

```{r, echo = T}

```

### 4.49

```{r, echo = T}

```

### 4.50

```{r, echo = T}

```

### 4.51

```{r, echo = T}

```

### 4.52

```{r, echo = T}

```

### 4.53

```{r, echo = T}

```

### 4.54

```{r, echo = T}

```

### 4.55

```{r, echo = T}

```

### 4.56

```{r, echo = T}

```

### 4.57

```{r, echo = T}

```

### 4.58

```{r, echo = T}

```

### 4.59

```{r, echo = T}

```

### 4.60

```{r, echo = T}

```

### 4.61

```{r, echo = T}

```

### 4.62

```{r, echo = T}

```