### 2020-01-17 Statistical Methods/10_Comparing_Independent_Groups.rmd
statistical methods, ch10
statistical methods, ch10

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Comparing N Independent Groups}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 10}
   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: 
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(here, quietly = T, warn.conflicts = F)
library(ggplot2, quietly = T, warn.conflicts = F)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(WRS, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### Chapter 10

#### 10.1

Consider the following data:

Group 1: 3, 5, 2, 4, 8, 4, 3, 9

Group 2: 4, 4, 3, 8, 7, 4, 2, 5

Group 3: 6, 7, 8, 6, 7, 9, 10, 9

```{r}
g1 <- c(3, 5, 2, 4, 8, 4, 3, 9)
g2 <- c(4, 4, 3, 8, 7, 4, 2, 5)
g3 <- c(6, 7, 8, 6, 7, 9, 10, 9)

data <- data.table(g1, g2, g3)

means <- colMeans(data)
sds <- apply(data, 2, sd) ** 2
```

Assume the three groups have a common population variance, $\sigma^2_p$. Estimate $\sigma^2_p$.

```{r}
# mean squares between groups:
sum(sds)/ncol(data)
```

#### 10.2

For the previous data, use R to test the hypothesis of equal means using the ANOVA F test with $\alpha = 0.05$.

```{r}
anova1(data)
```

#### 10.3

For the previous data, verify that Welch's test statistic is $F_w = 7.7$ with degrees of freedom $v_1 = 2$, $v_2 = 13.4$

```{r}
test <- t1way(data, tr = 0)

assertthat::are_equal(round(test$TEST, 2), 7.77)
assertthat::are_equal(test$nu1, 2)
assertthat::are_equal(round(test$nu2, 2), 13.4)

```

#### 10.4
#### 10.5
#### 10.6
#### 10.7
#### 10.8
#### 10.9
#### 10.10
#### 10.11
#### 10.12
#### 10.13
#### 10.15
#### 10.16
#### 10.17
#### 10.18
#### 10.19
#### 10.20
#### 10.21
#### 10.22
#### 10.23
#### 10.24
#### 10.25
#### 10.26
#### 10.27
#### 10.28
#### 10.29
#### 10.30