### 2020-01-17 Statistical Methods/10_Comparing_Independent_Groups.rmd
statistical methods, group comparisons
statistical methods, group comparisons

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Comparing N Independent Groups}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 10}
   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: 
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(here, quietly = T, warn.conflicts = F)
library(ggplot2, quietly = T, warn.conflicts = F)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(WRS, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### Chapter 10

#### 10.1

Consider the following data:

Group 1: 3, 5, 2, 4, 8, 4, 3, 9

Group 2: 4, 4, 3, 8, 7, 4, 2, 5

Group 3: 6, 7, 8, 6, 7, 9, 10, 9

```{r}
g1 <- c(3, 5, 2, 4, 8, 4, 3, 9)
g2 <- c(4, 4, 3, 8, 7, 4, 2, 5)
g3 <- c(6, 7, 8, 6, 7, 9, 10, 9)

data <- data.table(g1, g2, g3)

means <- colMeans(data)
sds <- apply(data, 2, sd) ** 2
```

Assume the three groups have a common population variance, $\sigma^2_p$. Estimate $\sigma^2_p$.

```{r}
# mean squares between groups:
sum(sds)/ncol(data)
```

#### 10.2

For the previous data, use R to test the hypothesis of equal means using the ANOVA F test with $\alpha = 0.05$.

```{r}
anova <- anova1(data)

anova

ifelse(anova$p.value < 0.05, "Reject", "Fail to Reject")

```

#### 10.3

For the previous data, verify that Welch's test statistic is $F_w = 7.7$ with degrees of freedom $v_1 = 2$, $v_2 = 13.4$

```{r}
test <- t1way(data, tr = 0)

assertthat::are_equal(round(test$TEST, 2), 7.77)
assertthat::are_equal(test$nu1, 2)
assertthat::are_equal(round(test$nu2, 2), 13.4)

```

#### 10.4

Using R, test the hypothesis of equal means using the ANOVA F test for the following data:

Group 1: 15, 17, 22

Group 2: 9, 12, 15

Group 3: 17, 20, 23

Group 4: 13, 12, 17

```{r}
g1 <- c(15, 17, 22)
g2 <- c(9, 12, 15)
g3 <- c(17, 20, 23)
g4 <- c(13, 12, 17)

data <- data.table(g1, g2, g3, g4)

anova <- anova1(data)

ggplot(stack(data)) +
   geom_boxplot(aes(ind, values, fill = ind))

anova

ifelse(anova$p.value < 0.05, "Reject Null", "Fail to Reject")
```

#### 10.5

Using the previous data, compute the common assumed variance.

```{r}
mean( apply(data, 2, var) )
```

#### 10.6

Use R to verify the p-value for the previous data based on Welch's test is 0.124.

```{r}
w <- t1way(data)

assertthat::are_equal(round(w$p.value, 3), 0.124)
```

#### 10.7

For the data in E4, explain why you would get the same result using a 20% trimmed mean.

_Sample size would not reduce with 3 values_

#### 10.8

Why would you not recommend the strategy of testing for equal variances, and if not significant, using kthe ANOVA F test rather than the Welch's method?

_Don't know when power is high enough._

#### 10.9

Five independent groups are compared with n = 15 observations for each group, SSBG = 20 and SSWG = 150.

Perform the ANOVA F test with $\alpha = 0.05$

```{r}
n <- 15; j <- 5
SSBG <- 20; SSWG <- 150

MSBG <- SSBG / (j -1); MSWG <- SSWG / (n * j - j)

f.stat <- MSBG / MSWG

v1 <- j - 1; v2 <- j *n - j
crit.val <- qf(1 - 0.05, v1, v2)

ifelse(f.stat > crit.val, "Reject", "Fail to Reject")

vals <- data.table(x = seq(0, 10, by = .001))
vals[, y := df(vals$x, v1, v2)]

ggplot(vals, aes(x, y)) +
   geom_point(size = .7) +
   geom_vline(xintercept = crit.val, col = "darkred") +
   geom_vline(xintercept = f.stat, col = "cornflowerblue")

```

#### 10.10

Use the R function anova1 to verify that for the following data, MSGB = 14.4 and MSGW = 12.59.

G1: 9, 10, 15

G2: 16, 8, 13, 6

G3: 7, 6, 9

```{r}
g1 <- c(9, 10, 15)
g2 <- c(16, 8, 13, 6)
g3 <- c(7, 6, 9)

data <- list(g1, g2, g3)

anova1(data)
```

#### 10.11

Consider five groups (J = 5) with population means 3, 4, 5, 6 and 7, and a common variance $\sigma^2_p = 2$. If the number of observations in each group is 10 (n = 10), indicate what is being estimated by MSBG, and based on the information given, determine its value. That is, if the population means and common variance were known, what is the value being estimated by MSBG?

MSBG estimates $\sigma^2_p + n\sigma^2_\mu$, the variation among the population means.

```{r}
var(c(3, 4, 5, 6, 7))
```

How does this differ from the value estimated by MSBG?

#### 10.12
#### 10.13
#### 10.15
#### 10.16
#### 10.17
#### 10.18
#### 10.19
#### 10.20
#### 10.21
#### 10.22
#### 10.23
#### 10.24
#### 10.25
#### 10.26
#### 10.27
#### 10.28
#### 10.29
#### 10.30