### 2020-02-19 HOML/05_Logistic-Regression.rmd
merged machine learning
merged machine learning

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Logistic Regression}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 5}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init()

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Logistic Regression

### Data Set

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition.h2o <- as.h2o(attrition)
```
### Overview

Linear Regression vs Logistic Regression

```{r, fig.height=3}
p1 <- ISLR::Default %>%
   mutate(prob = ifelse(default == "Yes", 1, 0)) %>%
   ggplot(aes(balance, prob)) +
   geom_point(alpha = .15) +
   geom_smooth(method = "lm") +
   ggtitle("Linear Regression model fit") +
   ylab("Balance") + xlab("Probability of Default")

p2 <- ISLR::Default %>%
   mutate(prob = ifelse(default == "Yes", 1, 0)) %>%
   ggplot(aes(balance, prob)) +
   geom_point(alpha = .15) +
   geom_smooth(method = "glm", method.args = list(family = "binomial")) +
   ggtitle("Logistic regression model fit") +
   xlab("Balance") + ylab("Probability of Default")

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

Create training (70%) and test (30%) tests.

```{r, echo = T}
set.seed(123)

churn_split <- initial_split(attrition, 
                             prop = 0.7, strata = "Attrition")

churn_train <- training(churn_split)
churn_test <- testing(churn_split)
```

### Simple Logistic Regression

Fit two generalized linear models to predict attrition.

```{r}
model1 <- glm(Attrition ~ MonthlyIncome, family = "binomial",
              data = churn_train)

model2 <- glm(Attrition ~ OverTime, family = "binomial",
              data = churn_train)

```

```{r}
churn_train2 <- churn_train %>% mutate(prob = ifelse(Attrition == "Yes", 1, 0))
churn_train2 <- broom::augment(model2, churn_train2) %>% mutate(.fitted = exp(.fitted))

p1 <- ggplot(churn_train2, aes(MonthlyIncome, prob)) +
   geom_point(alpha = 0.15) +
   geom_smooth(method = "glm", method.args = list(family = "binomial")) +
   ggtitle("Predicted probabilities for model1") +
   xlab("Monthly Income") +
   ylab("Probability of Attrition")

p2 <- ggplot(churn_train2, aes(OverTime, .fitted, color = OverTime)) +
   geom_boxplot(show.legend = F) +
   geom_rug(sides = "b", position = "jitter", alpha = 0.2, show.legend = F) +
   ggtitle("Predicted probabilities for model2") +
   xlab("Over Time") +
   scale_y_continuous("Probability of Attrition", limits = c(0, 1))

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

Model Diagnostics

```{r}
tidy(model1)

tidy(model2)
```

```{r}
exp(coef(model1))

exp(coef(model2))
```

```{r}
confint(model1)

confint(model2)
```

### Multiple Logistic Regression

```{r}
model3 <- glm(
   Attrition ~ MonthlyIncome + OverTime,
   family = "binomial",
   data = churn_train
)

tidy(model3)
```

```{r}
churn_train3 <- churn_train %>% mutate(prob = ifelse(Attrition == "Yes", 1, 0))
churn_train3 <- broom::augment(model3, churn_train3) %>% mutate(.fitted = exp(.fitted))

ggplot(churn_train3, aes(MonthlyIncome, prob, color = OverTime)) +
   geom_point(alpha = .15) +
   geom_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
   ggtitle("Predicted probabilites for model3") +
   xlab("Monthly Income") +
   ylab("Probability of Attrition")
```

### Assessing Model Accuracy

```{r}
set.seed(123)

cv_model1 <- train(
   Attrition ~ MonthlyIncome,
   data = churn_train,
   method = "glm",
   family = "binomial",
   trControl = trainControl(method = "cv", number = 10)
)

set.seed(123)

cv_model2 <- train(
   Attrition ~ MonthlyIncome + OverTime,
   data = churn_train,
   method = "glm",
   family = "binomial",
   trControl = trainControl(method = "cv", number = 10)
)

set.seed(123)

cv_model3 <- train(
   Attrition ~ .,
   data = churn_train,
   method = "glm",
   family = "binomial",
   trControl = trainControl(method = "cv", number = 10)
)

# extract out of sample performance measurse

summary(
   resamples(
      list(
         model_1 = cv_model1,
         model_2 = cv_model2,
         model_3 = cv_model3
      )
   )
)$statistics$Accuracy
```

```{r}
# predicted class
pred_class <- predict(cv_model3, churn_train)

# create confusion matrix
confusionMatrix(
   data = relevel(pred_class, ref = "Yes"),
   reference = relevel(churn_train$Attrition, ref = "Yes")
)
```

No-information rate

```{r}
table(churn_train$Attrition) %>% prop.table()
```

Basically, this is saying if we just predicted "No" for every instance we would have 83.8% accuracy.

```{r}
# Compute predicted probabilities
m1_prob <- predict(cv_model1, churn_train, type = "prob")$Yes
m3_prob <- predict(cv_model3, churn_train, type ="prob")$Yes

# Compute AUC metrics for cv_model1 and cv_model3
perf1 <- prediction(m1_prob, churn_train$Attrition) %>%
   performance(measure = "tpr", x.measure = "fpr")

perf2 <- prediction(m3_prob, churn_train$Attrition) %>%
   performance(measure = "tpr", x.measure = "fpr")

# Plot ROC curves for cv_model1 and cv_model3
plot(perf1, col = "black", lty = 2)
plot(perf2, col = "blue", add = T)

legend(0.8, 0.2, legend = c("cv_model1", "cv_model3"),
       col = c("black", "blue"), lty = 2:1, cex = 0.6)
```

```{r}
set.seed(123)
cv_model_pls <- train(
   Attrition ~ .,
   data = churn_train,
   method = "pls",
   family = "binomial",
   trControl = trainControl(method = "cv", number = 10),
   preProcess = c("zv", "center", "scale"),
   tuneLength = 16
)

cv_model_pls$bestTune

ggplot(cv_model_pls)
```

### Feature Interpretation

```{r}
vip::vip(cv_model3, num_features = 20)
```

```{r}
pred.fun <- function(object, newdata) {
   Yes <- mean(predict(object, newdata, type = "prob")$Yes)
   as.data.frame(Yes)
}

p1 <- pdp::partial(cv_model3, pred.var = "OverTime", pred.fun = pred.fun ) %>%
   autoplot(rug = T) + ylim(c(0, 1))

p2 <- pdp::partial(cv_model3, pred.var = "JobSatisfaction", pred.fun = pred.fun) %>%
   autoplot() + ylim(c(0, 1))

p3 <- pdp::partial(cv_model3, pred.var = "NumCompaniesWorked", pred.fun = pred.fun, gr = 10) %>%
   autoplot() + scale_x_continuous(breaks = 0:9) + ylim(c(0, 1))

p4 <- pdp::partial(cv_model3, pred.var = "EnvironmentSatisfaction", pred.fun = pred.fun) %>%
   autoplot() + ylim(c(0, 1))

grid.arrange(p1, p2, p3, p4, nrow = 2)
```

```{r}
# clean up
rm(list = ls())
```