### 2020-02-19 HOML/08_K-Nearest-Neighbors.rmd

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}K-Nearest Neighbors}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 8}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(earth, quietly = TRUE, warn.conflicts = FALSE)
library(mda, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Multivariate Adaptive Regression Splines

### Data Sets

Attrition

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition.h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")
churn.train <- training(churn)
churn.test <- testing(churn)

```

Ames, Iowa housing data.

```{r, echo = T}
set.seed(123)

ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)

ames.split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames.train <- training(ames.split)
ames.test <- testing(ames.split)
```

MNIST

```{r}
mnist <- dslabs::read_mnist()
names(mnist)
```

### Overview

K-nearest neighbors (KNN) is a very simple algorithm in which each observation is predicted based on its "similarity" to other observations.

```{r}
df <- recipe(Sale_Price ~., data = ames.train) %>%
   step_nzv(all_nominal()) %>%
   step_integer(matches("Qual|Cond|QC|Qu")) %>%
   step_center(all_numeric(), -all_outcomes()) %>%
   step_scale(all_numeric(), -all_outcomes()) %>%
   step_dummy(all_nominal(), -all_outcomes(), one_hot = T) %>%
   prep(training = ames.train, retain = T) %>%
   juice() %>%
   select(-Sale_Price)

home <- 30
k <- 10
index <- as.vector(FNN::knnx.index(df[-home, ], df[home, ], k = k))
knn.homes <- ames.train[c(home, index), ]

knn.homes %>%
   select(Longitude, Latitude) %>%
   mutate(desc = factor(c('House of Interest', rep('Closest Neighbors', k)),
                        levels = c("House of interest", "Closest Neighbors"))) %>%
   qmplot(Longitude, Latitude, data = .,
          maptype = "toner-background", darken = .7, color = desc, size = I(2.5)) +
   theme(legend.position = "top",
         legend.title = element_blank())
```

### Distance Measures

Euclidean: $\sqrt{\sum^{P}_{j=1}(x_{aj} - x_{bj})^2}$

Manhattan: $\sum^{P}_{j=1}|x_{aj} - x_{bj}|$

Example:

```{r}
two.homes <- ames.train %>%
   select(Gr_Liv_Area, Year_Built) %>%
   sample_n(2)

two.homes

# Euclidean
dist(two.homes, method = "euclidean")

# Manhattan
dist(two.homes, method = "manhattan")
```

Visually:

```{r}
p1 <- ggplot(two.homes, aes(Gr_Liv_Area, Year_Built)) +
  geom_point() +
  geom_line(lty = "dashed") +
  ggtitle("(A) Euclidean distance")

p2 <- ggplot(two.homes, aes(Gr_Liv_Area, Year_Built)) +
  geom_point() +
  geom_step(lty = "dashed") +
  ggtitle("(B) Manhattan distance")

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

### Pre-processing

```{r}
home1 <- ames %>%
  mutate(id = row_number()) %>%
  select(Bedroom_AbvGr, Year_Built, id) %>%
  filter(Bedroom_AbvGr == 4 & Year_Built == 2008) %>%
  slice(1) %>%
  mutate(home = "home1") %>%
  select(home, everything())

home2 <- ames %>%
  mutate(id = row_number()) %>%
  select(Bedroom_AbvGr, Year_Built, id) %>%
  filter(Bedroom_AbvGr == 2 & Year_Built == 2008) %>%
  slice(1) %>%
  mutate(home = "home2") %>%
  select(home, everything())

home3 <- ames %>%
  mutate(id = row_number()) %>%
  select(Bedroom_AbvGr, Year_Built, id) %>%
  filter(Bedroom_AbvGr == 3 & Year_Built == 1998) %>%
  slice(1) %>%
  mutate(home = "home3") %>%
  select(home, everything())

features <- c("Bedroom_AbvGr", "Year_Built")

# distance between home 1 and 2
dist(rbind(home1[,features], home2[,features]))

# distance between home 1 and 3
dist(rbind(home1[,features], home3[,features]))
```

```{r}
scaled.ames <- recipe(Sale_Price ~., ames.train) %>%
   step_center(all_numeric()) %>%
   step_scale(all_numeric()) %>%
   prep(training = ames, retain = T) %>%
   juice()

home1.std <- scaled.ames %>%
   mutate(id = row_number()) %>%
   filter(id == home1$id) %>%
   select(Bedroom_AbvGr, Year_Built, id) %>%
   mutate(home = "home1") %>%
   select(home, everything())

home2.std <- scaled.ames %>%
   mutate(id = row_number()) %>%
   filter(id == home2$id) %>%
   select(Bedroom_AbvGr, Year_Built, id) %>%
   mutate(home = "home2") %>%
   select(home, everything())

home3.std <- scaled.ames %>%
   mutate(id = row_number()) %>%
   filter(id == home1$id) %>%
   select(Bedroom_AbvGr, Year_Built, id) %>%
   mutate(home = "home3") %>%
   select(home, everything())

dist(rbind(home1.std[, features], home2.std[, features]))

dist(rbind(home1.std[, features], home3.std[, features]))
```

### Choosing K

Blueprint

```{r}
blueprint <- recipe(Attrition ~., data = churn.train) %>%
   step_nzv(all_nominal()) %>%
   step_integer(contains("Satisfaction")) %>%
   step_integer(WorkLifeBalance) %>%
   step_integer(JobInvolvement) %>%
   step_dummy(all_nominal(), -all_outcomes(), one_hot = T) %>%
   step_center(all_numeric(), -all_outcomes()) %>%
   step_scale(all_numeric(), -all_outcomes())

# Create a resampling method
cv <- trainControl(
   method = "repeatedcv",
   number = 10,
   repeats = 5,
   classProbs = T,
   summaryFunction = twoClassSummary
)

# Create a hyperparameter grid search
hyper.grid <- expand.grid(
   k = floor(seq(1, nrow(churn.train)/3, length.out = 20))
)

# Fit knn model and perform grid search
knn.grid <- train(
   blueprint,
   data = churn.train,
   method = "knn",
   trControl = cv,
   tuneGrid = hyper.grid,
   metric = "ROC"
)

ggplot(knn.grid)
```

MNIST

KNN doesn't work well with zero (or near zero) variance features. This is a particularly bad problem with MNIST:

```{r}
set.seed(123)

index <- sample(nrow(mnist$train$images), size = 10e3)

mnist.x <- mnist$train$images[index, ]
mnist.y <- factor(mnist$train$labels[index])

mnist.x %>%
   as.data.frame() %>%
   map_df(sd) %>%
   gather(feature, sd) %>%
   ggplot(aes(sd)) +
   geom_histogram(binwidth = 1)
```

```{r}
nzv <- nearZeroVar(mnist.x)

par(mfrow = c(1, 4))
i <- 2
image(1:28, 1:28, matrix(mnist$test$images[i,], nrow=28)[ , 28:1], 
      col = gray(seq(0, 1, 0.05)), xlab = "", ylab="", 
      xaxt="n", yaxt="n", main = "(A) Example image \nfor digit 2")
i <- 7
image(1:28, 1:28, matrix(mnist$test$images[i,], nrow=28)[ , 28:1], 
      col = gray(seq(0, 1, 0.05)), xlab = "", ylab="", 
      xaxt="n", yaxt="n", main = "(B) Example image \nfor digit 4")

i <- 9
image(1:28, 1:28, matrix(mnist$test$images[i,], nrow=28)[ , 28:1], 
      col = gray(seq(0, 1, 0.05)), xlab = "", ylab="", 
      xaxt="n", yaxt="n", main = "(C) Example image \nfor digit 5")
image(matrix(!(1:784 %in% nzv), 28, 28), col = gray(seq(0, 1, 0.05)), 
      xaxt="n", yaxt="n", main = "(D) Typical variability \nin images.")
```

Clean-up features w/ carat

```{r}
colnames(mnist.x) <- paste0("V", 1:ncol(mnist.x))

# Remove near zero variance

nzv <- nearZeroVar(mnist.x)
index <- setdiff(1:ncol(mnist.x), nzv)
mnist.x <- mnist.x[, index]
```

hyper-parameter search

```{r}
cv <- trainControl(
   method = "LGOCV",
   p = 0.7,
   number = 1,
   savePredictions = T
)

hyper.grid <- expand.grid(k = seq(3, 25, by = 2))

knn.mnist <- train(
   mnist.x,
   mnist.y,
   method = "knn",
   tuneGrid = hyper.grid,
   preProc = c("center", "scale"),
   trControl = cv
)

ggplot(knn.mnist)
```

Diagnostics

```{r}
cm <- confusionMatrix(knn.mnist$pred$pred, knn.mnist$pred$obs)
cm$byClass[, c(1:2, 11)] # sensitivity, specificity & accuracy
```

```{r}
vi <- varImp(knn.mnist)
vi
```

Most influential areas of the image:

```{r}
par(mfrow = c(1,1))
# Get median value for feature importance
imp <- vi$importance %>%
   rownames_to_column(var = "feature") %>%
   gather(response, imp, -feature) %>%
   group_by(feature) %>%
   summarize(imp = median(imp))

# Create tibble for all edge pixels
edges <- tibble(
   feature = paste0("V", nzv),
   imp = 0
)

# Combine and plot
imp <- rbind(imp, edges) %>%
   mutate(ID = as.numeric(str_extract(feature, "\\d+"))) %>%
   arrange(ID)

image(matrix(imp$imp, 28, 28), col = gray(seq(0, 1, 0.05)),
      xaxt = "n", yaxt = "n")
```

```{r}
# Get a few accurate predictions
set.seed(9)
good <- knn_mnist$pred %>%
  filter(pred == obs) %>%
  sample_n(4)

# Get a few inaccurate predictions
set.seed(9)
bad <- knn_mnist$pred %>%
  filter(pred != obs) %>%
  sample_n(4)

combine <- bind_rows(good, bad)

# Get original feature set with all pixel features
set.seed(123)
index <- sample(nrow(mnist$train$images), 10000)
X <- mnist$train$images[index,]

# Plot results
par(mfrow = c(4, 2), mar=c(1, 1, 1, 1))
layout(matrix(seq_len(nrow(combine)), 4, 2, byrow = FALSE))
for(i in seq_len(nrow(combine))) {
  image(matrix(X[combine$rowIndex[i],], 28, 28)[, 28:1], 
        col = gray(seq(0, 1, 0.05)),
        main = paste("Actual:", combine$obs[i], "  ", 
                     "Predicted:", combine$pred[i]),
        xaxt="n", yaxt="n") 
}
```

```{r}
# clean up
rm(list = ls())
```