### 2020-02-19 ISLR/08_Tree-Based-Methods.rmd

---
title: ''
mainfont: Arial
fontsize: 12pt
fig_width: 9
fig_height: 3.5
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 8}
   \fancyfoot[C]{\rmfamily\color{headergrey}Tree Based Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)
library(tree, quietly = TRUE, warn.conflicts = FALSE)
library(randomForest, quietly = TRUE, warn.conflicts = FALSE)
library(gbm, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select

```

## Chapter 8

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Lab

### Fitting Classification Trees

```{r}
carseats <- as.data.table(ISLR::Carseats)

carseats[, High := as.factor(ifelse(Sales <= 8, "No", "Yes"))]

tree.carseats <- tree(formula = High ~ .-Sales, data = carseats)

summary(tree.carseats)

plot(tree.carseats)
text(tree.carseats, pretty = 0)
```

```{r}
set.seed(2)

train <- sample(1:nrow(carseats), 200)

carseats.test <- carseats[-train]
high.test <- carseats[-train]$High

tree.carseats <- tree(High ~.-Sales, data = carseats, subset = train)
tree.pred <- predict(tree.carseats, carseats.test, type = "class")

table(tree.pred, high.test)
```

```{r}
set.seed(3)

cv.carseats <- cv.tree(tree.carseats, FUN = prune.misclass)
names(cv.carseats)

cv.carseats
```

```{r}
par(mfrow = c(1, 2))
plot(cv.carseats$size, cv.carseats$dev, type = "b")
plot(cv.carseats$k, cv.carseats$dev, type="b")
```

```{r}
par(mfrow = c(1,1))
prune.carseats <- prune.misclass(tree.carseats, best = 9)
plot(prune.carseats)
text(prune.carseats, pretty = 0)
```

```{r}
tree.pred <- predict(prune.carseats, carseats.test, type = "class")
table(tree.pred, high.test)
```

```{r}
prune.carseats <- prune.misclass(tree.carseats, best = 15)
plot(prune.carseats)
text(prune.carseats, pretty = 0)

tree.pred <- predict(prune.carseats, carseats.test, type = "class")
table(tree.pred, high.test)
```

### Regression Trees

```{r}

boston <- as.data.table(Boston)

N <- nrow(boston)

set.seed(1)

train <- sample(1:N, N /2)

tree.boston <- tree(medv ~ ., boston, subset = train)

summary(tree.boston)
```

```{r}
plot(tree.boston)
text(tree.boston, pretty = 0)
```

```{r}
cv.boston <- cv.tree(tree.boston)
plot(cv.boston$size, cv.boston$dev, type = 'b')
```

```{r}
prune.boston <- prune.tree(tree.boston, best = 5)
plot(prune.boston)
text(prune.boston, pretty = 0)
```

```{r}
yhat <- predict(tree.boston, newdata = boston[-train,])
boston.test <- boston[-train]$medv

plot(yhat, boston.test)
abline(0, 1)
mean((yhat - boston.test)^2)
```

### Bagging and Boosting

```{r}
set.seed(1)

bag.boston <- randomForest(medv ~ ., data = boston, subset = train, mtry = 13, importance = T)
bag.boston
```

```{r}
yhat.bag <- predict(bag.boston, newdata = boston[-train])
plot(yhat.bag, boston.test)
abline(0, 1)
mean((yhat.bag-boston.test)^2)
```

```{r}
bag.boston <- randomForest(medv ~ ., data = boston, subset = train, mtry = 13, ntree = 25)
yhat.bag <- predict(bag.boston, newdata = boston[-train,])
mean((yhat.bag - boston.test)^2)
```

```{r}
set.seed(1)

rf.boston <- randomForest(medv ~  ., data = boston, subset = train, mtry = 6, importance = T)
yhat.rf <- predict(rf.boston, newdata = boston[-train,])
mean((yhat.rf-boston.test)^2)
```

```{r}
importance(rf.boston)
```

```{r}
varImpPlot(rf.boston)
```

### Boosting

```{r}
boost.boston <- gbm(medv ~ ., data = boston[-train,], distribution = "gaussian", n.trees = 5000, interaction.depth = 4)

summary(boost.boston)
```

```{r}
par(mfrow = c(1, 2))
plot(boost.boston, i = "rm")
plot(boost.boston, i = "lstat")
```

```{r}
yhat.boost <- predict(boost.boston, newdata = boston[-train,], n.trees = 5000)
mean((yhat.boost - boston.test)^2)
```

## Applied

In the lab, we applied random forests to the *Boston* data using mtry = 6 and using ntree = 25 and ntree = 500. Create a plot displaying the test error resulting from random forests on this data set for a more comprehensive range of values for mtry ntree.

```{r}
N <- nrow(boston)

train <- sample(1:N, N * .7)

boston.train <- boston[train]
boston.test <- boston[!train]

ntrees <- seq(0, 500, 10)

results <- numeric(length(ntrees))

for(i in 1:length(ntrees))
{
   tree.count <- ntrees[i]

   rf <- randomForest(medv ~ ., data = boston.train, mtry = 6, n.trees = tree.count)

   pred <- predict(rf, newdata = boston.test)

   results[i] <- sqrt(mean((pred - boston.test$medv)^2)) # store the rmse
}

lowest.error <- which.min(results)

rf.results <- data.table(trees = ntrees, errors = results)[, lowest := .I == lowest.error]

ggplot(rf.results, aes(trees, errors, fill = lowest)) +
   geom_line()

```

In the lab, a classification tree was applied to the Carseats data set after converting Sales into a qualitative response variable. Now we will seek to predict Sales using regression trees and related approaches, treating the response as a quantitaive variable.

a.) Split the data into a t training set and a test set.

```{r}
N <- nrow(carseats)

index <- sample(1:N, N * .7)

carseats.train <- carseats[index]
carseats.test <- carseats[!index]
```

b.) Fit a regression tree to the training set. Plot the tree, and interpret the results. What test MSE do you obtain?

```{r}
carseat.tree <- tree(Sales ~., data = carseats.train)
pred <- predict(carseat.tree, newdata = carseats.test)
mse <- mean((pred - carseats.test$Sales)^2)

mse

par(mfrow = c(1,1))
plot(rf)
text(rf, pretty = 0)
```

c.) Use cross-validation in order to determine the optimal level of tree complexity. Does pruning the tree improve the test MSE?

```{r}
carseat.cv <- cv.tree(carseat.tree)

plot(carseat.cv$size, carseat.cv$dev, type = "b", col = "red")
plot(carseat.cv$size, carseat.cv$k, type = "b", col = "red")

prune.carseats <- prune.tree(carseat.tree, best = 5)

plot(prune.carseats)
text(prune.carseats)
```

d.) Use the bagging approach in order to analyze this data. What test MSE do you obtain? Use the importance() function to determine which variables are most important.

```{r}

```

```{r}

```