### 2020-02-19 Mathematical Statistics/06_Estimation-Notes.Rmd

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Estimation}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 6}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width="50%")
library(ggplot2)
library(dplyr)
library(stats)
```

### Wind speed case study: fitting the Weibull distribution

This function takes input shape parameter k and  the data to compute
 $(1/k)+ (1/n)*\sum (\log(x_i)) +(1/\alpha)\sum x_i^k\log(x_i)=0$  where $\alpha= \sum x_i^k$.

```{r}
weibull.shape <- function(k, data)
{
  numer <- colSums(outer(data, k, "^") * log(data))
  denom <- colSums(outer(data, k, "^"))
  numer/denom - 1/k - mean(log(data))
}
```

 This function takes input shape parameter k
 and data to compute   $k^{th}$ root of $(1/n) \sum x_i^k$
 where $n=$number of data values.

```{r}
weibull.scale <- function(k, data)
{
  mean(data^k)^(1/k)
}
```

 <tt>`uniroot`</tt> is a built-in function which estimates the roots of a function.
 Provide function, any arguments needed for function,
 and a guess of values two values around root.
 Function values must be opposite signs at lower
 and upper guess.

Now, we do the data specific commands:

```{r}
Turbine <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/Turbine.csv")
wind <- Turbine$AveSpeed

#alternatively, wind <- subset(Turbine, select = AveSpeed, drop = TRUE)
```
Estimate the shape parameter k:
```{r}
uniroot(weibull.shape, data = wind, lower = 1,upper = 5)
```
With estimate of shape parameter, now find estimate of scale parameters lambda:

```{r}
weibull.scale(3.169, wind)
```

Next, plot histogram with density curve overlap.
The <tt>`prob = TRUE`</tt> argument scales histogram to area 1.

```{r}
myfun <- function(x){
  dweibull(x, 3.169, 7.661)
}

df <- data.frame(wind)
ggplot(df, aes(wind)) + geom_histogram(aes(y = stat(density)), bins=12) + labs(title="Distribution of average wind speeds", x = "meters/sec") + stat_function(fun = myfun, colour="blue")

ggplot(df, aes(wind)) + stat_ecdf() + stat_function(fun=pweibull, args=list(shape=3.169, scale = 7.661), colour = "blue") + labs(title = "ECDF of wind data")
```

###Goodness-of-fit

Now for the chi-square goodness of fit test
Get the deciles of the Weibull distribution with the given parameters and also, the range of the wind data:

```{r}
q <- qweibull(seq(.1, .9, by = .1), 3.169, 7.661)

range(wind)
```
Now, encompass the range of wind and get the counts in each sub-interval. The
<tt>`plot = FALSE`</tt> argument suppresses the histogram and just gives statistics:
```{r}
q <- c(0, q, 14)

hist(wind, breaks = q, plot = F)$counts

#Now store the output
count <- hist(wind,breaks = q, plot = F)$counts

expected <- length(wind)*.1
```
Compute chi-square test statistic:
```{r}
(count-expected)^2/expected
```

### Example 6.13
Simulation comparing two estimators for uniform:

```{r}
Xbar <- numeric(1000)
maxY <- numeric(1000)

#set.seed(100)
for (i in 1:1000)
 {
   x <- runif(25,0,12) #sample n=25 from Unif[0,12]
   Xbar[i] <- 2*mean(x)
   maxY[i] <- 26/25*max(x)
 }

 #mean and standard deviation of the method of moments estimate
 mean(Xbar)
 sd(Xbar)
range(Xbar)  
 #mean and sd of estimate from MLE
 mean(maxY)
 sd(maxY)
 range(maxY)
```

To plot the results:

```{r}
library(gridExtra)

p1 <- ggplot() + geom_histogram(aes(Xbar), breaks=seq(7, 16.5, by =.5))  + ylim(c(0, 650)) + labs(title = "2*Sample mean", x = "means")
p2 <- ggplot() + geom_histogram(aes(maxY), breaks=seq(7, 16.5, by = .5)) + ylim(c(0, 650)) +labs(title = "25/24*maximum", x = "maximums")
grid.arrange(p1, p2)

```

### Example 6.14

Relative efficiency for the Wind Speed Case Study
```{r}
# Theta = P(wind > 5), nonparametric and parametric estimates
n <- length(wind)
theta1 <- mean(wind > 5)
theta2 <- pweibull(5, 3.169, 7.661, lower.tail = FALSE)
theta1 # 0.75
theta2 # 0.7720827
eta1 <- quantile(wind, .1)
eta2 <- qweibull(.1, 3.169, 7.661)
eta1 # 3.77
eta2 # 3.766038

# nonparametric bootstrap to find standard errors
set.seed(23)
N <- 10^4
boot.theta1 <- numeric(N)
boot.shape <-  numeric(N)
boot.scale <-  numeric(N)
boot.theta2 <- numeric(N)
boot.eta1 <- numeric(N)
boot.eta2 <- numeric(N)
for (i in 1:N)
{
  boot.wind <- wind[sample(1:n, replace=TRUE)]
  boot.theta1[i] <- mean(boot.wind > 5)
  boot.shape[i] <- uniroot(weibull.shape, data=boot.wind, lower=1,upper=5)$root
  boot.scale[i] <- weibull.scale(boot.shape[i], boot.wind)
  boot.theta2[i] <- pweibull(5, boot.shape[i], boot.scale[i], lower.t = FALSE)
  boot.eta1[i] <- quantile(boot.wind, .1)
  boot.eta2[i] <- qweibull(.1, boot.shape[i], boot.scale[i])
}

ggplot() + stat_qq(aes(sample = boot.theta1)) # discrete, normal
ggplot() + stat_qq(aes(sample = boot.theta2)) # continuous, normal
ggplot() + stat_qq(aes(sample = boot.eta1)) # discrete, irregular, roughly normal
ggplot() + stat_qq(aes(sample = boot.eta2)) # continuous, very normal

sd(boot.theta1) # 0.03331767
sd(boot.theta2) # 0.02342771
var(boot.theta1) / var(boot.theta2) # 2.022504

# Formula standard error, for comparison
sqrt(theta1 * (1-theta1) / n) # standard error = 0.03340766

sd(boot.eta1) # 0.2161054
sd(boot.eta2) # 0.1839509
var(boot.eta1) / var(boot.eta2) # 1.380154

# parametric bootstrap to find standard errors
set.seed(23)
pboot.theta1 <- numeric(N)
pboot.shape <-  numeric(N)
pboot.scale <-  numeric(N)
pboot.theta2 <- numeric(N)
pboot.eta1 <- numeric(N)
pboot.eta2 <- numeric(N)
for (i in 1:N)
{
  boot.wind <- rweibull(n, 3.169, 7.661)
  pboot.theta1[i] <- mean(boot.wind > 5)
  pboot.shape[i] <- uniroot(weibull.shape, data=boot.wind, lower=1,upper=5)$root
  pboot.scale[i] <- weibull.scale(pboot.shape[i], boot.wind)
  pboot.theta2[i] <- pweibull(5, pboot.shape[i], pboot.scale[i], lower.t=FALSE)
  pboot.eta1[i] <- quantile(boot.wind, .1)
  pboot.eta2[i] <- qweibull(.1, pboot.shape[i], pboot.scale[i])
}
etaRange <- range(boot.eta1, boot.eta2, pboot.eta1, pboot.eta2)

ggplot() + stat_qq(aes(sample = pboot.theta1)) # discrete, slight neg skewness
ggplot() + stat_qq(aes(sample = pboot.theta2)) # continuous, normal
ggplot() + stat_qq(aes(sample = pboot.eta1)) # very normal, continuous
ggplot() + stat_qq(aes(sample = pboot.eta2)) # very normal, continuous

sd(pboot.theta1) # 0.03198990
sd(pboot.theta2) # 0.02604409
var(pboot.theta1) / var(pboot.theta2) # 1.508716
# Similar to the nonparametric bootstrap, though relative efficiency differs

sd(pboot.eta1) # 0.2823103
sd(pboot.eta2) # 0.2100147
var(pboot.eta1) / var(pboot.eta2) # 1.806982
```

```{r}

```