### 2020-02-29 ISLR/10_Unsupervised-Learning.rmd
completed ISLR course & book
completed ISLR course & book

---
title: ''
mainfont: Arial
fontsize: 12pt
fig_width: 9
fig_height: 3.5
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 10}
   \fancyfoot[C]{\rmfamily\color{headergrey}Unsupervised Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select

```

## Chapter 10

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Lab

```{r}
load(file.path(here::here(), "ISLR", "10.R.RData"))
```

Suppose we want to fit a linear regression, but the number of variables is much larger than the number of observations. In some cases, we may improve the fit by reducing the dimension of the features before.

In this problem, we use a data set with n = 300 and p = 200, so we have more observations than variables, but not by much. Load the data x, y, x.test, and y.test from 10.R.RData.

First, concatenate x and x.test using the rbind functions and perform a principal components analysis on the concatenated data frame (use the "scale=TRUE" option). To within 10% relative error, what proportion of the variance is explained by the first five principal components?

```{r}
dat <- rbind(x, x.test)

pca.out <- prcomp(dat, scale = T)

summary(pca.out) # 0.3499

sum(head((pca.out$sdev)^2/ sum(pca.out$sdev^2), length = 5)) # 0.3566807

par(mfrow = c(1,1))
plot(pca.out)
```

The previous answer suggests that a relatively small number of "latent variables" account for a substantial fraction of the features' variability. We might believe that these latent variables are more important than linear combinations of the features that have low variance.

We can try forgetting about the raw features and using the first five principal components (computed on rbind(x,x.test)) instead as low-dimensional derived features. What is the mean-squared test error if we regress y on the first five principal components, and use the resulting model to predict y.test?

```{r}
xols <- pca.out$x[1:300,1:5]
fit0 <- lm(y ~ xols)
summary(fit0)

yhat0 <- predict(fit0, x.test)
mean((yhat0-y.test)^2)
```

### Lab

```{r}
states <- row.names(USArrests)
states
```

```{r}
names(USArrests)
```

```{r}
apply(USArrests, 2, mean)
```

```{r}
apply(USArrests, 2, var)
```

```{r}
pr.out <- prcomp(USArrests, scale = T)
```

```{r}
names(pr.out)
```

```{r}
pr.out$center
```

```{r}
pr.out$rotation
```

```{r}
dim(pr.out$x)
```

```{r}
pr.out$rotation =-pr.out$rotation
pr.out$x =-pr.out$x
biplot(pr.out, scale = 0)
```

```{r}
pr.out$sdev
```

```{r}
pr.var = pr.out$sdev^2
pr.out$sdev
```

```{r}
pve <- pr.var/sum(pr.var)
pve
```

```{r}
plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0, 1), type = "b")
plot(cumsum(pve), xlab = "Principal Component", ylab = "Cumulative Proportion of Variance Explained", ylim = c(0, 1), type = 'b')
```

```{r}
set.seed(2)
x <- matrix(rnorm(50*2), ncol = 2)
x[1:25, 1] = x[1:25, 1] + 3
x[1:25, 2] = x[1:25, 2] - 4
```

```{r}
km.out <- kmeans(x, 2, nstart = 20)
```

```{r}
km.out$cluster
```

```{r}
plot(x, col = (km.out$cluster + 1), main = "K-Means Clustering Results with K=2", pch = 20, cex = 2)
```

```{r}
set.seed(4)

km.out <- kmeans(x, 3, nstart = 20)
km.out
```

```{r}
set.seed(3)

km.out <- kmeans(x, 3, nstart = 1)
km.out$tot.withinss

km.out <- kmeans(x, 3, nstart = 20)
km.out$tot.withinss
```

```{r}
hc.complete <- hclust(dist(x), method = "complete")
hc.average <- hclust(dist(x), method = "average")
hc.single <- hclust(dist(x), method = "single")
```

```{r}
par(mfrow = c(1, 3))
plot(hc.complete, main = "Complete Linkage", cex = .9)
plot(hc.average, main = "Average Linkage", cex = .9)
plot(hc.single, main = "Single Linkage", cex = .9)
```

```{r}
xsc <- scale(x)

par(mfrow = c(1, 1))
plot(hclust(dist(xsc), method = "complete"), main = "Hierarchial Clustering with Scaled Features")
```

```{r}
x <- matrix(rnorm(30*3), ncol = 3)
dd <- as.dist(1 - cor(t(x)))

plot(hclust(dd, method = "complete"))
```

3.) In this problem, you will perform K-means clustering manually, with K=2, on a small example with n=6 observations and p=2 features. The observations are as follows.

a.) Plot the observations.

```{r}
x <- cbind(c(1, 1, 0, 5, 6, 4), c(4, 3, 4, 1, 2, 0))
plot(x[,1], x[,2])
```

b.) Randomly assign a cluster label to each observation. Report the cluster labels for each observation.

```{r}
set.seed(1)
labels <- sample(2, nrow(x), replace = T)
labels
```

```{r}
plot(x[, 1], x[, 2], col = (labels + 1), pch = 20, cex = 2)
```

c.) Compute the centroid for each cluster.

```{r}
centroid1 <- c(mean(x[labels == 1, 1]), mean(x[labels == 1, 2]))
centroid2 <- c(mean(x[labels == 2, 1]), mean(x[labels == 2, 2]))
plot(x[,1], x[,2], col=(labels + 1), pch = 20, cex = 2)
points(centroid1[1], centroid1[2], col = 2, pch = 4)
points(centroid2[1], centroid2[2], col = 3, pch = 4)
```

d.) Assign each observation to the centroid to which it is closest, in terms of Euclidean distance. Report the cluster labels for each observation.

```{r}
labels <- c(1, 1, 1, 2, 2, 2)
plot(x[, 1], x[, 2], col = (labels + 1), pch = 20, cex = 2)
points(centroid1[1], centroid1[2], col = 2, pch = 4)
points(centroid2[1], centroid2[2], col = 3, pch = 4)
```

e.) Repeat (c) and (d) until the answers obtained stop changing.

```{r}
centroid1 <- c(mean(x[labels == 1, 1]), mean(x[labels == 1, 2]))
centroid2 <- c(mean(x[labels == 2, 1]), mean(x[labels == 2, 2]))
plot(x[,1], x[,2], col=(labels + 1), pch = 20, cex = 2)
points(centroid1[1], centroid1[2], col = 2, pch = 4)
points(centroid2[1], centroid2[2], col = 3, pch = 4)
```

f.) In your plot from (a), color the observations according to the clusters labels obtained.

```{r}
plot(x[, 1], x[, 2], col=(labels + 1), pch = 20, cex = 2)
```

In words, describe the results that you would expect if you performed K-means clustering of the eight shoppers in Figure 10.14, on the basis of their sock and computer purchases, with K=2. Give three answers, one for each of the variable scalings displayed. Explain.

```{r}
socks <- c(8, 11, 7, 6, 5, 6, 7, 8)
computers <- c(0, 0, 0, 0, 1, 1, 1, 1)
x <- cbind(socks, computers)
labels <- c(1, 1, 2, 2, 2, 2, 1, 1)
plot(x[, 1], x[, 2], col=(labels + 1), pch = 20, cex = 2, asp = 1)
```

If we take into consideration the scaled variables, the number of computers plays a much larger role than the number of socks, so we have the clusters {5,6,7,8} (purchased computer) and {1,2,3,4} (no computer purchased).

```{r}
x <- cbind(scale(socks, center = FALSE), scale(computers, center = FALSE))
sd(computers)
```

```{r}
labels <- c(1, 1, 2, 2, 2, 2, 1, 1)
plot(x[, 1], x[, 2], col=(labels + 1), pch = 20, cex = 2, asp = 1)
```

A researcher collects expression measurements for 1000 genes in 100 tissue samples. The data can be written as a 1000x1000 matrix, which we call X, in which each row represents a gene and each column a tissue sample. Each tissue sample was processed on a different day, and the columns of X are ordered so that the samples that were processed earliest are on the left, and the samples that were processed later are on the right. The tissue samples belong to two groups : control (C) and treatment (T). The C and T samples were processed in a random order across the days. The researcher wishes to determine whether each gene’s expression measurements differ between the treatment and control groups.

As a pre-analysis (before comparing T versus C), the researcher performs a principal component analysis of the data, and finds that the first principal component (a vector of length 100) has a strong linear trend from left to right, and explains 10% of the variation. The researcher now remembers that each patient sample was run on one of two machines, A and B, and machine A was used more often in the earlier times while B was used more often later. The researcher has a record of which sample was run on which machine.

Explain what it means that the first principal component “explains 10% of the variation”.
The first principal component “explains 10% of the variation” means 90% of the information in the gene data set is lost by projecting the tissue sample observations onto the first principal component. Another way of explaining it is 90% of the variance in the data is not contained in the first principal component.

The researcher decides to replace the (i,j)th element of X with
xij−zi1ϕj1
where zi1 is the ith score, and ϕj1 is the jth loading, for the first principal component. He will then perform a two-sample t-test on each gene in this new data set in order to determine whether its expression differs between the two conditions. Critique this idea, and suggest a better approach.
Given the flaw shown in pre-analysis of a time-wise linear trend amongst the tissue samples’ first principal component, I would advise the researcher to include the machine used (A vs B) as a feature of the data set. This should enhance the PVE of the first principal component before applying the two-sample t-test.

Design and run a small simulation experiment to demonstrate the superiority of your idea.

```{r}
set.seed(1)
Control <- matrix(rnorm(50 * 1000), ncol = 50)
Treatment <- matrix(rnorm(50 * 1000), ncol = 50)
X <- cbind(Control, Treatment)
X[1, ] <- seq(-18, 18 - .36, .36) # linear trend in one dimension
pr.out <- prcomp(scale(X))
summary(pr.out)$importance[, 1]
```

We have 9.911% variance explained by the first principal component. Now, adding in A vs B via 10 vs 0 encoding.

```{r}
X <- rbind(X, c(rep(10, 50), rep(0, 50)))
pr.out <- prcomp(scale(X))
summary(pr.out)$importance[, 1]
```

7.) In the chapter, we mentioned the use of correlation-based distance and Euclidean distance as dissimilarity measures for hierarchical clustering. It turns out that these two measures are almost equivalent : if each observation has been centered to have mean zero and standard deviation one, and if we let rij denote the correlation between the ith and jth observations, then the quantity 1−rij is proportional to the squared Euclidean distance between the ith and jth observations. On the “USArrests” data, show that this proportionality holds.

```{r}
set.seed(1)
dsc <- scale(USArrests)
d1 <- dist(dsc)^2
d2 <- as.dist(1 - cor(t(dsc)))
summary(d2 / d1)
```

In Section 10.2.3, a formula for calculating PVE was given in Equation 10.8. We also saw that the PVE can be obtained using the “sdev” output of the “prcomp()” function. On the “USArrests” data, calculate PVE in two ways :

a.) Using the “sdev” output of the “prcomp()” function, as was done in Section 10.2.3.

```{r}
pr.out <- prcomp(USArrests, scale = TRUE)
pr.var <- pr.out$sdev^2
pve <- pr.var / sum(pr.var)
sum(pr.var)
```

b.) By applying Equation 10.8 directly. That is, use the “prcomp()” function to compute the principal component loadings. Then, use those loadings in Equation 10.8 to obtain the PVE.

```{r}
loadings <- pr.out$rotation
USArrests2 <- scale(USArrests)
sumvar <- sum(apply(as.matrix(USArrests2)^2, 2, sum))
apply((as.matrix(USArrests2) %*% loadings)^2, 2, sum) / sumvar
```

9.) Consider the “USArrests” data. We will now perform hierarchical clustering on the states.

a.) Using hierarchical clustering with complete linkage and Euclidean distance, cluster the states.

```{r}
set.seed(2)
hc.complete <- hclust(dist(USArrests), method = "complete")
plot(hc.complete)
```

b.) Cut the dendrogram at a height that results in three distinct clusters. Which states belong to which clusters ?

```{r}
cutree(hc.complete, 3)
```

c.) Hierachically cluster the states using complete linkage and Euclidean distance, after scaling the variables to have standard deviation one.

```{r}
sd.data <- scale(USArrests)
hc.complete.sd <- hclust(dist(sd.data), method = "complete")
plot(hc.complete.sd)
```

d.) What effect does scaling the variables have on the hierarchical clustering obtained ? In your opinion, should the variables be scaled before the inter-observation dissimilarities are computed ? Provide a justification for your answer.

```{r}
cutree(hc.complete.sd, 3)
```

```{r}
table(cutree(hc.complete, 3), cutree(hc.complete.sd, 3))
```

10.) In this problem, you will generate simulated data, and then perform PCA and K-means clustering on the data.

a.) Generate a simulated data set with 20 observations in each of three classes (i.e. 60 observations total), and 50 variables.

```{r}
set.seed(2)
x <- matrix(rnorm(20 * 3 * 50, mean = 0, sd = 0.001), ncol = 50)
x[1:20, 2] <- 1
x[21:40, 1] <- 2
x[21:40, 2] <- 2
x[41:60, 1] <- 1
true.labels <- c(rep(1, 20), rep(2, 20), rep(3, 20))
```

b.) Perform PCA on the 60 observations and plot the first two principal component score vectors. Use a different color to indicate the observations in each of the three classes. If the three classes appear separated in this plot, then continue on to part (c). If not, the return to part (a) and modify the simulation so that there is greater separation between the three classes. Do not continue to part (c) until the three classes show at least some separation in the first two principal component score vectors.

```{r}
pr.out <- prcomp(x)
plot(pr.out$x[, 1:2], col = 1:3, xlab = "Z1", ylab = "Z2", pch = 19)
```

c.) Perform K-means clustering of the observations with K=3. How well do the clusters that you obtained in K-means clustering compare to the true class labels ?

```{r}
km.out <- kmeans(x, 3, nstart = 20)
table(true.labels, km.out$cluster)
```

d.) Perform K-means clustering with K=2. Describe your results.

```{r}
km.out <- kmeans(x, 2, nstart = 20)
table(true.labels, km.out$cluster)
```

e.) Now perform K-means clustering with K=4, and describe your results.

```{r}
km.out <- kmeans(x, 4, nstart = 20)
table(true.labels, km.out$cluster)
```

f.) Now perform K-means clustering with K=3 on the first two principal component score vectors, rather than on the raw data. That is, perform K-means clustering on the 60x2 matrix of which the first column is the first principal component score vector, and the second column is the second principal component score vector. Comment on the results.

```{r}
km.out <- kmeans(pr.out$x[, 1:2], 3, nstart = 20)
table(true.labels, km.out$cluster)
```

g.) Using the “scale()” function, perform K-means clustering with K=3 on the data after scaling each variable to have standard deviation one. How do these results compare to those obtained in (b) ? Explain.

```{r}
km.out <- kmeans(scale(x), 3, nstart = 20)
table(true.labels, km.out$cluster)
```

11.) On the book website, there is a gene expression data set that consists of 40 tissue samples with measurements on 1000 genes. The first 20 samples are from healthy patients, while the second 20 are from a diseased group.

a.) Load the data using “read.csv()”. You will need to select “header = F”.

```{r}
genes <- read.csv(file.path(here::here(), "ISLR", "Ch10Ex11.csv"), header = FALSE)
```

b.) Apply hierarchical clustering to the samples using correlation-based distance, and plot the dendrogram. Do the genes separate the samples into two groups ? Do your results depend on the type of linkage used ?

```{r}
hc.complete <- hclust(as.dist(1 - cor(genes)), method = "complete")
plot(hc.complete)
```

```{r}
hc.single <- hclust(as.dist(1 - cor(genes)), method = "single")
plot(hc.single)
```

```{r}
hc.average <- hclust(as.dist(1 - cor(genes)), method = "average")
plot(hc.average)
```

The results are pretty different when using different linkage methods as we obtain two clusters for complete and single linkages or three clusters for average cluster.

c.) Your collaborator wants to know which genes differ the most across the two groups. Suggest a way to answer this question, and apply it here.
We may use PCA to see which genes differ the most. We will examine the absolute values of the total loadings for each gene as it characterizes the weight of each gene.

```{r}
pr.out <- prcomp(t(genes))
head(pr.out$rotation)
```

```{r}
total.load <- apply(pr.out$rotation, 1, sum)
index <- order(abs(total.load), decreasing = TRUE)
index[1:10]
```

```{r}
rm(list = ls())
```