### 2020-03-01 Hands-on Machine Learning/04_Linear-Regression.rmd

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Linear Regression}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 4}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)

# Additional Data
library(AppliedPredictiveModeling, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init()

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Linear Regression

### Data Set

h2o

```{r, echo = T}
ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)
```

stratified (_Sale_Price_) training sample

```{r, echo = T}
set.seed(123)

split <- initial_split(ames, prop = 0.7,
                       strata = "Sale_Price")

ames_train <- training(split)
ames_test <- testing(split)
```

## Simple Linear Model

```{r, echo = T}
model1 <- lm(Sale_Price ~ Gr_Liv_Area, data = ames_train)
```

```{r, echo = T, fig.height = 4.5}
# Fitted regression line (full training)
p1 <- model1 %>%
   broom::augment() %>%
   ggplot(aes(Gr_Liv_Area, Sale_Price)) +
   geom_point(size = 1, alpha = 0.3) +
   geom_smooth(se = F, method = "lm") +
   scale_y_continuous(labels = scales::dollar) +
   ggtitle("Fitted regression line")

# Fitted regression line (restricted range)
p2 <- model1 %>%
   broom::augment() %>%
   ggplot(aes(Gr_Liv_Area, Sale_Price)) +
   geom_segment(aes(x = Gr_Liv_Area, y = Sale_Price,
                    xend = Gr_Liv_Area, yend = .fitted),
                alpha = .3) +
   geom_point(size = 1, alpha = 0.3) +
   geom_smooth(se = F, method = "lm") +
   scale_y_continuous(labels = scales::dollar) +
   ggtitle("Fitted regression line (with residuals)")

grid.arrange(p1, p2, nrow = 2)
```

```{r, echo=T}
summary(model1)
```

```{r}
sigma(model1) # RMSE, also Residual Standard Error in summary()
sigma(model1)^2 # MSE
```

### Inference

The variability of an estimate is its _standard error (SE)_, the square root of its variance.

t-test for the coefficents are simply the estimated coefficent divided by the standard error (t value = Estimate / Std. Error)

t-test measure the number of standard deviations each coefficent is away from zero (basically abs(T) > 2 is significant at 95% conf)

The confidence interval for coefficents is:

\[\hat{\beta_j} \pm t_{1-\alpha/2, n-p}\hat{SE}(\hat{\beta_j})\]

```{r, echo = T}
confint(model1, level = .95)
```

Interpretation: We are 95% confident that each one unit increase in Gr_Liv_Area adds between 109.9 and 119.8 dollars to the sale price.

Linear Regression Assumptions:

+ 1.) Independent observations
+ 2.) The random errors have mean zero, and constant variance
+ 3.) The random errors are normally distributed

## Multiple Linear Regression

```{r, echo = T}
(model2 <- lm(Sale_Price ~ Gr_Liv_Area + Year_Built, data = ames_train))
# Equivalent
(model2 <- update(model1, . ~ . + Year_Built))

round(coef(model2), 3)
```

```{r, echo = T}
summary(model3 <- lm(Sale_Price ~ Gr_Liv_Area + Year_Built + Gr_Liv_Area:Year_Built, data = ames_train))

round(coef(model3), 3)
```

```{r, echo = T, fig.height=5}
fit1 <- lm(Sale_Price ~ Gr_Liv_Area + Year_Built, data = ames_train)
fit2 <- lm(Sale_Price ~ Gr_Liv_Area * Year_Built, data = ames_train)

# Regression plane
plot_grid <- expand.grid(
   Gr_Liv_Area = seq(from = min(ames_train$Gr_Liv_Area), to = max(ames_train$Gr_Liv_Area),
                     length = 100),
   Year_Built = seq(from = min(ames_train$Year_Built), to = max(ames_train$Year_Built),
                     length = 100)
)

plot_grid$y1 <- predict(fit1, newdata = plot_grid)
plot_grid$y2 <- predict(fit2, newdata = plot_grid)

# Level plots
p1 <- ggplot(plot_grid, aes(x = Gr_Liv_Area, y = Year_Built,
                      z = y1, fill = y1)) +
   geom_tile() +
   geom_contour(color = "white") +
   viridis::scale_fill_viridis(name = "Predicted\nvalue", option = "inferno") +
   theme_bw() +
   ggtitle("Main Effects Only")

p2 <- ggplot(plot_grid, aes(x = Gr_Liv_Area, y = Year_Built,
                      z = y2, fill = y2)) +
   geom_tile() +
   geom_contour(color = "white") +
   viridis::scale_fill_viridis(name = "Predicted\nvalue", option = "inferno") +
   theme_bw() +
   ggtitle("Main effect with two-way interaction")

gridExtra::grid.arrange(p1, p2, nrow = 2)
```

Full Model

```{r, echo = T}
model3 <- lm(Sale_Price ~ ., data = ames_train)
broom::tidy(model3)
```

### Assessing Model Accuracy

Models 1/2/3:

```{r, echo = T}
# Train model using 10-fold cross-validation

set.seed(123) # for reproducibility

(cv_model <- train(
   form = Sale_Price ~ Gr_Liv_Area,
   data = ames_train,
   method = "lm",
   trControl = trainControl(method = "cv", number = 10)
))

set.seed(123)

cv_model2 <- train(
   Sale_Price ~ Gr_Liv_Area + Year_Built,
   data = ames_train,
   method = "lm",
   trControl = trainControl(method = "cv", number = 10)
)

set.seed(123)

suppressWarnings({
   cv_model3 <- train(
      Sale_Price ~ .,
      data = ames_train,
      method = "lm",
      trControl = trainControl(method = "cv", number = 10)
   )
})
```

__Accuracy__:

```{r}
summary(resamples(list(
   model1 = cv_model,
   model2 = cv_model2,
   model3 = cv_model3
)))
```

### Model Concerns / Assumptions

#### 1.) Linear Relationships

Possible solution to non-linear relationship is by variable transformations:

```{r, echo = T}

p1 <- ggplot(ames_train, aes(Year_Built, Sale_Price)) +
   geom_point(size = 1, alpha = .4) +
   geom_smooth(se = F) +
   scale_y_continuous(labels = scales::dollar) +
   xlab("Year Built") +
   ggtitle(paste("Non-transformed variables with a \n", "non-linear relationship"))

p2 <- ggplot(ames_train, aes(Year_Built, Sale_Price)) +
   geom_point(size = 1, alpha = .4) +
   geom_smooth(se = F, method = "lm") +
   scale_y_log10("Sale Price", labels = scales::dollar, breaks = seq(0, 400000, by = 100000)) +
   xlab("Year Built") +
   ggtitle(paste("Transforming variables can provide a \n", "near-linear relationship"))

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

#### 2.) Constant variance among residuals

Linear models assume constant variance amoung error terms (_homoscedasticity_).

Notice the cone shape in Model 1:

```{r, echo = T}
df1 <- broom::augment(cv_model$finalModel, data = ames_train)

p1 <- ggplot(df1, aes(.fitted, .resid)) +
   geom_point(size = 1, alpha = .4) +
   xlab("Predicted Values") +
   ylab("Residuals") +
   ggtitle("Model 1", subtitle = "Sale_Price ~ Gr_Liv_Area")

df2 <- broom::augment(cv_model3$finalModel, data = ames_train)

p2 <- ggplot(df2, aes(.fitted, .resid)) +
   geom_point(size = 1, alpha = .4) +
   xlab("Predicted Values") +
   ylab("Residuals") +
   ggtitle("Model 3", subtitle = "Sale_Price ~ .")

gridExtra::grid.arrange(p1, p2, nrow =1)
```

#### 3.) No autocorrelation

Residuals should be uncorrelated and i.i.d.

```{r}
df1 <- mutate(df1, id = row_number())
df2 <- mutate(df2, id = row_number())

p1 <- ggplot(df1, aes(id, .resid)) +
   geom_point(size = 1, alpha = .4) +
   geom_smooth(se = F) +
   xlab("Row ID") +
   ylab("Residuals") +
   ggtitle("Model 1", subtitle = "Correlated Residuals")

p2 <- ggplot(df2, aes(id, .resid)) +
   geom_point(size = 1, alpha = .4) +
   geom_smooth(se = F) +
   xlab("Row ID") +
   ylab("Residuals") +
   ggtitle("Model 3", subtitle = "Uncorrelated Residuals")

gridExtra::grid.arrange(p1, p2, nrow = 1)

```

Model 1 has a distict pattern to the residuals (due to being ordered by neighborhood, which is unaccounted for in the model)

#### 4.) More observations than predictors

The number of features cannot exceed the number of observations.

(not a problem in this example)

#### 5.)

No or little multicollinearity.

Collinearity refers to the situation where two or more predictor variables are closely related to one another.

```{r}
cor(ames_train$Garage_Area, ames_train$Garage_Cars)
```

Example: _Garage_Area_ and _Garage_Cars_ are highly correlated.

```{r}
summary(cv_model3) %>%
   broom::tidy() %>%
   filter(term %in% c("Garage_Area", "Garage_Cars"))
```

Only _Garage_Area_ is significant (p < 0.05),

However, refit w/o _Garage_Area_ term, 
```{r}
# model without Garage Area
set.seed(123)

mod_wo_Garage_Cars <- train(
   Sale_Price ~ .,
   data = select(ames_train, -Garage_Area),
   method = "lm",
   trControl = trainControl(method = "cv", number = 10)
)

summary(mod_wo_Garage_Cars) %>%
   broom::tidy() %>%
   filter(term == "Garage_Cars")
```

_Garage_Cars_ becomes significant ( p < 0.001 )

Best to check VIF scores (variance inflation factor)

### Principal Component Analysis

PCA -> Linear factorization of features

PCR -> Principal Component Regression

```{r}
# performs 10-fold cross validation on a PCR model tuning the
# number of principal components to use as predictors from 1-20

set.seed(123)

cv_model_pcr <- train(
   Sale_Price ~ .,
   data = ames_train,
   method = "pcr",
   trControl = trainControl(method = "cv", number = 10),
   preProcess = c("zv", "center", "scale"),
   tuneLength = 20
)

cv_model_pcr$bestTune

ggplot(cv_model_pcr)

summary(cv_model_pcr)
```

### Partial least squares

PLS can be viewed as a supervised dimension reduction procedure.

Below is a comparision between PCR and PLS:

```{r}
data(solubility)

df <- cbind(solTrainX, solTrainY)

pca_df <- recipe(solTrainY ~ ., data = df) %>%
   step_center(all_predictors()) %>%
   step_scale(all_predictors()) %>%
   step_pca(all_predictors()) %>%
   prep(training = df, retain = T) %>%
   juice() %>%
   select(PC1, PC2, solTrainY) %>%
   rename(`PCR Component 1` = "PC1", `PCR Component 2` = "PC2") %>%
   gather(component, value, -solTrainY)

pls_df <- recipe(solTrainY ~., data = df) %>%
   step_center(all_predictors()) %>%
   step_scale(all_predictors()) %>%
   step_pls(all_predictors(), outcome = "solTrainY") %>%
   prep(training = df, retain = T) %>%
   juice() %>%
   rename(`PLS Component 1` = "PLS1", `PLS Component 2` = "PLS2") %>%
   gather(component, value, -solTrainY)

pca_df %>%
   bind_rows(pls_df) %>%
   ggplot(aes(value, solTrainY)) +
   geom_point(alpha = .15) +
   geom_smooth(method = "lm", se = F, lty = "dashed") +
   facet_wrap(~ component, scales = "free") +
   labs(x = "PC Eigenvalues", y = "Response")
```

Fit PLS to ames:

```{r}

# perform 10-fold cross validation on a PLS model tuning the
# number of principal compoents to use as predictors from 1-20

set.seed(123)

cv_model_pls <- train(
   Sale_Price ~ .,
   data = ames_train,
   method = "pls",
   trControl = trainControl(method = "cv", number = 10),
   preProcess = c("zv", "center", "scale"),
   tuneLength = 20
)

# model with lowest RMSE

cv_model_pls$bestTune

ggplot(cv_model_pls)

```

### Feature interpretation

Linear models are monotonic, meaning 1 unit change in X means a constant change in Y.

In PLS we can measure the importance of features by an absolute weighted sum of the regression coefficents.

We can use vip  to extract and plot the most important variables:

```{r}
vip(cv_model_pls, num_features = 20, method = "model")
```

Alternatively, we can use PDP (partial dependence plots) to summarize the relationship.

```{r}
p1 <- pdp::partial(cv_model_pls, pred.var = "Gr_Liv_Area", grid.resolution = 20) %>% 
  autoplot() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

p2 <- pdp::partial(cv_model_pls, pred.var = "First_Flr_SF", grid.resolution = 20) %>% 
  autoplot() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

p3 <- pdp::partial(cv_model_pls, pred.var = "Total_Bsmt_SF", grid.resolution = 20) %>% 
  autoplot() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

p4 <- pdp::partial(cv_model_pls, pred.var = "Garage_Cars", grid.resolution = 4) %>% 
  autoplot() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

grid.arrange(p1, p2, p3, p4, nrow = 2)
```

We see all 4 of the top predictive features have a positive linear relationship.

```{r}
# clean up
rm(list = ls())
```