### 2020-03-01 Hands-on Machine Learning/11_Random-Forests.rmd

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Random Forests}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 11}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)
library(foreach, quietly = T, warn.conflicts = F)
library(doParallel, quietly = T, warn.conflicts = F)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(earth, quietly = TRUE, warn.conflicts = FALSE)
library(ranger, quietly = T, warn.conflicts = F)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

##  Random Forests

### Data Sets

Attrition

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition.h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")
churn.train <- training(churn)
churn.test <- testing(churn)
```

Ames, Iowa housing data.

```{r, echo = T}
set.seed(123)

ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)

ames.split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames.train <- training(ames.split)
ames.test <- testing(ames.split)
```

### Random Forest Overview

Random forests are modifications of bagged decision trees that build a large collection of _de-correlated_ trees to further improve the predictive performance.

### Extended Bagging

The bootstrap aggregation procedure (bagging) has a limited effect on the variance reduction of decision trees.

Random forests help reduce tree correlation by injecting more randomness into the tree-growing process. More specifically, while growing a decision tree during the bagging process, random forests perform split-variable randomization where each time a split is to be peformed, the search for the split variable is limited to a random subset of $m_{try}$ of the original p features.

Typical default: $m_{try} = \frac{p}{3}$ (regression) and $m_{try} = \sqrt{p}$ for classification.

Basic algorithm is as follows::

+ 1.) Given a training data set

+ 2.) Select number of trees to build (n_trees)

+ 3.) for i = 1 to n_trees do:

+ 4.) Generate a bootstrap sample of the original data

+ 5.) Grow a regression/classification tree to the bootstrapped data

+ 6.) for each split

+ 7.) Select m_try variables at random for all p variables

+ 8.) Pick the best variable/split-point among the m_try

+ 9.) split the node into two child nodes

+ 10.) end

+ 11.) Use typical tree model stopping criteria to determine when a tree is complete (do not prune)

+ 12.) end

+ 13.) output ensemble of trees

Note: when $m_{try} = p$, the algorithm is equivalent to bagging decission trees.

### Out-of-the-box Performance

```{r}
# number of features
n.features <- length(setdiff(names(ames.train), "Sale_Price"))

# train a default random forest model
ames.rf1 <- ranger(
  Sale_Price ~ ., 
  data = ames.train,
  mtry = floor(n.features / 3),
  respect.unordered.factors = "order",
  seed = 123
)

# get OOB RMSE
(default.rmse <- sqrt(ames.rf1$prediction.error))
```

### Hyperparamteres

+ 1.) The number of trees in the forest

+ 2.) The number of features to consider at any given split

+ 3.) The complexity of each tree

+ 4.) The sampling scheme

+ 5.) The splitting rule to use during tree construction

#### Number of Trees

Typically start with 10x the number of features

```{r}
# number of features
n.features <- ncol(ames.train) - 1

# tuning grid
tuning.grid <- expand.grid(
  trees = seq(10, 1000, by = 20),
  rmse  = NA
)

for(i in seq_len(nrow(tuning.grid))) {

  # Fit a random forest
  fit <- ranger(
    formula = Sale_Price ~ ., 
    data = ames.train, 
    num.trees = tuning.grid$trees[i],
    mtry = floor(n.features / 3),
    respect.unordered.factors = 'order',
    verbose = FALSE,
    seed = 123
  )

  # Extract OOB RMSE
  tuning.grid$rmse[i] <- sqrt(fit$prediction.error)
}

ggplot(tuning.grid, aes(trees, rmse)) +
  geom_line(size = 1) +
  ylab("OOB Error (RMSE)") +
  xlab("Number of trees")
```

#### $m_{try}$

Hyperparameter that control the split-variable randomization feature of random forests.

With a high number of predictive features a high value of $m_{try}$ is likely to perform better. When many features are relevant, lower might perform better.

```{r}
tuning.grid <- expand.grid(
  trees = seq(10, 1000, by = 20),
  mtry  = floor(c(seq(2, 80, length.out = 5), 26)),
  rmse  = NA
)

for(i in seq_len(nrow(tuning.grid))) {
  fit <- ranger(
    formula    = Sale_Price ~ ., 
    data       = ames.train, 
    num.trees  = tuning.grid$trees[i],
    mtry       = tuning.grid$mtry[i],
    respect.unordered.factors = 'order',
    verbose    = FALSE,
    seed       = 123
  )

  tuning.grid$rmse[i] <- sqrt(fit$prediction.error)

}

labels <- tuning.grid %>%
  filter(trees == 990) %>%
  mutate(mtry = as.factor(mtry))

tuning.grid %>%
  mutate(mtry = as.factor(mtry)) %>%
  ggplot(aes(trees, rmse, color = mtry)) +
  geom_line(size = 1, show.legend = FALSE) +
  ggrepel::geom_text_repel(data = labels, aes(trees, rmse, label = mtry), nudge_x = 50, show.legend = FALSE) +
  ylab("OOB Error (RMSE)") +
  xlab("Number of trees")
```

#### Tree Complexity

Random forests are built on individual decision trees; consequently most random forest implementations have one or more hyperparameters.

```{r}

tuning.grid <- expand.grid(
  min.node.size = 1:20,
  run_time  = NA,
  rmse = NA
)

for(i in seq_len(nrow(tuning.grid))) {
  fit.time <- system.time({
    fit <- ranger(
      formula    = Sale_Price ~ ., 
      data       = ames.train, 
      num.trees  = 1000,
      mtry       = 26,
      min.node.size = tuning.grid$min.node.size[i],
      respect.unordered.factors = 'order',
      verbose    = FALSE,
      seed       = 123
  )
})

  tuning.grid$run_time[i] <- fit.time[[3]]
  tuning.grid$rmse[i] <- sqrt(fit$prediction.error)
}

min.node.size <- tuning.grid %>% 
  mutate(
    error_first = first(rmse),
    runtime_first = first(run_time),
    `Error Growth` = (rmse / error_first) - 1,
    `Run Time Reduction` = (run_time / runtime_first) - 1
    )

p1 <-  ggplot(min.node.size, aes(min.node.size, `Error Growth`)) +
  geom_smooth(size = 1, se = FALSE, color = "black") +
  scale_y_continuous("Percent growth in error estimate", labels = scales::percent) +
  xlab("Minimum node size") +
  ggtitle("A) Impact to error estimate")

p2 <-  ggplot(min.node.size, aes(min.node.size, `Run Time Reduction`)) +
  geom_smooth(size = 1, se = FALSE, color = "black") +
  scale_y_continuous("Reduction in run time", labels = scales::percent) +
  xlab("Minimum node size") +
  ggtitle("B) Impact to run time")

gridExtra::grid.arrange(p1, p2, nrow = 1)

```

#### Sampling Scheme

```{r}
tuning.grid <- expand.grid(
  sample.fraction = seq(.05, .95, by = .05),
  replace  = c(TRUE, FALSE),
  rmse = NA
)

for(i in seq_len(nrow(tuning.grid))) {
  fit <- ranger(
    formula    = Sale_Price ~ ., 
    data       = ames.train, 
    num.trees  = 1000,
    mtry       = 26,
    sample.fraction = tuning.grid$sample.fraction[i],
    replace = tuning.grid$replace[i],
    respect.unordered.factors = 'order',
    verbose    = FALSE,
    seed       = 123
  )

  tuning.grid$rmse[i] <- sqrt(fit$prediction.error)
}

tuning.grid %>%
  ggplot(aes(sample.fraction, rmse, color = replace)) +
  geom_line(size = 1) +
  scale_x_continuous("Sample Fraction", breaks = seq(.1, .9, by = .1), labels = scales::percent) +
  ylab("OOB Error (RMSE)") +
  scale_color_discrete("Sample with Replacement") +
  theme(legend.position = c(0.8, 0.85),
        legend.key = element_blank(),
        legend.background = element_blank())
```

### Tuning Strategies

Simple Brute Force

```{r}
# create hyperparameter grid
hyper_grid <- expand.grid(
  mtry = floor(n.features * c(.05, .15, .25, .333, .4)),
  min.node.size = c(1, 3, 5, 10), 
  replace = c(TRUE, FALSE),                               
  sample.fraction = c(.5, .63, .8),                       
  rmse = NA                                               
)

# execute full cartesian grid search
for(i in seq_len(nrow(hyper_grid))) {
  # fit model for ith hyperparameter combination
  fit <- ranger(
    formula         = Sale_Price ~ ., 
    data            = ames.train, 
    num.trees       = n.features * 10,
    mtry            = hyper_grid$mtry[i],
    min.node.size   = hyper_grid$min.node.size[i],
    replace         = hyper_grid$replace[i],
    sample.fraction = hyper_grid$sample.fraction[i],
    verbose         = FALSE,
    seed            = 123,
    respect.unordered.factors = 'order',
  )
  # export OOB error 
  hyper_grid$rmse[i] <- sqrt(fit$prediction.error)
}

# assess top 10 models
hyper_grid %>%
  arrange(rmse) %>%
  mutate(perc_gain = (default.rmse - rmse) / default.rmse * 100) %>%
  head(10)
```

#### h20 Random Grid Search

```{r}
h2o.no_progress()
h2o.init(max_mem_size = "5g")
```

```{r}
# convert training data to h2o object
train_h2o <- as.h2o(ames.train)

# set the response column to Sale_Price
response <- "Sale_Price"

# set the predictor names
predictors <- setdiff(colnames(ames.train), response)
```

```{r}
h2o_rf1 <- h2o.randomForest(
    x = predictors, 
    y = response,
    training_frame = train_h2o, 
    ntrees = n.features * 10,
    seed = 123
)

h2o_rf1
```

```{r}
# hyperparameter grid
hyper_grid <- list(
  mtries = floor(n.features * c(.05, .15, .25, .333, .4)),
  min_rows = c(1, 3, 5, 10),
  max_depth = c(10, 20, 30),
  sample_rate = c(.55, .632, .70, .80)
)

# random grid search strategy
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "mse",
  stopping_tolerance = 0.001,   # stop if improvement is < 0.1%
  stopping_rounds = 10,         # over the last 10 models
  max_runtime_secs = 60*5      # or stop search after 5 min.
)
```

```{r}
# perform grid search 
random_grid <- h2o.grid(
  algorithm = "randomForest",
  grid_id = "rf_random_grid",
  x = predictors, 
  y = response, 
  training_frame = train_h2o,
  hyper_params = hyper_grid,
  ntrees = n.features * 10,
  seed = 123,
  stopping_metric = "RMSE",   
  stopping_rounds = 10,           # stop if last 10 trees added 
  stopping_tolerance = 0.005,     # don't improve RMSE by 0.5%
  search_criteria = search_criteria
)
```

```{r}
# collect the results and sort by our model performance metric 
# of choice
random_grid_perf <- h2o.getGrid(
  grid_id = "rf_random_grid", 
  sort_by = "mse", 
  decreasing = FALSE
)
random_grid_perf
```

```{r}
# re-run model with impurity-based variable importance
rf_impurity <- ranger(
  formula = Sale_Price ~ ., 
  data = ames.train, 
  num.trees = 2000,
  mtry = 32,
  min.node.size = 1,
  sample.fraction = .80,
  replace = FALSE,
  importance = "impurity",
  respect.unordered.factors = "order",
  verbose = FALSE,
  seed  = 123
)

# re-run model with permutation-based variable importance
rf_permutation <- ranger(
  formula = Sale_Price ~ ., 
  data = ames.train, 
  num.trees = 2000,
  mtry = 32,
  min.node.size = 1,
  sample.fraction = .80,
  replace = FALSE,
  importance = "permutation",
  respect.unordered.factors = "order",
  verbose = FALSE,
  seed  = 123
)
```

```{r, fig.height=7}
p1 <- vip::vip(rf_impurity, num_features = 25, bar = FALSE)
p2 <- vip::vip(rf_permutation, num_features = 25, bar = FALSE)

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

```{r}
# clean up
rm(list = ls())
```