### 2020-03-01 Hands-on Machine Learning/12_Gradient-Boosting.rmd

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Gradient Boosting}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 12}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)
library(foreach, quietly = T, warn.conflicts = F)
library(doParallel, quietly = T, warn.conflicts = F)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(earth, quietly = TRUE, warn.conflicts = FALSE)
library(ranger, quietly = T, warn.conflicts = F)
library(gbm, quietly = T, warn.conflicts = F)
library(xgboost, quietly = T, warn.conflicts = F)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Gradient Boosting

### Data Sets

Attrition

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition_h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")

churn_train <- training(churn)
churn_test <- testing(churn)

rm(churn)
```

Ames, Iowa housing data.

```{r, echo = T}
set.seed(123)

ames <- AmesHousing::make_ames()
ames_h2o <- as.h2o(ames)

ames_split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames_train <- training(ames_split)
ames_test <- testing(ames_split)

rm(ames_split)

h2o.init(max_mem_size = "10g", strict_version_check = F)

train_h2o <- as.h2o(ames_train)
response <- "Sale_Price"
predictors <- setdiff(colnames(ames_train), response)
```

### Gradient Boosting Overview

Whereas random forests build an ensemble of deep independent trees, GBMs build an ensemble of shallow trees in sequence with each tree learning and improving on the previous one. Although shallow trees by themselves are rather weak predictive models, they can be "boosted" to produce a powerful "committee"

### Boosting vs Bagging

Boosting is a general concept that aggregates the predictions of simpler models. It is typically used with models that have high bias and low variance. This works particularly well with decision trees.

Boosting is a sequential algorithm that attacks the bias-variance trade-off by starting with a _weak_ model, and sequentially boost its performance by continuiing to build new trees, where each new tree in the sequence tries to "fix up" where the previous one made the biggest error.

### Base Learners

Boosting is a framework that iteratively improves any weak learning model. Many gradient boosting applications allow you to "plug in" various classes of weak learners at your disposal.

In practice, these are almost always decision trees as base learners.

### Traning Weak Models

A weak model is one whose error rate is only slighly better than random guessing. The idea behind boosting is that eadch model in the sequence slighly improves upon the performance of the previous one.

### Sequential Training

Boosted trees are grown sequentially; each tree is grown using information from previously grown trees to improve performance.

### Basic GBM Algorithm

+ 1.) Fit a decision tree to the data: $F_1(x) = y$

+ 2.) We then fit the next decision tree to the residuals of the previous: $h_1(x) = y - F_1(x)$

+ 3.) Add this new tree to our algorithm: $F_2(x) = F_1(x) + h_1(x)$

+ 4.) Fit the next decision tree to the residuals of $F_2 : h_2(x) = y - F_2(x)$

+ 5.) Add this new tree to our algorithm: $F_3(x) = F_2(x) + h_1(x)$

+ 6.) Continue this process until some mechanism (i.e., cross validation) tells us to stop.

The final model here is a stagewise additive model of b individual trees:

$f(x) = \sum^{B}_{b=1}f^b(x)$

Visually, the process looks like this:

```{r, fig.height=7}
# Simulate sine wave data
set.seed(1112)  # for reproducibility
df <- tibble::tibble(
  x = seq(from = 0, to = 2 * pi, length = 1000),
  y = sin(x) + rnorm(length(x), sd = 0.5),
  truth = sin(x)
)

# Function to boost `rpart::rpart()` trees
rpartBoost <- function(x, y, data, num_trees = 100, learn_rate = 0.1, tree_depth = 6) {
  x <- data[[deparse(substitute(x))]]
  y <- data[[deparse(substitute(y))]]
  G_b_hat <- matrix(0, nrow = length(y), ncol = num_trees + 1)
  r <- y
  for(tree in seq_len(num_trees)) {
    g_b_tilde <- rpart(r ~ x, control = list(cp = 0, maxdepth = tree_depth))
    g_b_hat <- learn_rate * predict(g_b_tilde)
    G_b_hat[, tree + 1] <- G_b_hat[, tree] + matrix(g_b_hat)
    r <- r - g_b_hat
    colnames(G_b_hat) <- paste0("tree_", c(0, seq_len(num_trees)))
  }
  cbind(df, as.data.frame(G_b_hat)) %>%
    gather(tree, prediction, starts_with("tree")) %>%
    mutate(tree = stringr::str_extract(tree, "\\d+") %>% as.numeric())
}

# Plot boosted tree sequence
rpartBoost(x, y, data = df, num_trees = 2^10, learn_rate = 0.05, tree_depth = 1) %>%
  filter(tree %in% c(0, 2^c(0:10))) %>%
  ggplot(aes(x, prediction)) +
    ylab("y") +
    geom_point(data = df, aes(x, y), alpha = .1) +
    geom_line(data = df, aes(x, truth), color = "blue") +
    geom_line(colour = "red", size = 1) +
    facet_wrap(~ tree, nrow = 3)
```

The Gradient in GBM is a gradient descent algorithm (from Calculus), where we can use it to find the minima/maximua of a function. Here, we apply the gradient descent to the loss function in the algorithm.

```{r}
# create data to plot
x <- seq(-5, 5, by = .05)
y <- x^2 + 3
df <- data.frame(x, y)
step <- 5
step_size <- .2
for(i in seq_len(18)) {
  next_step <- max(step) + round(diff(range(max(step), which.min(df$y))) * step_size, 0)
  step <- c(step, next_step)
  next
}
steps <- df[step, ] %>%
  mutate(x2 = lag(x), y2 = lag(y)) %>%
  dplyr::slice(1:18)
# plot
ggplot(df, aes(x, y)) +
  geom_line(size = 1.5, alpha = .5) +
  theme_classic() +
  scale_y_continuous("Loss function", limits = c(0, 30)) +
  xlab(expression(theta)) +
  geom_segment(data = df[c(5, which.min(df$y)), ], aes(x = x, y = y, xend = x, yend = -Inf), lty = "dashed") +
  geom_point(data = filter(df, y == min(y)), aes(x, y), size = 4, shape = 21, fill = "yellow") +
  geom_point(data = steps, aes(x, y), size = 3, shape = 21, fill = "blue", alpha = .5) +
  geom_curve(data = steps, aes(x = x, y = y, xend = x2, yend = y2), curvature = 1, lty = "dotted") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank()
  ) +
  annotate("text", x = df[5, "x"], y = 1, label = "Initial value", hjust = -0.1, vjust = .8) +
  annotate("text", x = df[which.min(df$y), "x"], y = 1, label = "Minimium", hjust = -0.1, vjust = .8) +
  annotate("text", x = df[5, "x"], y = df[5, "y"], label = "Learning step", hjust = -.8, vjust = 0)
```

Gradient descent is the process of gradually decreasing the cost function (i.e, MSE) by tweaking parameter(s) iteratively until you have reached a minimuim.

Gradient descent can be performed on any loss function that is differentiable.

An important parameter is the step size with is controlled by the learning rate. If the learning rate is too small, then the algorithm will take many interations to find the minimum. If its too big, it can stip over a possible minimum value.

Visually:

```{r}
# create too small of a learning rate
step <- 5
step_size <- .05
for(i in seq_len(10)) {
  next_step <- max(step) + round(diff(range(max(step), which.min(df$y))) * step_size, 0)
  step <- c(step, next_step)
  next
}
too_small <- df[step, ] %>%
  mutate(x2 = lag(x), y2 = lag(y))
# plot
p1 <- ggplot(df, aes(x, y)) +
  geom_line(size = 1.5, alpha = .5) +
  theme_classic() +
  scale_y_continuous("Loss function", limits = c(0, 30)) +
  xlab(expression(theta)) +
  geom_segment(data = too_small[1, ], aes(x = x, y = y, xend = x, yend = -Inf), lty = "dashed") +
  geom_point(data = too_small, aes(x, y), size = 3, shape = 21, fill = "blue", alpha = .5) +
  geom_curve(data = too_small, aes(x = x, y = y, xend = x2, yend = y2), curvature = 1, lty = "dotted") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank()
  ) +
  annotate("text", x = df[5, "x"], y = 1, label = "Start", hjust = -0.1, vjust = .8) +
  ggtitle("b) too small")
# create too large of a learning rate
too_large <- df[round(which.min(df$y) * (1 + c(-.9, -.6, -.2, .3)), 0), ] %>%
  mutate(x2 = lag(x), y2 = lag(y))
# plot
p2 <- ggplot(df, aes(x, y)) +
  geom_line(size = 1.5, alpha = .5) +
  theme_classic() +
  scale_y_continuous("Loss function", limits = c(0, 30)) +
  xlab(expression(theta)) +
  geom_segment(data = too_large[1, ], aes(x = x, y = y, xend = x, yend = -Inf), lty = "dashed") +
  geom_point(data = too_large, aes(x, y), size = 3, shape = 21, fill = "blue", alpha = .5) +
  geom_curve(data = too_large, aes(x = x, y = y, xend = x2, yend = y2), curvature = 1, lty = "dotted") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank()
  ) +
  annotate("text", x = too_large[1, "x"], y = 1, label = "Start", hjust = -0.1, vjust = .8) +
  ggtitle("a) too big")
gridExtra::grid.arrange(p2, p1, nrow = 1)
```

Not all loss functions are convex, however. There are many local minimas, plateaus, and other irregular terain of the loss function that makes finding the global minimum difficult.

**Stochastic gradient descent** can help us address this problem by sampling a fraction of the training observations (typically w/o replacement), and growing the next tree using that subsample.

Visually:

```{r}
# create random walk data
set.seed(123)
x <- sample(seq(3, 5, by = .05), 10, replace = TRUE)
set.seed(123)
y <- seq(2, 28, length.out = 10)

random_walk <- data.frame(
  x = x,
  y = y[order(y, decreasing = TRUE)]
)

optimal <- data.frame(x = 0, y = 0)

# plot
ggplot(df, aes(x, y)) + 
  coord_polar() +
  theme_minimal() +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank()
  ) +
  xlab(expression(theta[1])) +
  ylab(expression(theta[2])) +
  geom_point(data = random_walk, aes(x, y), size = 3, shape = 21, fill = "blue", alpha = .5) + 
  geom_point(data = optimal, aes(x, y), size = 2, shape = 21, fill = "yellow") + 
  geom_path(data = random_walk, aes(x, y), lty = "dotted") +
  annotate("text", x = random_walk[1, "x"], y = random_walk[1, "y"], label = "Start", hjust = 1, vjust = -1) +
  annotate("text", x = optimal[1, "x"], y = optimal[1, "y"], label = "Minimum", hjust = -.2, vjust = 1) +
  ylim(c(0, 28)) + 
  xlim(-5, 5)
```

Stochastic gradent descent will often find a near-optimal solution by jumping out of local minimas and off plateaus.

### Basic GBM

The first boosting algorithm was AdaBoost, which was then generalized into a regression framework.

### Hyperparameters

Typically there are two kinds of hyperparamters in GBM, boosting and tree-specific.

+ 1.) Number of Trees

Total number of trees in the ensemble.

+ 2.) Learning Rate

Determines the contribution of each tree on the final outcome and control how quickly the algorithm proceeds down the gradient descent.

Implementation:

```{r}
# run a basic GBM model
set.seed(123)  # for reproducibility
ames_gbm1 <- gbm(
  formula = Sale_Price ~ .,
  data = ames_train,
  distribution = "gaussian",  # SSE loss function
  n.trees = 5000,
  shrinkage = 0.1,
  interaction.depth = 3,
  n.minobsinnode = 10,
  cv.folds = 10
)

# find index for number trees with minimum CV error
best <- which.min(ames_gbm1$cv.error)

# get MSE and compute RMSE
sqrt(ames_gbm1$cv.error[best])
```

```{r}
# plot error curve
gbm.perf(ames_gbm1, method = "cv")
```

### General Tuning Strategy

```{r}
# create grid search
hyper_grid <- expand.grid(
  learning_rate = c(0.3, 0.1, 0.05, 0.01, 0.005),
  RMSE = NA,
  trees = NA,
  time = NA
)

# execute grid search
for(i in seq_len(nrow(hyper_grid))) {

  # fit gbm
  set.seed(123)  # for reproducibility
  train_time <- system.time({
    m <- gbm(
      formula = Sale_Price ~ .,
      data = ames_train,
      distribution = "gaussian",
      n.trees = 5000, 
      shrinkage = hyper_grid$learning_rate[i], 
      interaction.depth = 3, 
      n.minobsinnode = 10,
      cv.folds = 10 
   )
  })

  # add SSE, trees, and training time to results
  hyper_grid$RMSE[i]  <- sqrt(min(m$cv.error))
  hyper_grid$trees[i] <- which.min(m$cv.error)
  hyper_grid$Time[i]  <- train_time[["elapsed"]]

}

# results
arrange(hyper_grid, RMSE)
```

Exaustive Method

```{r}
# search grid
hyper_grid <- expand.grid(
  n.trees = 6000,
  shrinkage = 0.01,
  interaction.depth = c(3, 5, 7),
  n.minobsinnode = c(5, 10, 15)
)

# create model fit function
model_fit <- function(n.trees, shrinkage, interaction.depth, n.minobsinnode) {
  set.seed(123)
  m <- gbm(
    formula = Sale_Price ~ .,
    data = ames_train,
    distribution = "gaussian",
    n.trees = n.trees,
    shrinkage = shrinkage,
    interaction.depth = interaction.depth,
    n.minobsinnode = n.minobsinnode,
    cv.folds = 10
  )
  # compute RMSE
  sqrt(min(m$cv.error))
}

# perform search grid with functional programming
hyper_grid$rmse <- purrr::pmap_dbl(
  hyper_grid,
  ~ model_fit(
    n.trees = ..1,
    shrinkage = ..2,
    interaction.depth = ..3,
    n.minobsinnode = ..4
    )
)

# results
arrange(hyper_grid, rmse)
```

### Stochastic GBMs

```{r}
# refined hyperparameter grid
hyper_grid <- list(
  sample_rate = c(0.5, 0.75, 1),              # row subsampling
  col_sample_rate = c(0.5, 0.75, 1),          # col subsampling for each split
  col_sample_rate_per_tree = c(0.5, 0.75, 1)  # col subsampling for each tree
)

# random grid search strategy
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "mse",
  stopping_tolerance = 0.001,   
  stopping_rounds = 10,         
  max_runtime_secs = 60*60      
)

# perform grid search 
grid <- h2o.grid(
  algorithm = "gbm",
  grid_id = "gbm_grid",
  x = predictors, 
  y = response,
  training_frame = train_h2o,
  hyper_params = hyper_grid,
  ntrees = 6000,
  learn_rate = 0.01,
  max_depth = 7,
  min_rows = 5,
  nfolds = 10,
  stopping_rounds = 10,
  stopping_tolerance = 0,
  search_criteria = search_criteria,
  seed = 123
)

# collect the results and sort by our model performance metric of choice
grid_perf <- h2o.getGrid(
  grid_id = "gbm_grid", 
  sort_by = "mse", 
  decreasing = FALSE
)

grid_perf
```

```{r}
# Grab the model_id for the top model, chosen by cross validation error
best_model_id <- grid_perf@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)

# Now let’s get performance metrics on the best model
h2o.performance(model = best_model, xval = TRUE)
```

### XGBoost

#### Regularization

**XGBoost** has multiple tuning parameters to help prevent overfitting.

$\gamma$ = Lagrangian multiplier which controls the complexity of a given tree.

$\alpha$ = $L_1$ regularization (simmilar to ridge regresssion)

$\lambda$ = $L_2$ regularization (similar to lasso)

#### Dropout

Loosly defined as a regularization technique:

```{r}
xgb_prep <- recipe(Sale_Price ~ ., data = ames_train) %>%
  step_integer(all_nominal()) %>%
  prep(training = ames_train, retain = TRUE) %>%
  juice()

X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "Sale_Price")])
Y <- xgb_prep$Sale_Price
```

```{r}
set.seed(123)
ames_xgb <- xgb.cv(
  data = X,
  label = Y,
  nrounds = 6000,
  objective = "reg:linear",
  early_stopping_rounds = 50, 
  nfold = 10,
  params = list(
    eta = 0.1,
    max_depth = 3,
    min_child_weight = 3,
    subsample = 0.8,
    colsample_bytree = 1.0),
  verbose = 0
)

# minimum test CV RMSE
min(ames_xgb$evaluation_log$test_rmse_mean)
```

```{r}
# hyperparameter grid
hyper_grid <- expand.grid(
  eta = 0.01,
  max_depth = 3, 
  min_child_weight = 3,
  subsample = 0.5, 
  colsample_bytree = 0.5,
  gamma = c(0, 1, 10, 100, 1000),
  lambda = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  alpha = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  rmse = 0,          # a place to dump RMSE results
  trees = 0          # a place to dump required number of trees
)

# grid search
for(i in seq_len(nrow(hyper_grid))) {
  set.seed(123)
  m <- xgb.cv(
    data = X,
    label = Y,
    nrounds = 4000,
    objective = "reg:linear",
    early_stopping_rounds = 50, 
    nfold = 10,
    verbose = 0,
    params = list( 
      eta = hyper_grid$eta[i], 
      max_depth = hyper_grid$max_depth[i],
      min_child_weight = hyper_grid$min_child_weight[i],
      subsample = hyper_grid$subsample[i],
      colsample_bytree = hyper_grid$colsample_bytree[i],
      gamma = hyper_grid$gamma[i], 
      lambda = hyper_grid$lambda[i], 
      alpha = hyper_grid$alpha[i]
    ) 
  )
  hyper_grid$rmse[i] <- min(m$evaluation_log$test_rmse_mean)
  hyper_grid$trees[i] <- m$best_iteration
}

# results
hyper_grid %>%
  filter(rmse > 0) %>%
  arrange(rmse) %>%
  glimpse()
```

```{r}
# optimal parameter list
params <- list(
  eta = 0.01,
  max_depth = 3,
  min_child_weight = 3,
  subsample = 0.5,
  colsample_bytree = 0.5
)

# train final model
xgb.fit.final <- xgboost(
  params = params,
  data = X,
  label = Y,
  nrounds = 3944,
  objective = "reg:linear",
  verbose = 0
)
```

### Feature Interpretation

```{r}
# variable importance plot
vip::vip(xgb.fit.final) 
```

```{r}
h2o.shutdown(prompt = FALSE)
```

```{r}
# clean up
rm(list = ls())
```