### 2020-03-01 Hands-on Machine Learning/13_Deep-Learning.rmd

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Deep Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 13}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)
library(foreach, quietly = T, warn.conflicts = F)
library(doParallel, quietly = T, warn.conflicts = F)
library(compiler, quietly = T, warn.conflicts = F)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)

library(keras, quietly = TRUE, warn.conflicts = FALSE)
library(tfruns, quietly = TRUE, warn.conflicts = FALSE)
library(tfestimators, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Deep Learning

### Data Sets

MNIST

```{r, echo = T}
# Import MNIST training data
mnist <- dslabs::read_mnist()
mnist_x <- mnist$train$images
mnist_y <- mnist$train$labels

# Rename columns and standardize feature values
colnames(mnist_x) <- paste0("V", 1:ncol(mnist_x))
mnist_x <- mnist_x / 255

# One-hot encode response
mnist_y <- to_categorical(mnist_y, 10)

```

### Deep Learning Overview

Deep neural networks are designed to handle unique data sets. Multiple types exists RNN for text / speach, CNN image/video, etc.

```{r}
model <- keras_model_sequential() %>%
  layer_dense(units = 128, input_shape = ncol(mnist_x)) %>%
  layer_dense(units = 64) %>%
  layer_dense(units = 10)
```

```{r}
model <- keras_model_sequential() %>%
  layer_dense(units = 128, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dense(units = 10, activation = "softmax")
```

```{r}
model <- keras_model_sequential() %>%

  # Network architecture
  layer_dense(units = 128, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dense(units = 10, activation = "softmax") %>%

  # Backpropagation
  compile(
    loss = 'categorical_crossentropy',
    optimizer = optimizer_rmsprop(),
    metrics = c('accuracy')
  )
```

```{r}
# Train the model
fit1 <- model %>%
  fit(
    x = mnist_x,
    y = mnist_y,
    epochs = 25,
    batch_size = 128,
    validation_split = 0.2,
    verbose = FALSE
  )

# Display output
fit1
```

```{r}
plot(fit1)
```

### Model Tuning

```{r}
data_frame(
  Size = c("small", "medium", "large"), 
  `1` = c("16", "64", "256"),
  `2` = c("16, 8", "64, 32", "256, 128"),
  `3` = c("16, 8, 4", "64, 32, 16", "256, 128, 64")
) %>%
  kable(align = "c", caption = "Model capacities assessed represented as number of layers and nodes per layer.") %>%
  add_header_above(c(" ", "Hidden Layers" = 3)) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

### Batch Normalization

```{r}
model_w_norm <- keras_model_sequential() %>%

  # Network architecture with batch normalization
  layer_dense(units = 256, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 128, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%

  # Backpropagation
  compile(
    loss = "categorical_crossentropy",
    optimizer = optimizer_rmsprop(),
    metrics = c("accuracy")
  )
```

```{r}
# One layer models -------------------------------------------------------------
# small capacity model
`1 layer_small` <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# medium
`1 layer_medium` <- keras_model_sequential() %>%
  layer_dense(units = 64, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# large
`1 layer_large` <- keras_model_sequential() %>%
  layer_dense(units = 256, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# Two layer models -------------------------------------------------------------
# small capacity model
`2 layer_small` <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 8, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# medium
`2 layer_medium` <- keras_model_sequential() %>%
  layer_dense(units = 64, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# large
`2 layer_large` <- keras_model_sequential() %>%
  layer_dense(units = 256, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 128, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# Three layer models -------------------------------------------------------------
# small capacity model
`3 layer_small` <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 8, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 4, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# medium
`3 layer_medium` <- keras_model_sequential() %>%
  layer_dense(units = 64, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# large
`3 layer_large` <- keras_model_sequential() %>%
  layer_dense(units = 256, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 128, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

models <- ls(pattern = "layer_") 
df_batch <- models %>%
  map(get) %>%
  map(~ data.frame(
    `Validation error` = .$metrics$val_loss,
    `Training error`   = .$metrics$loss,
    epoch = seq_len(.$params$epoch)
    )) %>%
  map2_df(models, ~ mutate(.x, model = .y)) %>%
  separate(model, into = c("Middle layers", "Number of nodes"), sep = "_") %>%
  gather(Validation, Loss, Validation.error:Training.error) %>%
  mutate(
    Validation = str_replace_all(Validation, "\\.", " "),
    `Number of nodes` = factor(`Number of nodes`, levels = c("small", "medium", "large")),
    `Batch normalization` = TRUE
    )
```

```{r}
h2o.shutdown(prompt = FALSE)
```

```{r}
# clean up
rm(list = ls())
```