### 2020-03-02 Mathematics and Statistics for FRM/01_Basics.rmd
frm ch1
frm ch1

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Basics}
   \rfoot{\color{headergrey}Chapter 1}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematics and Statistics for Financial Risk Management}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 1

### 1

Solve for $y$, where:

a.) $y = ln(e^5)$

5

b.) $y = ln(1/e)$

$y = ln(1/e) = ln(1) - 1 = 0 - 1 = -1$

c.) $y = ln(10e)$

$y = ln(10e) = ln(10) + ln(e) = ln(10) + 1$

```{r}
log(10) + 1
```

### 2

The nominal monthly rate for a loan is quoted at 5%. What is the equivalent annual rate? Semiannual rate? Continious rate?

```{r}
effective_rate <- function(rate, periods) {
   ((1 + rate/periods)^periods - 1)
}
```

Annual:

```{r}
effective_rate(.05, 12) * 100
```

Semiannual:

```{r}
effective_rate(.05, 2) * 100
```

Continous:

```{r}
log(1 + effective_rate(.05, 12)) * 100
```

### 3

Over the course of a year, the log return on a stock market index is 11.2%. The starting value of the index is 100. What is the value at the end of the year?

```{r}
100 * exp(.112)
```

### 4

You have a portfolio of 10 bonds. In how many different ways can exactly two bonds default? Assume the order in which the bonds default is unimportant.

```{r}
choose(10, 2)
```

### 5

What is the present value of a perpetuity that pays $100 per year? Use an actual discount rate of 4%, and assume the first payment will be made in exactly one year.

```{r}
100 * (1/0.04)
```

### 6

ABC stock will pay a $1 dividend in one year. Assume the dividend will continue to be paid annually forever and the dividend payments will increase in size at a rate of 5%. Value this stream of dividends using a 6% annual discount rate.

```{r}
1 / (0.06 - 0.05)
```

### 7

What is the present value of a 10-year bond with a $100 face value, which pays a 6% coupon annually? Use an 8% discount rate.

```{r}

present_value <- function(face, int.rate, periods, discount.rate) {
   delta <- 1/(1 + discount.rate)

   coupon <- face * int.rate

   discount <- (delta - delta^(periods+1)) / (1 - delta)

   coupon * discount + ( face / (1 + discount.rate)^periods )
}

present_value(100, .06, 10, .08)
```

### 8

Solve for $x$, where $e^{e^x} = 10$

```{r}
log(log(10))
```

### 9

Calculate the value of the following summation: $\sum^{9}_{i=0}(-0.5)^i$

```{r}
r <- -0.5; n <- 9

(1 - r^n) / (1 - r)
```

### 10

The risk department of your firm has 10 analysts. You need to select four analysts to serve on a special audit committee. How many possible groupings of four analysts can be put together?

```{r}
choose(10, 4)
```

long-hand:

```{r}
n <- 10; k <- 4

factorial(10) / ( factorial(k) * factorial(n - k))
```

### 11

What is the present value of a newly issued 10-year bond with a notial value of $100 and a 2% annual coupon? Assume a constant 5% annual discount rate and no risk of default.

```{r}
present_value(100, .02, 10, 0.05)
```