### 2020-03-03 Hands-on Machine Learning/16_Interpretable-Machine-Learning.rmd
iml
iml

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Interpretable Machine Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 16}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling
library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)
library(svmpath, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(kernlab, quietly = TRUE, warn.conflicts = FALSE)
library(mlbench, quietly = TRUE, warn.conflicts = FALSE)

library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(iml, quietly = TRUE, warn.conflicts = FALSE)
library(DALEX, quietly = TRUE, warn.conflicts = FALSE)
library(lime, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Interpretable Machine Learning

### Data Sets

Attrition

```{r, attrition_data}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition_h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")

churn_train <- training(churn)
churn_test <- testing(churn)

rm(churn)
```

Ames, Iowa housing data.

```{r, ames_data}

ames <- AmesHousing::make_ames()
ames_h2o <- as.h2o(ames)

set.seed(123)

ames_split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames_train <- training(ames_split)
ames_test <- testing(ames_split)

rm(ames_split)

h2o.init(max_mem_size = "10g", strict_version_check = F)

train_h2o <- as.h2o(ames_train)
response <- "Sale_Price"
predictors <- setdiff(colnames(ames_train), response)
```

```{r, data_recipe}

# ensure consistent categorical levels
blueprint <- recipe(Sale_Price ~., data = ames_train) %>%
  step_other(all_nominal(), threshold = 0.005)

# Create training / test h2o frames
train_h2o <- prep(blueprint, training = ames_train, retain = T) %>%
  juice() %>%
  as.h2o()

test_h2o <- prep(blueprint, training = ames_train) %>%
  bake(new_data = ames_test) %>%
  as.h2o()

Y <- "Sale_Price"
X <- setdiff(names(ames_train), Y)
```

### h2o ML setup

```{r, train_h2o_models}

# Train & cross-validate a GLM model
best_glm <- h2o.glm(
  x = X, y = Y, training_frame = train_h2o, alpha = 0.1,
  remove_collinear_columns = TRUE, nfolds = 10, fold_assignment = "Modulo",
  keep_cross_validation_predictions = TRUE, seed = 123
)

# Train & cross-validate a RF model
best_rf <- h2o.randomForest(
  x = X, y = Y, training_frame = train_h2o, ntrees = 1000, mtries = 20,
  max_depth = 30, min_rows = 1, sample_rate = 0.8, nfolds = 10,
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
  seed = 123, stopping_rounds = 5, stopping_metric = "RMSE",
  stopping_tolerance = 0
)

# Train & cross-validate a GBM model
best_gbm <- h2o.gbm(
  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.01,
  max_depth = 7, min_rows = 5, sample_rate = 0.8, nfolds = 10,
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
  seed = 123, stopping_rounds = 5, stopping_metric = "RMSE",
  stopping_tolerance = 0
)

# Train & cross-validate an XGBoost model

# not avaliable on windows, yet..(3/2/20)

#best_xgb <- h2o.xgboost(
#  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.05,
#  max_depth = 3, min_rows = 3, sample_rate = 0.8, categorical_encoding = "Enum",
#  nfolds = 10, fold_assignment = "Modulo", 
#  keep_cross_validation_predictions = TRUE, seed = 123, stopping_rounds = 50,
#  stopping_metric = "RMSE", stopping_tolerance = 0
#)

# Train a stacked tree ensemble
ensemble_tree <- h2o.stackedEnsemble(
  x = X, y = Y, training_frame = train_h2o, model_id = "my_tree_ensemble_01",
  base_models = list(best_glm, best_rf, best_gbm),
  metalearner_algorithm = "drf"
)
```

### Local Interpretation

Predictions using ensemble tree:

```{r, local_predictions}

# Compute predictions
predictions <- predict(ensemble_tree, train_h2o) %>% as.vector()

# Print the highest and lowest predicted sales price
paste("Observation", which.max(predictions), 
      "has a predicted sale price of", scales::dollar(max(predictions)))

paste("Observation", which.min(predictions), 
      "has a predicted sale price of", scales::dollar(min(predictions)))

# Grab feature values for observations with min/max predicted sales price
high_ob <- as.data.frame(train_h2o)[which.max(predictions), ] %>% select(-Sale_Price)
low_ob  <- as.data.frame(train_h2o)[which.min(predictions), ] %>% select(-Sale_Price)
```

### Model specific vs. model agnostic

```{r, model_features}

# 1) create a data frame with just the features
features <- as.data.frame(train_h2o) %>% select(-Sale_Price)

# 2) Create a vector with the actual responses
response <- as.data.frame(train_h2o) %>% pull(Sale_Price)

# 3) Create custom predict function that returns the predicted values as a vector
pred <- function(object, newdata)  {
  results <- as.vector(h2o.predict(object, as.h2o(newdata)))
  return(results)
}

# Example of prediction output
pred(ensemble_tree, features) %>% head()
```

```{r, iml_dalex_components}
# iml model agnostic object
components_iml <- Predictor$new(
  model = ensemble_tree, 
  data = features, 
  y = response, 
  predict.fun = pred
)

# DALEX model agnostic object
components_dalex <- DALEX::explain(
  model = ensemble_tree,
  data = features,
  y = response,
  predict_function = pred
)
```

### Permutation based feature importance

Implementation

```{r, vip_plot}

vip(
  ensemble_tree,
  train = as.data.frame(train_h2o),
  method = "permute",
  target = "Sale_Price",
  metric = "RMSE",
  nsim = 5,
  sample_frac = 0.5,
  pred_wrapper = pred
)

```

Partial dependence

```{r, pdp_partial}

# Compute partial dependence values

pd_values <- partial(
  ensemble_tree,
  train = as.data.frame(train_h2o), 
  pred.var = "Gr_Liv_Area",
  pred.fun = function(object, newdata) {
      results <- mean(as.vector(h2o.predict(object, as.h2o(newdata))))
      return(results)
    },
  grid.resolution = 20
)

head(pd_values)  # take a peak

# Partial dependence plot
autoplot(pd_values, rug = TRUE, train = as.data.frame(train_h2o))
```

### Individual conditional expectation

```{r, ice_curves}

# Construct ICE curves
ice_non_centered <- partial(
  ensemble_tree,
  train = as.data.frame(train_h2o), 
  pred.var = "Gr_Liv_Area",
  pred.fun = pred,
  grid.resolution = 20
) %>%
  autoplot(alpha = 0.05, center = FALSE) +
  ggtitle("A) Non-centered ICE curves")

# Construct c-ICE curves
ice_centered <- partial(
  ensemble_tree,
  train = as.data.frame(train_h2o), 
  pred.var = "Gr_Liv_Area",
  pred.fun = pred,
  grid.resolution = 20
) %>%
  autoplot(alpha = 0.05, center = TRUE) +
  ggtitle("B) Centered ICE curves")

# Display plots side by side
gridExtra::grid.arrange(ice_non_centered, ice_centered, ncol = 2)
```

Implementation

### Feature interactions

```{r, pdp_ice_curves}
# Construct c-ICE curves
partial(
  ensemble_tree,
  train = as.data.frame(train_h2o), 
  pred.var = "Gr_Liv_Area",
  pred.fun = pred,
  grid.resolution = 20,
  plot = TRUE,
  center = TRUE,
  plot.engine = "ggplot2"
)
```

```{r, interactions}

interact <- Interaction$new(components_iml)

interact$results %>%
  arrange(desc(.interaction)) %>%
  head()

plot(interact)
```

### Local interpretable model-agnostic explanations

```{r, lime_explainer}

# Create explainer object
components_lime <- lime(
  x = features,
  model = ensemble_tree, 
  n_bins = 10
)

class(components_lime)

# Use LIME to explain previously defined instances: high_ob and low_ob
lime_explanation <- lime::explain(
  x = rbind(high_ob, low_ob), 
  explainer = components_lime, 
  n_permutations = 10,
  dist_fun = "gower",
  kernel_width = 0.25,
  n_features = 10, 
  feature_select = "highest_weights"
)

plot_features(lime_explanation, ncol = 1)
```

### Tuning

```{r, tuning}
# Tune the LIME algorithm a bit
lime_explanation2 <- explain(
  x = rbind(high_ob, low_ob), 
  explainer = components_lime, 
  n_permutations = 5000,
  dist_fun = "euclidean",
  kernel_width = 0.75,
  n_features = 10, 
  feature_select = "lasso_path"
)

# Plot the results
plot_features(lime_explanation2, ncol = 1)

# Two-way PDP using iml
interaction_pdp <- Partial$new(
  components_iml, 
  c("First_Flr_SF", "Overall_Qual"), 
  ice = FALSE, 
  grid.size = 20
)

labels <- interaction_pdp$results %>% filter(First_Flr_SF == max(First_Flr_SF))
plot(interaction_pdp) + 
  ggrepel::geom_label_repel(
    data = labels, 
    aes(label = Overall_Qual),
    label.size = .05, 
    label.padding = .15
)

# Construct c-ICE curves
partial(
  ensemble_tree,
  train = as.data.frame(train_h2o), 
  pred.var = "Gr_Liv_Area",
  pred.fun = pred,
  grid.resolution = 20,
  plot = TRUE,
  center = TRUE,
  plot.engine = "ggplot2"
)
```

### Shapley values

```{r}
# Compute (approximate) Shapley values
(shapley <- Shapley$new(components_iml, x.interest = high_ob, sample.size = 1000))

# Plot results
plot(shapley)

# Reuse existing object
shapley$explain(x.interest = low_ob)

# Plot results
shapley$results %>%
  top_n(25, wt = abs(phi)) %>%
  ggplot(aes(phi, reorder(feature.value, phi), color = phi > 0)) +
  geom_point(show.legend = FALSE)
```

#### Clean up

```{r}
h2o.shutdown(prompt = FALSE)
```

```{r}
# clean up
rm(list = ls())
```