### 2020-03-07 Hands-on Machine Learning/17-20_Unsupervised-Learning.rmd
homl unsupervised
homl unsupervised

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Principal Component Analysis}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapters 17-20}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling
library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)
library(svmpath, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(kernlab, quietly = TRUE, warn.conflicts = FALSE)
library(mlbench, quietly = TRUE, warn.conflicts = FALSE)

library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(iml, quietly = TRUE, warn.conflicts = FALSE)
library(DALEX, quietly = TRUE, warn.conflicts = FALSE)
library(lime, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Unsupervised Learning

### Data Sets

Attrition

```{r, attrition_data}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition_h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")

churn_train <- training(churn)
churn_test <- testing(churn)

rm(churn)
```

Ames, Iowa housing data.

```{r, ames_data}
ames <- AmesHousing::make_ames()
ames_h2o <- as.h2o(ames)

set.seed(123)

ames_split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames_train <- training(ames_split)
ames_test <- testing(ames_split)

rm(ames_split)

h2o.init(max_mem_size = "10g", strict_version_check = F)

train_h2o <- as.h2o(ames_train)
response <- "Sale_Price"
predictors <- setdiff(colnames(ames_train), response)
```

Market Basket:

```{r}
url <- "https://koalaverse.github.io/homlr/data/my_basket.csv"

my_basket <- readr::read_csv(url)

dim(my_basket)
```

MNIST:

```{r}
mnist <- dslabs::read_mnist()
```

### Principal Components Analysis

h2o PCA

```{r}
my_basket_h2o <- as.h2o(my_basket)

# PCA
my_pca <- h2o.prcomp(
  training_frame = my_basket_h2o,
  pca_method = "GramSVD",
  k = ncol(my_basket_h2o),
  transform = "STANDARDIZE",
  impute_missing = T,
  max_runtime_secs = 1000
)
```

Loadings

```{r}
my_pca@model$eigenvectors %>%
  as.data.frame() %>%
  mutate(feature = row.names(.)) %>%
  ggplot(aes(pc1, reorder(feature, pc1))) +
  geom_point()
```

```{r}
my_pca@model$eigenvectors %>%
  as.data.frame() %>%
  mutate(feature = row.names(.)) %>%
  ggplot(aes(pc1, pc2, label = feature)) +
  geom_text()
```

Eigenvalue Criterion

```{r}
eigen <- my_pca@model$importance["Standard deviation", ] %>%
  as.vector() %>%
  .^2

sum(eigen)

which(eigen >= 1)
```

Proportion of Variance Explained

```{r}
data.frame(
  PC = my_pca@model$importance %>% seq_along(),
  PVE = my_pca@model$importance %>% .[2,] %>% unlist(),
  CVE = my_pca@model$importance %>% .[3,] %>% unlist()
) %>%
  tidyr::gather(metric, variance_explained, -PC) %>%
  ggplot(aes(PC, variance_explained)) +
  geom_point() +
  facet_wrap( ~ metric, ncol = 1, scales = "free")
```

```{r}
data.frame(
  PC = my_pca@model$importance %>% seq_along,
  PVE = my_pca@model$importance %>% .[2,] %>% unlist()
) %>%
  ggplot(aes(PC, PVE, group = 1, label = PC)) +
  geom_point() +
  geom_line() +
  geom_text(nudge_y = 0.002)
```

## Generalized Low Rank Models

h2o implementation

```{r}
basic_glrm <- h2o.glrm(
  training_frame = my_basket_h2o,
  k = 20,
  loss = "Quadratic",
  regularization_x = "None",
  regularization_y = "None",
  transform = "STANDARDIZE",
  max_iterations = 2000,
  seed = 123
)

summary(basic_glrm)

plot(basic_glrm)
```

```{r}
data.frame(
  PC = basic_glrm@model$importance %>% seq_along(),
  PVE = basic_glrm@model$importance %>% .[2,] %>% unlist(),
  CVE = basic_glrm@model$importance %>% .[3,] %>% unlist()
) %>%
  gather(metric, variance_explained, -PC) %>%
  ggplot(aes(PC, variance_explained)) +
  geom_point() +
  facet_wrap( ~ metric, ncol = 1, scales = "free")
```

```{r}
t(basic_glrm@model$archetypes)[1:5, 1:5]
```

```{r}
p1 <- t(basic_glrm@model$archetypes) %>%
  as.data.frame() %>%
  mutate(feature = row.names(.)) %>%
  ggplot(aes(Arch1, reorder(feature, Arch1))) +
  geom_point()

p2 <- t(basic_glrm@model$archetypes) %>%
  as.data.frame() %>%
  mutate(feature = row.names(.)) %>%
  ggplot(aes(Arch1, Arch2, label = feature)) +
  geom_text()

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

```{r}
k8_glrm <- h2o.glrm(
  training_frame = my_basket_h2o,
  k = 8,
  loss = "Quadratic",
  regularization_x = "None",
  regularization_y = "None",
  transform = "STANDARDIZE",
  max_iterations = 2000,
  seed = 123
)

my_reconstruction <- h2o.reconstruct(k8_glrm, my_basket_h2o,
                                     reverse_transform = T)

my_reconstruction[1:5, 1:4] %>% round(0)
```

```{r}
# Use non-negative regularization
k8_glrm_regularized <- h2o.glrm(
  training_frame = my_basket_h2o,
  k = 8, 
  loss = "Quadratic",
  regularization_x = "NonNegative", 
  regularization_y = "NonNegative",
  gamma_x = 0.5,
  gamma_y = 0.5,
  transform = "STANDARDIZE", 
  max_iterations = 2000,
  seed = 123
)

# Show predicted values
predict(k8_glrm_regularized, my_basket_h2o)[1:5, 1:5]

# Compare regularized versus non-regularized loss
par(mfrow = c(1, 2))
plot(k8_glrm)
plot(k8_glrm_regularized)
```

```{r}
# Split data into train & validation
split <- h2o.splitFrame(my_basket_h2o, ratios = 0.75, seed = 123)
train <- split[[1]]
valid <- split[[2]]

# Create hyperparameter search grid
params <- expand.grid(
  regularization_x = c("None", "NonNegative", "L1"),
  regularization_y = c("None", "NonNegative", "L1"),
  gamma_x = seq(0, 1, by = .25),
  gamma_y = seq(0, 1, by = .25),
  error = 0,
  stringsAsFactors = FALSE
)

# Perform grid search
for(i in seq_len(nrow(params))) {

  # Create model
  glrm_model <- h2o.glrm(
    training_frame = train,
    k = 8, 
    loss = "Quadratic",
    regularization_x = params$regularization_x[i], 
    regularization_y = params$regularization_y[i],
    gamma_x = params$gamma_x[i],
    gamma_y = params$gamma_y[i],
    transform = "STANDARDIZE", 
    max_runtime_secs = 1000,
    seed = 123
  )

  # Predict on validation set and extract error
  validate <- h2o.performance(glrm_model, valid)
  params$error[i] <- validate@metrics$numerr
}

# Look at the top 10 models with the lowest error rate
params %>%
  arrange(error) %>%
  head(10)
```

```{r}
# Apply final model with optimal hyperparamters
final_glrm_model <- h2o.glrm(
  training_frame = my_basket_h2o,
  k = 8, 
  loss = "Quadratic",
  regularization_x = "L1", 
  regularization_y = "NonNegative",
  gamma_x = 1,
  gamma_y = 0.25,
  transform = "STANDARDIZE", 
  max_iterations = 2000,
  seed = 123
)

# New observations to score
new_observations <- as.h2o(sample_n(my_basket, 2))

# Basic scoring
predict(final_glrm_model, new_observations) %>% round(0)
```

### Autoencoders

```{r}
# Convert mnist features to an h2o input data set
features <- as.h2o(mnist$train$images)

# Train an autoencoder
ae1 <- h2o.deeplearning(
  x = seq_along(features),
  training_frame = features,
  autoencoder = TRUE,
  hidden = 2,
  activation = 'Tanh',
  sparse = TRUE
)

# Extract the deep features
ae1_codings <- h2o.deepfeatures(ae1, features, layer = 1)
ae1_codings
```

### Stacked Autoencoders

```{r}
# Hyperparameter search grid
hyper_grid <- list(hidden = list(
  c(50),
  c(100), 
  c(300, 100, 300),
  c(100, 50, 100),
  c(250, 100, 50, 100, 250)
))

# Execute grid search
ae_grid <- h2o.grid(
  algorithm = 'deeplearning',
  x = seq_along(features),
  training_frame = features,
  grid_id = 'autoencoder_grid',
  autoencoder = TRUE,
  activation = 'Tanh',
  hyper_params = hyper_grid,
  sparse = TRUE,
  ignore_const_cols = FALSE,
  seed = 123
)

# Print grid details
h2o.getGrid('autoencoder_grid', sort_by = 'mse', decreasing = FALSE)
```

Visualizing Reconstruction

```{r}
# Get sampled test images
index <- sample(1:nrow(mnist$test$images), 4)
sampled_digits <- mnist$test$images[index, ]
colnames(sampled_digits) <- paste0("V", seq_len(ncol(sampled_digits)))

# Predict reconstructed pixel values
best_model_id <- ae_grid@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)
reconstructed_digits <- predict(best_model, as.h2o(sampled_digits))
names(reconstructed_digits) <- paste0("V", seq_len(ncol(reconstructed_digits)))

combine <- rbind(sampled_digits, as.matrix(reconstructed_digits))

# Plot original versus reconstructed
par(mfrow = c(1, 3), mar=c(1, 1, 1, 1))
layout(matrix(seq_len(nrow(combine)), 4, 2, byrow = FALSE))
for(i in seq_len(nrow(combine))) {
  image(matrix(combine[i, ], 28, 28)[, 28:1], xaxt="n", yaxt="n")
}
```

```{r}
ae100_codings <- h2o.deepfeatures(best_model, features, layer = 1)
ae100_codings %>% 
    as.data.frame() %>% 
    tidyr::gather() %>%
    summarize(average_activation = mean(value))
```

```{r}
# Hyperparameter search grid
hyper_grid <- list(sparsity_beta = c(0.01, 0.05, 0.1, 0.2))

# Execute grid search
ae_sparsity_grid <- h2o.grid(
  algorithm = 'deeplearning',
  x = seq_along(features),
  training_frame = features,
  grid_id = 'sparsity_grid',
  autoencoder = TRUE,
  hidden = 100,
  activation = 'Tanh',
  hyper_params = hyper_grid,
  sparse = TRUE,
  average_activation = -0.1,
  ignore_const_cols = FALSE,
  seed = 123
)

# Print grid details
h2o.getGrid('sparsity_grid', sort_by = 'mse', decreasing = FALSE)
```

```{r}
# Extract reconstruction errors
(reconstruction_errors <- h2o.anomaly(best_model, features))

# Plot distribution
reconstruction_errors <- as.data.frame(reconstruction_errors)

ggplot(reconstruction_errors, aes(Reconstruction.MSE)) +
  geom_histogram(aes(fill = ..count..), bins = 30)

```

#### Clean up

```{r}
h2o.shutdown(prompt = FALSE)
```

```{r}
# clean up
rm(list = ls())
```