### 2020-03-08 Hands-on Machine Learning/21-22_Clustering.rmd
homl, clustering
homl, clustering

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Clustering}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapters 17-20}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling
library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)
library(svmpath, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)

library(cluster, quietly = TRUE, warn.conflicts = FALSE)
library(factoextra, quietly = TRUE, warn.conflicts = FALSE)
library(yardstick, quietly = TRUE, warn.conflicts = FALSE)
library(NbClust, quietly = TRUE, warn.conflicts = FALSE)
library(mclust, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Clustering

### Data Sets

Attrition

```{r, attrition_data}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")

churn_train <- training(churn)
churn_test <- testing(churn)

rm(churn)
```

Ames, Iowa housing data.

```{r, ames_data}
ames <- AmesHousing::make_ames()

set.seed(123)

ames_split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames_train <- training(ames_split)
ames_test <- testing(ames_split)

rm(ames_split)

response <- "Sale_Price"
predictors <- setdiff(colnames(ames_train), response)
```

Market Basket:

```{r}
url <- "https://koalaverse.github.io/homlr/data/my_basket.csv"

my_basket <- readr::read_csv(url)

dim(my_basket)
```

MNIST:

```{r}
mnist <- dslabs::read_mnist()
```

```{r}
data(geyser, package = 'MASS')
```

### K-means Clustering

```{r}
features <- mnist$train$images

# Use k-means model with 10 centers and 10 random starts
mnist_clustering <- kmeans(features, centers = 10, nstart = 10)

# Print contents of the model output
str(mnist_clustering)
```

```{r}
# Extract cluster centers
mnist_centers <- mnist_clustering$centers

# Plot typical cluster digits
par(mfrow = c(2, 5), mar=c(0.5, 0.5, 0.5, 0.5))
layout(matrix(seq_len(nrow(mnist_centers)), 2, 5, byrow = FALSE))
for(i in seq_len(nrow(mnist_centers))) {
  image(matrix(mnist_centers[i, ], 28, 28)[, 28:1], 
        col = gray.colors(12, rev = TRUE), xaxt="n", yaxt="n")
}
```

```{r}
# Create mode function
mode_fun <- function(x){  
  which.max(tabulate(x))
}

mnist_comparison <- data.frame(
  cluster = mnist_clustering$cluster,
  actual = mnist$train$labels
) %>%
  group_by(cluster) %>%
  mutate(mode = mode_fun(actual)) %>%
  ungroup() %>%
  mutate_all(factor, levels = 0:9)

# Create confusion matrix and plot results
yardstick::conf_mat(
  mnist_comparison, 
  truth = actual, 
  estimate = mode
) %>%
  autoplot(type = 'heatmap')
```

```{r}
fviz_nbclust(
  my_basket, 
  kmeans, 
  k.max = 25,
  method = "wss",
  diss = get_dist(my_basket, method = "spearman")
)
```

# Mixed Data

```{r}
# Full ames data set --> recode ordinal variables to numeric
ames_full <- AmesHousing::make_ames() %>%
  mutate_if(str_detect(names(.), 'Qual|Cond|QC|Qu'), as.numeric)

# One-hot encode --> retain only the features and not sale price
full_rank  <- caret::dummyVars(Sale_Price ~ ., data = ames_full, 
                               fullRank = TRUE)
ames_1hot <- predict(full_rank, ames_full)

# Scale data
ames_1hot_scaled <- scale(ames_1hot)

# New dimensions
dim(ames_1hot_scaled)
```

```{r}
set.seed(123)

fviz_nbclust(
  ames_1hot_scaled, 
  kmeans, 
  method = "wss", 
  k.max = 25, 
  verbose = FALSE
)
```

```{r}
# Original data minus Sale_Price
ames_full <- AmesHousing::make_ames() %>% select(-Sale_Price)

# Compute Gower distance for original data
gower_dst <- daisy(ames_full, metric = "gower")
```

```{r}
# You can supply the Gower distance matrix to several clustering algos
pam_gower <- pam(x = gower_dst, k = 8, diss = TRUE)
diana_gower <- diana(x = gower_dst, diss = TRUE)
agnes_gower <- agnes(x = gower_dst, diss = TRUE)
```

```{r}
fviz_nbclust(
  ames_1hot_scaled, 
  pam, 
  method = "wss", 
  k.max = 25, 
  verbose = FALSE
)
```

```{r}
# k-means computation time on MNIST data
system.time(kmeans(features, centers = 10))

# CLARA computation time on MNIST data
system.time(clara(features, k = 10))
```

### Hierachial Clustering

```{r}
ames_scale <- AmesHousing::make_ames() %>%
  select_if(is.numeric) %>%  # select numeric columns
  select(-Sale_Price) %>%    # remove target column
  mutate_all(as.double) %>%  # coerce to double type
  scale()                    # center & scale the resulting columns
```

```{r}
# For reproducibility
set.seed(123)

# Dissimilarity matrix
d <- dist(ames_scale, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d, method = "complete" )
```

```{r}
# For reproducibility
set.seed(123)

# Compute maximum or complete linkage clustering with agnes
hc2 <- agnes(ames_scale, method = "complete")

# Agglomerative coefficient
hc2$ac
```

```{r}
# methods to assess
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(ames_scale, method = x)$ac
}

# get agglomerative coefficient for each linkage method
purrr::map_dbl(m, ac)
```

```{r}
# compute divisive hierarchical clustering
hc4 <- diana(ames_scale)

# Divise coefficient; amount of clustering structure found
hc4$dc
```

### Optimal Clusters

```{r}
# Plot cluster results
p1 <- fviz_nbclust(ames_scale, FUN = hcut, method = "wss", 
                   k.max = 10) +
  ggtitle("(A) Elbow method")
p2 <- fviz_nbclust(ames_scale, FUN = hcut, method = "silhouette", 
                   k.max = 10) +
  ggtitle("(B) Silhouette method")
p3 <- fviz_nbclust(ames_scale, FUN = hcut, method = "gap_stat", 
                   k.max = 10) +
  ggtitle("(C) Gap statistic")

# Display plots side by side
gridExtra::grid.arrange(p1, p2, p3, nrow = 1)
```

```{r}
# Construct dendorgram for the Ames housing example
hc5 <- hclust(d, method = "ward.D2" )
dend_plot <- fviz_dend(hc5)
dend_data <- attr(dend_plot, "dendrogram")
dend_cuts <- cut(dend_data, h = 8)
fviz_dend(dend_cuts$lower[[2]])
```

```{r}
# Ward's method
hc5 <- hclust(d, method = "ward.D2" )

# Cut tree into 4 groups
sub_grp <- cutree(hc5, k = 8)

# Number of members in each cluster
table(sub_grp)
```

```{r}
# Plot full dendogram
fviz_dend(
  hc5,
  k = 8,
  horiz = TRUE,
  rect = TRUE,
  rect_fill = TRUE,
  rect_border = "jco",
  k_colors = "jco",
  cex = 0.1
)
```

```{r}
dend_plot <- fviz_dend(hc5)                # create full dendogram
dend_data <- attr(dend_plot, "dendrogram") # extract plot info
dend_cuts <- cut(dend_data, h = 70.5)      # cut the dendogram at 
                                           # designated height
# Create sub dendrogram plots
p1 <- fviz_dend(dend_cuts$lower[[1]])
p2 <- fviz_dend(dend_cuts$lower[[1]], type = 'circular')

# Side by side plots
gridExtra::grid.arrange(p1, p2, nrow = 1)
```

### Model-Based Clustering

```{r}
data(geyser, package = 'MASS')

# Apply GMM model with 3 components
geyser_mc <- Mclust(geyser, G = 3)

par(mfrow = c(1, 2))
# Plot results
plot(geyser_mc, what = "density")
plot(geyser_mc, what = "uncertainty")
```

```{r}
# Observations with high uncertainty
sort(geyser_mc$uncertainty, decreasing = TRUE) %>% head()
```

```{r}
summary(geyser_mc)
```

```{r}
geyser_optimal_mc <- Mclust(geyser)

summary(geyser_optimal_mc)
```

```{r}
my_basket_mc <- Mclust(my_basket, 1:20)

summary(my_basket_mc)
## ---------------------------------------------------- 
## Gaussian finite mixture model fitted by EM algorithm 
## ---------------------------------------------------- 
## 
## Mclust EEV (ellipsoidal, equal volume and shape) model with 6 components: 
## 
##  log-likelihood    n   df      BIC       ICL
##        8308.915 2000 5465 -24921.1 -25038.38
## 
## Clustering table:
##   1   2   3   4   5   6 
## 391 403  75 315 365 451

plot(my_basket_mc, what = 'BIC', 
     legendArgs = list(x = "bottomright", ncol = 5))
```

```{r}
probabilities <- my_basket_mc$z 
colnames(probabilities) <- paste0('C', 1:6)

probabilities <- probabilities %>%
  as.data.frame() %>%
  mutate(id = row_number()) %>%
  tidyr::gather(cluster, probability, -id)

ggplot(probabilities, aes(probability)) +
  geom_histogram() +
  facet_wrap(~ cluster, nrow = 2)
```

```{r}
uncertainty <- data.frame(
  id = 1:nrow(my_basket),
  cluster = my_basket_mc$classification,
  uncertainty = my_basket_mc$uncertainty
)

uncertainty %>%
  group_by(cluster) %>%
  filter(uncertainty > 0.25) %>%
  ggplot(aes(uncertainty, reorder(id, uncertainty))) +
  geom_point() +
  facet_wrap(~ cluster, scales = 'free_y', nrow = 1)
```

```{r}
cluster2 <- my_basket %>%
  scale() %>%
  as.data.frame() %>%

  mutate(cluster = my_basket_mc$classification) %>%
  filter(cluster == 2) %>%
  select(-cluster)

cluster2 %>%
  tidyr::gather(product, std_count) %>%
  group_by(product) %>%
  summarize(avg = mean(std_count)) %>%
  ggplot(aes(avg, reorder(product, avg))) +
  geom_point() +
  labs(x = "Average standardized consumption", y = NULL)
```

#### Clean up

```{r}
# clean up
rm(list = ls())
```