### 2020-04-06 Computational Methods/01_intro-numerical-analysis.rmd

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Intoduction to Numerical Analysis}
   \rfoot{\color{headergrey}Chapter 1}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Computational Methods for Numerical Analysis}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

library(cmna, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- file.path(here::here(), "Case Studies", "datasets")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Introduction to Numerical Analysis

### Numerical Analysis in R

Example, factorial:

```{r}
factorial <- function(n) {
   if(n == 0)
      return(0)
   if(n == 1)
      return(1)
   return(factorial(n - 1) + factorial(n - 2))
}

sapply(1:10, factorial)
```

```{r}
isPrime <- function(n) {
   if( n == 2)
      return(TRUE)

   for(i in 2:sqrt(n))
      if(n %% i == 0)
         return(FALSE)

   return(TRUE)
}

nums <- 1:25

rbind(nums, sapply(nums, isPrime))
```

```{r}
x <- c(TRUE, FALSE, TRUE, FALSE, TRUE)
sum(x); mean(x); length(x)
```

```{r}
naievesum <- function(x) {

   s <- 0
   n <- length(x)

   for(i in 1:n) {
      s <- S + x[i]
   }

   s
}

naivesum(c(1, 2, 3, 4.5, -6))
```

```{r}
pwisesum <- function(x) {

   n <- length(x)

   if(n == 1)
      return(x)

   m = floor(n/2)

   return(pwisesum(x[1:m]) + pwisesum(x[(m + 1):n]))
}

pwisesum(c(1, 2, 3, 4.5, -6))

```

```{r}
kahan_sum <- function(x) {
   comp <- s <- 0
   n <- length(x)

   for(i in 1:n) {
      y <- x[i] - comp
      t <- x[i] + s
      comp <- (t - s) - y
      s <- t
   }

   return(s)
}

kahan_sum(1:8)
```

```{r}
naive_poly <- function(x, coefs) {
   y <- rep(0, length(x))

   for(i in 1:length(coefs)) {
      y <- y + coefs[i] * (x ^ (i - 1))
   }

   return(y)
}

# example poly:

# f(x) = (x-5)(x+3)(x+2)(x-1)
# = x^4 + x^3 - x^2 - 11x + 30
# = 30 - 11 x - 19 x^2 - x^3 + x^4

f <- c(30, -11, -19, -1, 1)

x <- c(-1, 0, 1)

naive_poly(x, f)

```

```{r}
better_poly <- function(x, coefs) {
   y <- rep(0, length(x))

   cached.x <- 1

   for(i in 1:length(coefs)) {
      y <- y + coefs[i] * cached.x
      cached.x <- cached.x * x
   }

   return(y)
}

better_poly(x, f)
```

```{r}
horner <- function(x, coefs) {
   y <- rep(0, length(x))

   for(i in length(coefs):1)
      y <- coefs[i] + x * y

   return(y)
}

horner(x, f)
```

```{r}
nth_root <- function(a, n, tol = 1 / 1000) {
   x <- 1
   deltax <- tol * 10

   while(abs(deltax) > tol) {
      deltax <- (1 / n) * (a / x ^ (n - 1) - x)
      x <- x + deltax
   }

   return(x)
}

nth_root(100, 2)
nth_root(65536, 4)
nth_root(1000, 3)
nth_root(pi, 2)
```

## Exercises

#### 1.)

What is the circumference of a 10-meter-diameter circle symbolically?

$2\pi r$, r = 10

$10\pi$

What is it numerically?

```{r}
10*pi
```

Which one is correct? *Both*

#### 2.)

Describe an application for which a numerical solution are preferable to a symbolic solution.

_An exact symbolic solution to some problems can be difficult or near impossible, so a good approximation via numerical methods is preferable than no solution._

Describe an application where symbolic solutions are preferable to numerical.

_An analytical/symbolic solution can solve a related family of problems._

#### 3.)

If an unsigned variable were stored using 16 bits, what is the largest number it could hold?

```{r}
2^16
```

#### 4.)

Using R's internal trigonometric functions, implement a secant function.

```{r}
sec <- function(x) 1/cos(x)

assertthat::are_equal(sec(pi), -1)

```

#### 5.)

Using the exponentiation operator, create a function that finds the nth root of any number.

```{r}
nth_root <- function(x, n) x^(1/n)

assertthat::are_equal(nth_root(32, 5), 2)

```

#### 6.)

Function 1.1 provides a recursive implementation of the Fibonacci sequence. Rewrite the function to provide a result without recursion.

```{r}
fib_iter <- function(x) {

   # build look-up
   fib_values <- list()
   fib_values[[1]] <- 1

   for(index in 2:x) {

      fib1 <- ifelse(index <= 1, 1, fib_values[[index - 1]])
      fib2 <- ifelse(index <= 1, 1, ifelse(index == 2, 1, fib_values[[index - 2]]))

      fib_values[index] <- fib1 + fib2
   }

   return(fib_values[[x]])
}

sapply(1:10, fib_iter)
```

What are the relative strenghts and weaknesses of each implementation?

_The recursive version is a bit more readable._

#### 7.)

Implement a product function in R without using recursion.

```{r}

```

#### 8.)
#### 9.)
#### 10.)
#### 11.)
#### 12.)
#### 13.)
#### 14.)
#### 15.)
#### 16.)
#### 17.)
#### 18.)
#### 19.)