### 2020-04-06 Mathematics and Statistics for FRM/03_Statistics.rmd

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistics}
   \rfoot{\color{headergrey}Chapter 3}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematics and Statistics for Financial Risk Management}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 3

### 1

Compute the mean and the median of the following series of returns:

```{r}
dat <- data.table(Returns = c(.12, .5, -.08, .2, .04, .1, .02))

pretty_kable(dat, "Returns")
```

```{r}
mean(dat$Returns)

median(dat$Returns)
```

### 2

Compute the sample mean and the standard deviation of the following returns:

```{r}
dat <- data.table(Returns = c(.12, .5, -.08, .2, .04, .1, .02))

pretty_kable(dat, "Returns")
```

### 3

Prove that Equation 3.2 is an unbiased estimator of the mean. That is, show that $\mathbb{E}[\hat\mu] = \mu$

$\mathbb{E}(\bar{x}) = \mathbb{E}[\frac{x_1 + x_2 + x_3 + \ldots + x_n}{n}]$

$\ldots = \mathbb{E}\frac{1}{n}(x_1 + x_2 + x_3 + \ldots + x_n)$

$\ldots = \frac{1}{n}\mathbb{E}(x_1 + x_2 + x_3 + \ldots + x_n)$

$\ldots = \frac{1}{n}(\mathbb{E}[x_1] + \mathbb{E}[x_2] + \mathbb{E}[x_3] + \ldots + \mathbb{E}[x_n])$

$\ldots = \frac{1}{n}(\mu_1 + \mu_2 + \mu_3 + \ldots + \mu_n)$

$\ldots = \frac{1}{n}n(u)$

$\ldots = \mu$

### 4

What is the standard deviation of the estimator in Eq. 3.2? Assume the various data points are i.i.d.

$\mu = \frac{1}{n}\sum^{n}_{i=1}r_i$

$\frac{1}{n}\sum^{n}_{i=1}r_i = \frac{1}{n}(r_1 + r_2 + r_3 \ldots + r_n)$

$\sigma^2 = \frac{1}{n}\sum(x_i - \mu)^2$

$\sigma_{\mu} = \frac{\sigma_{}}{\sqrt{n}}$

### 5

Calculate the population covariance and correlation of the following series:

```{r}
dat <- data.table(S1 = c(.21, .53, .83, .19), S2 = c(.2, .32, .8, .4))

pretty_kable(dat, "Returns")
```

```{r}
cov(dat$S1, dat$S2)

cor(dat$S1, dat$S2)
```

### 6

Calculate the population mean, standard deviation and skewness of each of the following two series:

```{r}
dat <- data.table(S1 = c(-51, -21, 21, 51), S2 = c(-61, -7, 33, 35))

pretty_kable(dat, "Returns")
```

```{r}

skew <- function(x) {
   mean(x - mean(x))^3 / sd(x)^3
}

apply(dat, 2, mean)
apply(dat, 2, sd)
apply(dat, 2, skew)

```

### 7

Calculate the population mean, standard deviation and skewness of each of the following two series:

```{r}
dat <- data.table(S1 = c(-23, -7, 7, 23), S2 = c(-17, -17, 17, 17))

pretty_kable(dat, "Returns")
```

```{r}

skew <- function(x) {
   mean(x - mean(x))^3 / sd(x)^3   
}

apply(dat, 2, mean)
apply(dat, 2, sd)
apply(dat, 2, skew)

```

### 8

Given the probability density function for a random variable, X

$f(x) = \frac{x}{18}$ for $0 \le x \le 6$

$\mu = \int_0^6 \frac{x^3}{18 * 3} = \frac{6^3}{3*18} - \frac{0^3}{3*18}$

$\frac{6^2}{3^2} = 4$

$\sigma^2 = \int_0^6(x - 4)^2\frac{x}{18}dx$

$\frac{1}{18}(x^3 - 8x^2 + 16x)dx$

$\sigma^2 = 2(9 - 16 + 8) = 2$

### 9

Prove that Equation 3.19, reproduced here, is an unbiased estimator of the variance.

$\hat\sigma^2_x = \frac{1}{n - 1}\sum_{i=1}^n(x_i - \hat\mu_x)$

$\hat\sigma^2_x = \frac{1}{n - 1}\sum_{i=1}^n = \frac{1}{n - 1}\sum_{i=1}^n(\frac{n-1}{n}x_i - \frac{1}{n}\sum_{j \ne x}x_j)^2$

$\hat\sigma^2_x = \frac{1}{n}\sum_{i=1}^nx^2_i - \frac{1}{n(n - 1)}\sum{i=1}^n\sum_{i \ne j}x_ix_j$

$\mathbb{E}[x^2_i] = \sigma^2 + \mu^2$

$\mathbb{E}[x_i x_j] = \mu_i\mu_j + iff \ \sigma_{ij} = 0 \ \forall i \ne j$

### 10

Given two random variables, $X_A$ and $X_B$, with the corresponding means $\mu_A$ and $\mu_B$ and standard deviations $\sigma_A$ and $\sigma_B$, prove that the variance of $X_A$ plus $X_B$ is:

$\mathbb{Var}[X_A + X_B] = \sigma^2_A + \sigma^2_B + 2\rho_{AB}\sigma_A\sigma_B$

$\mathbb{E}[X_A + X_B] = \mathbb{E}[X_A] + \mathbb{E}[X_B] = \mu_A + \mu_B$

$\mathbb{Var}[X_A + X_B] = \mathbb{E}[(X_A + X_B - \mathbb{E}[X_A + X_b])^2]$

$\mathbb{Var}[X_A + X_B] = \sigma^2_A + \sigma^2_B + 2Cov[X_A, X_B]$

### 11

A $100 notional, zero coupon bond has one year to expiry. The probability of default is 10%. In the event of default, assume that the recovery rate is 40%.

The continiously compounded discount rate is 5%. What is the present value of this bond?

```{r}
EV <- .9 * 100 + 0.10 * 40
```

$\mathbb{E}(V) = `r EV`$

```{r}
PV <- exp(-.05) * EV
```

$\mathbb{P}(V)$ = $`r round(PV, 2)`