### 2020-06-27 Bayesian Data Analysis/03_Bayes_vs_Freq_.rmd
bayesian data analysis
bayesian data analysis

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Bayesian vs Frequentist}
   \rfoot{\color{headergrey}Chapter 6}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Bayesian Data Analysis}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 4.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(cowplot, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

library(lattice, quietly = TRUE, warn.conflicts = FALSE)

library(blmeco, quietly = TRUE, warn.conflicts = FALSE)
library(tidybayes, quietly = TRUE, warn.conflicts = FALSE)
library(arm, quietly = TRUE, warn.conflicts = FALSE)
library(coda, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

# The Bayesian Way

```{r}

# utility function for plotting a posterior

plot_posterior <- function(bsim) {
   res <- data.table(mu = bsim@coef, sigma = bsim@sigma)
   colnames(res) <- c("mu", "sigma")

   pmain <- ggplot(res, aes(mu, sigma)) +
     geom_point(alpha = .65, col = "darkgrey") +
     labs(x = "", y = "") +
     stat_density_2d(aes(fill = stat(level), alpha = ..level..), geom = "polygon")

   xbox <- axis_canvas(pmain, axis = "x") +
     geom_histogram(data = res, aes(mu, y = ..density.., fill = ..count..), size = .2, bins = 50, alpha = .85) +
     geom_density(data = res, aes(mu), col = "darkgrey", alpha = .85) +
     theme(legend.position = "none")

   ybox <- axis_canvas(pmain, axis = "y", coord_flip = T) +
     geom_histogram(data = res, aes(sigma, y = ..density.., fill = ..count..), size = .2, bins = 50, alpha = .85) +
     geom_density(data = res, aes(sigma), col = "darkgrey", alpha = .85) +
     coord_flip() +
     theme(legend.position = "none")

   suppressWarnings({
     p1 <- insert_xaxis_grob(pmain, xbox, grid::unit(.2, "null"), position = "top")
     p2 <- insert_yaxis_grob(p1, ybox, grid::unit(.2, "null"), position = "right")
   })

   ggdraw(p2)
}
```

## Bayes Therom

$P(\theta | y) = \frac{P(\theta)P(y|\theta)}{P(y)}$

### Estimating the Mean

The model of the data is:

$y \sim Norm(\theta, \sigma^2)$

Given the data has three measurements,

$y_1 = 27.1, y_2 = 14.6, y_3 = 14.6$

$\sigma^2 = 20$

```{r}
n <- 200

x <- rnorm(n, mean = 17, sd = .5)
triplot.normal.knownvariance(n=n, theta.data = x, variance.known = 20, prior.theta = 0, prior.variance = 17)
```

### Estimating Mean and Variance using Simulating

```{r}
# Simulate hypothetical body height measurements

true.mean <- 165
true.sd <- 10
y <- round(rnorm(10, mean = true.mean, sd = true.sd))

mod <- lm(y ~ 1)
mod

summary(mod)$sigma
```

$y \sim Norm(\hat{\theta} = 166.3, \sigma^2 = 13.3)$

```{r}
nsim <- 5000
bsim <- sim(mod, n.sim=nsim)

str(bsim)

str(bsim, max.level = 2)

plot_posterior(bsim)

quantile(bsim@coef, prob = c(.025, 0.5, 0.975))
quantile(bsim@sigma, prob = c(0.025, 0.5, 0.975))

HPDinterval(as.mcmc(bsim@coef))

sum(bsim@coef > 160) / nsim
mean(bsim@coef > 160)

cvsim <- bsim@sigma / bsim@coef
quantile(cvsim, prob = c(0.025, 0.5, 0.975))
```

### The Frequentist Way

```{r}

```